### 2019-03-26 paper/main.tex
Init repository

\documentclass[12pt]{book}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage[a4paper, margin=25mm]{geometry}
\usepackage{fancyhdr}
\usepackage{mathtools}
\usepackage{hyperref}
\usepackage{float}
\pagestyle{fancy}
\fancyhead{}
\fancyhead[L]{Open digital signature software}
\fancyhead[R]{Jakub Ďuraš}
\fancyfoot{}
\fancyfoot[C]{\thepage}
\graphicspath{{./}}

\title{Open digital signature software}
\author{Jakub Ďuraš}

\begin{document}

\maketitle

\end{document}

### 2019-04-07 paper/main.tex
Update paper template

@@
-\documentclass[12pt]{book}
+\documentclass[thesismargins, english, thesislinespacing, twoside, openright, upjsfrontpage]{rnthesis}
+\usepackage[english]{babel}
+\usepackage[T1]{fontenc}
 \usepackage[utf8]{inputenc}
-\usepackage{graphicx}
-\usepackage[a4paper, margin=25mm]{geometry}
-\usepackage{fancyhdr}
-\usepackage{mathtools}
-\usepackage{hyperref}
-\usepackage{float}
-\pagestyle{fancy}
-\fancyhead{}
-\fancyhead[L]{Open digital signature software}
-\fancyhead[R]{Jakub Ďuraš}
-\fancyfoot{}
-\fancyfoot[C]{\thepage}
-\graphicspath{{./}}
+\usepackage{lmodern}
 
-\title{Open digital signature software}
+\usepackage{rnt-pic}
+\usepackage{rnt-thm}
+
+\usepackage{pdfpages} % inserting pdf pages
+
+\usepackage[hyphens]{url} % format and linebreak of URLs
+
+\title{Open electronic signature software}
 \author{Jakub Ďuraš}
+\typprace{bachelor's}
+\rok{2020}
+\miesto{Košice}
+\podakovanie{
+  Thanks to the Planet!
+}
+\veduci{RNDr. Viliam Kačala}
+\pracovisko{Institute of Computer Science}
+\pdfzadanie{zadanie.pdf}
+
+\abstract{
+With the recent changes in the legal status of electronic signatures in many parts of the world, there is a need for easily accessible solutions intended as an alternative to handwritten signatures.
+This may be more necessary than ever since electronic communication is the preferred way of communication.
+This bachelor thesis aims to explore the principles and legal status of electronic signatures, review current electronic signature software, and explore possible obstacles the open-source community is facing to develop such specialized applications.
+We propose an open-source, cross-platform, and user-friendly software compliant with the eIDAS Regulation (Regulation No 910/2014).
+Our application should allow ordinary users to quickly sign and verify signatures of different types of documents and therefore easily use them in everyday life.
+}
+
+\keywords{electronic signature, electronic seal, qualified, open-source, XAdES, PAdES, CAdES, desktop software}
+
+\abstrakt{
+Abstrakt obsahuje informáciu o cieľoch práce, jej stručnom obsahu,
+výsledkoch a význame celej práce. Súčasťou abstraktu je 3--5 kľúčových 
+slov. Abstrakt sa píše súvisle ako jeden odsek a jeho rozsah je 
+spravidla 100 až 500 slov.
+}
+
+\klucoveslova{lorem ipsum, dolor, sit amet}
+
+\bibliographystyle{alpha}
 
 \begin{document}
+\maketitle
+\newpage
+\tableofcontents
+\listoffigures
+\listoftables
+% zoznam značiek a skratiek. Pre tento koncept neexistuje
+% LaTeXovsky príkaz
 
-\maketitle
+\uvod
+
+Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
+Integer lacinia, nulla porta varius tempus, lacus metus blandit
+lorem, a rutrum justo wisi id sapien. Integer risus libero,
+feugiat eleifend, ornare ac, volutpat nec, sem. In facilisis,
+quam eu elementum aliquet, lorem quam euismod dui, aliquet
+laoreet purus ipsum ac quam.
+
+\chapter{A Traditional Lipsum}
+
+Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
+Integer lacinia, nulla porta varius tempus, lacus metus blandit
+lorem, a rutrum justo wisi id sapien. Integer risus libero,
+feugiat eleifend, ornare ac, volutpat nec, sem. In facilisis,
+quam eu elementum aliquet, lorem quam euismod dui, aliquet
+laoreet purus ipsum ac quam. 
+%
+$$\int\!\!\!\int x^y\,{\rm d}x\,{\rm d}y = \int\!\!\!\int (1 + x) - (\sin^2 x + \cos^2 x)^y\,{\rm d}x\,{\rm d}y$$
+
+Donec dolor arcu, posuere at, vehicula vitae, accumsan ut,
+lacus. Nulla tristique eros eu diam. Vivamus nec tortor vel
+ligula elementum lacinia. Curabitur euismod eros adipiscing
+ipsum. Donec sed quam at felis suscipit egestas. Morbi faucibus
+libero sit amet libero. Nullam laoreet ipsum eu eros. Donec in
+diam. Ut facilisis eros vel leo. Nunc vitae mauris. Donec leo
+erat, luctus porttitor, laoreet eget, facilisis non, erat.
+Integer nec elit.
+
+\section{Dolor sit amet}
+
+Turabitur condimentum. Quisque ut risus. Vestibulum non arcu a
+est feugiat porttitor. Lorem ipsum dolor sit amet, consectetuer
+adipiscing elit. Duis in metus. 
+%
+$$\forall \epsilon>0 \exists \delta>0:\forall x\in(-\delta,\delta):|f(x)-\epsilon|<\delta$$
+%
+Integer ligula sapien, rutrum et, pulvinar ac, viverra a,
+neque. Suspendisse hendrerit lectus id ante. Fusce mattis nunc
+non ipsum. Praesent tristique hendrerit lorem. Morbi risus
+erat, euismod quis.
+
+\section{Sit transit gloria mundi}
+\begin{df}
+\pojem{Definíciou} nazývame každý pojem uzavretý v prostredí 
+\texttt{df}.
+\end{df}
+
+Autorom predchádzajúcej definície je \osoba{František Taraba},
+ktorý už v roku fusce elit enim, commodo eget, blandit eu,
+faucibus sed, nisl. Maecenas adipiscing. Proin non risus in
+erat dapibus hendrerit. Sed nonummy, velit sit amet dictum
+eleifend, arcu lacus molestie mauris, a sollicitudin felis
+lacus eget velit. Vivamus imperdiet. Vivamus accumsan
+sollicitudin leo. Morbi et orci. Sed at sem. Vestibulum mattis
+augue. Nam sit amet wisi. Donec vel est. Praesent vehicula.
+Nunc convallis hendrerit nisl.
+%
+$$m^2=n_1^2+n_2^2$$
+
+\begin{veta}[Cauchy's Integral Theorem]
+Let $U$ be an open subset of $C$ which is simply connected, let $f : U \to C$ be a holomorphic function, and let $\gamma$ be a rectifiable path in $U$ whose start point is equal to its end point. Then
+
+$$ \oint_{\gamma }f(z)\,dz=0.$$.
+\end{veta}
+%
+\begin{dokaz}
+Dôkaz prenechávame na pozorného čitateľa.
+\end{dokaz}
+%
+\begin{dokaz}
+A teraz:
+$$\pmatrix{3 & 3\cr 1 & 2} \cdot 3$$
+\end{dokaz}
+Dôkaz prenechávame na pozorného čitateľa. Už \cite{knuth84} ukázal 
+použitie tejto metódy. Na druhej strane, \cite{lamport86} použil
+alternatívny prístup.\footnote{Donec venenatis rutrum odio. Fusce porta. Curabitur
+a ipsum sit amet arcu semper posuere. Donec orci felis, auctor sit
+amet, semper non, suscipit vel, tellus. Sed eu ipsum nec mi egestas
+ultrices. Sed mauris. Aliquam purus.}
+
+Suspendisse lobortis. Donec ornare elit sit amet nibh. Mauris nec
+augue. Sed dignissim dictum mauris. Morbi tincidunt leo at est.
+Pellentesque habitant morbi tristique senectus et netus et
+malesuada fames ac turpis egestas. Etiam eu ante in sem dictum
+eleifend. Nunc eget odio. Sed vitae elit at justo tristique
+dapibus. Quisque sit amet urna at velit faucibus cursus,
+ako bolo ukázané v \cite{1}.
+
+Aliquam commodo wisi sed ipsum. Donec quis ligula. Ut porttitor,
+nibh nec interdum fringilla, risus est nonummy nibh, at rutrum
+massa odio molestie dolor. Morbi metus. Nulla nec velit vitae elit
+nonummy lobortis. Phasellus facilisis, urna ac viverra tristique,
+tellus turpis commodo dui, id pharetra erat nibh et mi. Mauris
+iaculis nisl sit amet tellus molestie fringilla.
+
+Duis ligula lectus, condimentum in, lacinia eget, pellentesque sit
+amet, mi. Aliquam convallis euismod arcu. Aliquam erat volutpat.
+Fusce erat elit, congue eu, pretium in, congue non, neque.
+Vestibulum non massa eu nisl condimentum blandit. Integer pretium
+wisi id metus. Donec venenatis rutrum odio. Fusce porta. Curabitur
+a ipsum sit amet arcu semper posuere. Donec orci felis, auctor sit
+amet, semper non, suscipit vel, tellus. Sed eu ipsum nec mi egestas
+ultrices. Sed mauris. Aliquam purus. In hac habitasse platea
+dictumst. Phasellus ut urna. Etiam sit amet ligula ultrices massa
+iaculis suscipit. Integer odio lacus, interdum eget, tempor eu,
+aliquam nec, elit.
+
+\begin{table}
+\begin{center}
+\begin{tabular}{ccc}
+Level & Requirement\\
+1 & 0\\
+2 & 1000\\
+3 & 3000\\
+4 & 6000
+\end{tabular}
+\end{center}
+\caption{Requirements of XP for given level.}
+\end{table}
+Suspendisse lobortis. Donec ornare elit sit amet nibh. Mauris nec
+augue. Sed dignissim dictum mauris. Morbi tincidunt leo at est.
+Pellentesque habitant morbi tristique senectus et netus et
+malesuada fames ac turpis egestas. Etiam eu ante in sem dictum
+eleifend. Nunc eget odio. Sed vitae elit at justo tristique
+dapibus. Quisque sit amet urna at velit faucibus cursus.
+Morbi metus. Nulla nec velit vitae elit
+nonummy lobortis. Phasellus facilisis, urna ac viverra tristique,
+tellus turpis commodo dui, id pharetra erat nibh et mi. Mauris
+iaculis nisl sit amet tellus molestie fringilla. Sed nonummy mollis
+dui. Sed fermentum suscipit metus. Curabitur quis wisi. In
+condimentum pellentesque ante. Integer in odio. Curabitur arcu
+metus, convallis eu, euismod in, ultricies nec, felis. Suspendisse
+hendrerit ipsum ac neque pretium consectetuer. Vestibulum et diam
+vitae nulla faucibus mattis.
+
+\chapter{Contemporary Pseudo-Random Texts}
+
+Nullam neque felis, accumsan a cursus at, fringilla a massa.
+Praesent suscipit purus eget enim volutpat, bibendum imperdiet odio convallis.
+Vivamus congue ultricies dapibus.
+Sed eu pretium mauris, consequat scelerisque ex.
+Vestibulum ut sem venenatis, pharetra orci eu, tempus metus.
+Ut auctor porta tincidunt.
+Interdum et malesuada fames ac ante ipsum primis in faucibus.
+Sed eget risus fringilla, volutpat mauris non, cursus ex.
+Donec faucibus ante et nisl sodales, vel venenatis turpis rhoncus.
+Nunc eget congue elit.
+In placerat leo in libero hendrerit maximus.
+Vestibulum efficitur aliquet sapien blandit semper.
+Nulla est diam, fermentum a commodo id, ullamcorper vitae tellus.
+Aenean scelerisque fringilla velit id condimentum.
+Ut vel libero turpis.
+
+\begin{equation}
+P(X_n = x_n | X_{n - 1} = x_{n - 1}, X_{n - 2} = x_{n - 2}, \dots, X_0 = x_0) = P(X_n = x_n | X_{n - 1} = x_{n - 1})
+\end{equation}
+
+Sed ornare ac mauris semper efficitur.
+Aenean id sollicitudin augue.
+Quisque at efficitur libero, at scelerisque arcu.
+Suspendisse consectetur luctus posuere.
+Cras vel nisl non neque feugiat tempor.
+Aliquam dapibus, mi eget fringilla rutrum, ligula nulla mattis sem, in pharetra augue neque quis ipsum.
+Cras bibendum quis lacus eget interdum.
+Pellentesque facilisis auctor magna, vitae mattis metus fringilla in.
+Ut tempor vel tellus sed ultricies.
+Donec accumsan mi ac fermentum cursus.
+Donec congue, arcu et scelerisque ornare, velit leo placerat augue, in laoreet velit odio quis odio.
+Proin tempor gravida mauris.
+Nulla facilisi.
+Nam pellentesque porttitor enim non malesuada.
+
+Nullam lacus turpis, tempus et nisi non, dignissim tincidunt urna.
+Praesent vehicula dui urna, ac ullamcorper justo laoreet in.
+Pellentesque at arcu id risus ultricies consequat ut in dolor.
+Duis posuere, enim at ullamcorper facilisis, dolor erat auctor augue, vitae congue ipsum velit vitae nibh.
+Maecenas tincidunt magna luctus felis tempus blandit.
+Pellentesque mattis lorem lacus, pretium egestas tellus maximus in.
+Donec ligula felis, laoreet non tortor id, ultrices lobortis arcu.
+Proin in fermentum libero, vitae ultricies nibh.
+In non euismod sapien, ullamcorper luctus tellus.
+Donec nibh massa, malesuada a metus ac, molestie consectetur purus.
+Morbi elementum molestie libero, quis condimentum ipsum.
+
+Ut lobortis semper risus, non condimentum dui convallis ut.
+Nulla eget volutpat tellus.
+Vestibulum lobortis tincidunt massa eu rhoncus.
+Suspendisse luctus eu dui non vehicula.
+Vivamus elementum auctor felis, placerat maximus magna lobortis ut.
+Donec placerat sem a mi sagittis blandit.
+Maecenas pellentesque laoreet mauris, dictum viverra ante sodales sed.
+Nullam non ligula quis ante ultricies finibus non quis ex.
+Ut tempor vitae ipsum sed imperdiet.
+Fusce aliquam nisl sit amet nunc tempor convallis.
+Vivamus vehicula magna sit amet purus commodo, et tincidunt purus accumsan.
+Nulla velit dolor, lacinia nec scelerisque a, euismod a sapien.
+
+\zaver
+
+Ut lobortis semper risus, non condimentum dui convallis ut.
+Nulla eget volutpat tellus.
+Vestibulum lobortis tincidunt massa eu rhoncus.
+Suspendisse luctus eu dui non vehicula.
+Vivamus elementum auctor felis, placerat maximus magna lobortis ut.
+Donec placerat sem a mi sagittis blandit.
+Maecenas pellentesque laoreet mauris, dictum viverra ante sodales sed.
+Nullam non ligula quis ante ultricies finibus non quis ex.
+Ut tempor vitae ipsum sed imperdiet.
+Fusce aliquam nisl sit amet nunc tempor convallis.
+Vivamus vehicula magna sit amet purus commodo, et tincidunt purus accumsan.
+Nulla velit dolor, lacinia nec scelerisque a, euismod a sapien.
+%
+
+%\renewcommand{\bibname}{Zoznam použitej literatúry}
+\begin{thebibliography}{9}
+% Príklady popisu dokumentov citácií podľa systému meno a dátum (Harvardský systém)
+% ----
+% Varianty zápisov autorov:
+% [1] GUZANIN, Štefan, Robert SABOVČÍK a Pavol KAČMÁR. Priezviská vždy VEĽKÝMI PÍSMENAMI,
+%   priezvisko prvého autora je vždy pred menom, druhý a ďalší autor majú zápis
+%   Meno PRIEZVISKO
+% [2] Neuvádzať rodné mená autorov.
+% [3] Verzálky nie sú povinné, možno použiť aj iné indikatívnejšie označenie
+%
+% --- 
+% 1. Knižná publikácia (monografia, učebnica, zborník ...)
+%   1 autor
+\bibitem{2}
+  \osoba{Beck, G.}, 2007. \emph{Zakázaná rétorika: 30 manipulatívních technik}. Preklad
+\osoba{Pomikálová, M.}. Praha: Grada Publishing. ISBN 978-80-247-1743-2.
+\bibitem{3}
+  \osoba{Vojčík, P.}, 2010. \emph{Občianske právo hmotné II.} 3. prep. a dopl. vyd. Košice: UPJŠ v Košiciach. ISBN 978-80-7097-817-7.
+%  2 autori
+\bibitem{4}
+  \osoba{Šoltés, M.} a \osoba{Radoňák, J.}, 2013. \emph{Základné princípy laparoskopickej chirurgie.} Košice: UPJŠ v Košiciach. ISBN 978-80-8152-074-7.
+% 3 autori
+\bibitem{5}
+  \osoba{Guzanin, Š.}, \osoba{Sabovčík, R.} a \osoba{Kačmár, P.}, 2004.
+\emph{Selected Chapters of Plastic and Reconstructive Surgery: vysokoškolské učebné texty}.
+Košice: Univerzita Pavla Jozefa Šafárika v Košiciach, Lekárska fakulta.
+ISBN 80-7097-557-1.
+% 4 a viac autorov
+\bibitem{6}
+  \osoba{Nagyová, I.} et al. 2009. \emph{Measuring health and quality of life in the chronically ill}. Košice: Equilibria. ISBN 978-80-892-8446-7.
+% Elektronická kniha
+\bibitem{7}
+  \osoba{Speight, J.
+G.}, 2005.
+\emph{Lange's Handbook of Chemistry} [online].
+London: McGraw-Hill.
+[cit.
+2009.06.10.] ISBN 978-1-60119-261-5.
+Dostupné na: \url{http://www.knovel.com/web/portal/basic_search/display?_EXT_KNOVEL_DISPLAY_bookid=1347&_EXT_KNOVEL_DISPLAY_fromSearch=true&_EXT_KNOVEL_DISPLAY_searchType=basic}
+% zborník
+\bibitem{8}
+  \osoba{Bačkor, M.} a \osoba{Mihaličová, S.}, zost., 2013.
+\emph{Zborník príspevkov z konferencie 11.
+dni doktorandov experimentálnej biológie rastlín a 13.
+konferencie experimentálnej biológie rastlín} [online].
+Košice: Univerzita Pavla Jozefa Šafárika v Košiciach, Prírodovedecká fakulta [cit.
+2009-06-10].
+ISBN 9788081520327.
+Dostupné na: \url{
+http://www.upjs.sk/public/media/5596/PF-Zbornik-prispevkov-konferencie-11-dni-doktorandov.pdf}
+%
+% 2. Časopis (ako celok)
+\bibitem{9}
+  \emph{Thaiszia: Journal of Botany}. Košice: P.J.Safarik University, Botanic Garden, \mbox{1990--\ .} ISSN 1210-0420.
+\bibitem{10}
+  \emph{Ikaros: elektronický časopis o informační bezpečnosti} [online], 2002. [Praha]: Ikaros. 1997--{} [cit. 2002-03-08]. Dostupné na: \url{http://www.ikaros.cz/}. ISSN 1212-5075.
+% Jedno číslo časopisu
+\bibitem{11}
+  \emph{CHIP: magazín informačních technologií}, 2013. Praha: Burda Praha, roč. 23, říjen. ISSN 1210-0684.
+%  ------
+%  3. Príspevok v knihe/zborníku
+%  ------
+\bibitem{12}
+  \osoba{Sabol, J.}, 2000. Jazyk ako ľudské posolstvo: (namiesto doslovu). In: \emph{O jazyku a štýle kriticky aj prakticky}. Prešov: Náuka, s. 149--159. ISBN 809676022X.
+\bibitem{13}
+  \osoba{Tóthová, E.} a kol., 2013.
+A rare t(9,22,16)(q34,q11,q24) translocation in chronic myeloid leukemia for which imatinib mesylate was effective: a case report.
+In: \emph{XXVII.
+Olomoucké hematologické dny s mezinárodní účastí, 12.--14.5.2013, Olomouc: sborník abstrakt}.
+Olomouc: Univerzita Palackého v Olomouci, s.
+75--76.
+ISBN 9788024434803.
+%  ------
+% 4. Článok v časopise
+%  ------
+\bibitem{14}
+  \osoba{Beňačka, J.} et al., 2009.
+A better cosine approximate solution to pendulum equation.
+In: \emph{International Journal of Mathematical Education in Science and Technology}.
+Vol.
+40, no.
+2, p.
+206--215.
+ISSN 0020-739X.
+\bibitem{15} 
+  \osoba{Dubayová, T.} et al., 2010.
+The impact of the intensity of fear on patient's delay regarding health care seeking behavior: a systematic review vyhľadaní zdravotníckej starostlivosti.
+In: \emph{International Journal of Public Health}.
+Vol.
+55, no.
+5, p.
+459--468.
+ISSN 1661-8556.
+\bibitem{16} 
+  \osoba{Steinerová, J.}, 2000. Princípy formovania vzdelania v informačnej vede. In: \emph{Pedagogická revue}. Roč. 2, č. 3, s. 8--16. ISSN 1335-1982.
+\bibitem{17}
+  \osoba{Hoggan, D.}, 2002.
+Challenges, Strategies, and Tools for Research Scientists.
+In: \emph{Electronic Journal of Academic and Special Librarianship} [online].
+Vol.
+3, no.
+3 [cit.
+2013-01-10].
+ISSN 1525-321X.
+Dostupné na: \url{http://southernlibrarianship.icaap.org/content/v03n03/Hoggan_d01.htm}
+\bibitem{18}
+  \osoba{Srbecká, Gabriela}, 2010.
+Rozvoj kompetencí studentů ve vzdělávání.
+In: \emph{Inflow: information journal} [online].
+Roč.
+3, č.
+7 [cit.
+2013-08-06].
+ISSN 1802-9736.
+Dostupné na: \url{http://www.inflow.cz/rozvoj-kompetenci-studentu-ve-vzdelavani}
+%  ------
+% 5. Príspevok v zborníku na CD-ROM
+%  ------
+\bibitem{19}
+  \osoba{Zemánek, P.}, 2001.
+The machines for ``green works'' in vineyards and their economical evaluation.
+In: \emph{9th International Conference: proceedings.
+Vol.
+2.
+Fruit Growing and viticulture} [CD-ROM].
+Lednice: Mendel University of Agriculture and Forestry, p.
+262--268.
+ISBN 80-7157-524-0.
+%  ------
+% 6. Záverečné a kvalifikačné práce
+%  ------
+\bibitem{20} 
+  \osoba{Mikulášiková, M.}, 1999. \emph{Didaktické pomôcky pre praktickú výučbu na hodinách výtvarnej výchovy pre 2. stupeň základných škôl}: diplomová práca. Nitra: UKF.
+\bibitem{21} 
+  \osoba{Urdzík, P.}, 2007. \emph{Predikcia intrauterinnej rastovej retardácie a preeklampsie pomocou biochemických a ultrazvukových markerov}: dizertačná práca. Košice: UPJŠ v Košiciach.
+%  ------
+% 7. Výskumné správy
+%  ------
+\bibitem{22} 
+  \osoba{Baumgartner, J.} a kol., 1998. \emph{Ochrana a udržiavanie genofondu zvierat, šľachtenie zvierat}: výskumná správa. Nitra: VÚŽV.
+%  ------
+% 8. Normy
+%  ------
+\bibitem{23}
+  STN ISO 690: 2012. \emph{Informácie a dokumentácia. Návod na tvorbu bibliografických odkazov na informačné pramene a ich citovanie}.
+%  ------
+% 9. Mapa
+%  ------
+\bibitem{24} 
+  VKÚ, 2003. \emph{Košice: mapa okolia}. [1:15000]. 3. vyd. Harmanec: VKÚ. ISBN 80-8042-223-0.
+%  ------
+% 10. Zákon
+%  ------
+\bibitem{25} 
+  \emph{Zákon č.131/2002 Zb. o vysokých školách a o zmene a doplnení niektorých zákonov.}
+\bibitem{26} 
+  \emph{Zákon č. 313/2001 o verejnej službe.}
+
+\end{thebibliography}
 

### 2019-12-05 paper/main.tex
Add structure to the thesis, add link

@@
 \rok{2020}
+\odbor{Applied Informatics}
 \miesto{Košice}
@@
 \abstrakt{
-Abstrakt obsahuje informáciu o cieľoch práce, jej stručnom obsahu,
-výsledkoch a význame celej práce. Súčasťou abstraktu je 3--5 kľúčových 
-slov. Abstrakt sa píše súvisle ako jeden odsek a jeho rozsah je 
-spravidla 100 až 500 slov.
+Abstrakt v SK jazyku bude pridaný neskôr ako voľný preklad z EN verzie (pozri ďalšiu stranu). Predíde sa tak zbytočnej práci keďže EN abstrakt sa môže časom meniť.
 }
 
-\klucoveslova{lorem ipsum, dolor, sit amet}
+\klucoveslova{elektronický podpis, elektronická pečať, kvalifikovaný, open-source, XAdES, PAdES, CAdES, počítačový softvér}
 
@@
 
-\chapter{A Traditional Lipsum}
+\chapter{Background and theory}
 
@@
 laoreet purus ipsum ac quam.
-%
-$$\int\!\!\!\int x^y\,{\rm d}x\,{\rm d}y = \int\!\!\!\int (1 + x) - (\sin^2 x + \cos^2 x)^y\,{\rm d}x\,{\rm d}y$$
+
+\section{Legal}
+
+-
+
+\section{Cryptography}
+
+-
+
+\section{Software engineering}
+
+-
+
+\chapter{Problem setting}
+
+Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
+Integer lacinia, nulla porta varius tempus, lacus metus blandit
+lorem, a rutrum justo wisi id sapien. Integer risus libero,
+feugiat eleifend, ornare ac, volutpat nec, sem. In facilisis,
+quam eu elementum aliquet, lorem quam euismod dui, aliquet
+laoreet purus ipsum ac quam.
+
+\chapter{Analysis}
+
+Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
+Integer lacinia, nulla porta varius tempus, lacus metus blandit
+lorem, a rutrum justo wisi id sapien. Integer risus libero,
+feugiat eleifend, ornare ac, volutpat nec, sem. In facilisis,
+quam eu elementum aliquet, lorem quam euismod dui, aliquet
+laoreet purus ipsum ac quam.
+
+\chapter{Implementation}
+
+Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
+Integer lacinia, nulla porta varius tempus, lacus metus blandit
+lorem, a rutrum justo wisi id sapien. Integer risus libero,
+feugiat eleifend, ornare ac, volutpat nec, sem. In facilisis,
+quam eu elementum aliquet, lorem quam euismod dui, aliquet
+laoreet purus ipsum ac quam.
+
+\chapter{Results}
+
+Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
+Integer lacinia, nulla porta varius tempus, lacus metus blandit
+lorem, a rutrum justo wisi id sapien. Integer risus libero,
+feugiat eleifend, ornare ac, volutpat nec, sem. In facilisis,
+quam eu elementum aliquet, lorem quam euismod dui, aliquet
+laoreet purus ipsum ac quam.
 
@@
 Integer nec elit.
-
-\section{Dolor sit amet}
-
-Turabitur condimentum. Quisque ut risus. Vestibulum non arcu a
-est feugiat porttitor. Lorem ipsum dolor sit amet, consectetuer
-adipiscing elit. Duis in metus. 
-%
-$$\forall \epsilon>0 \exists \delta>0:\forall x\in(-\delta,\delta):|f(x)-\epsilon|<\delta$$
-%
-Integer ligula sapien, rutrum et, pulvinar ac, viverra a,
-neque. Suspendisse hendrerit lectus id ante. Fusce mattis nunc
-non ipsum. Praesent tristique hendrerit lorem. Morbi risus
-erat, euismod quis.
-
-\section{Sit transit gloria mundi}
-\begin{df}
-\pojem{Definíciou} nazývame každý pojem uzavretý v prostredí 
-\texttt{df}.
-\end{df}
-
-Autorom predchádzajúcej definície je \osoba{František Taraba},
-ktorý už v roku fusce elit enim, commodo eget, blandit eu,
-faucibus sed, nisl. Maecenas adipiscing. Proin non risus in
-erat dapibus hendrerit. Sed nonummy, velit sit amet dictum
-eleifend, arcu lacus molestie mauris, a sollicitudin felis
-lacus eget velit. Vivamus imperdiet. Vivamus accumsan
-sollicitudin leo. Morbi et orci. Sed at sem. Vestibulum mattis
-augue. Nam sit amet wisi. Donec vel est. Praesent vehicula.
-Nunc convallis hendrerit nisl.
-%
-$$m^2=n_1^2+n_2^2$$
-
-\begin{veta}[Cauchy's Integral Theorem]
-Let $U$ be an open subset of $C$ which is simply connected, let $f : U \to C$ be a holomorphic function, and let $\gamma$ be a rectifiable path in $U$ whose start point is equal to its end point. Then
-
-$$ \oint_{\gamma }f(z)\,dz=0.$$.
-\end{veta}
-%
-\begin{dokaz}
-Dôkaz prenechávame na pozorného čitateľa.
-\end{dokaz}
-%
-\begin{dokaz}
-A teraz:
-$$\pmatrix{3 & 3\cr 1 & 2} \cdot 3$$
-\end{dokaz}
-Dôkaz prenechávame na pozorného čitateľa. Už \cite{knuth84} ukázal 
-použitie tejto metódy. Na druhej strane, \cite{lamport86} použil
-alternatívny prístup.\footnote{Donec venenatis rutrum odio. Fusce porta. Curabitur
-a ipsum sit amet arcu semper posuere. Donec orci felis, auctor sit
-amet, semper non, suscipit vel, tellus. Sed eu ipsum nec mi egestas
-ultrices. Sed mauris. Aliquam purus.}
-
-Suspendisse lobortis. Donec ornare elit sit amet nibh. Mauris nec
-augue. Sed dignissim dictum mauris. Morbi tincidunt leo at est.
-Pellentesque habitant morbi tristique senectus et netus et
-malesuada fames ac turpis egestas. Etiam eu ante in sem dictum
-eleifend. Nunc eget odio. Sed vitae elit at justo tristique
-dapibus. Quisque sit amet urna at velit faucibus cursus,
-ako bolo ukázané v \cite{1}.
-
-Aliquam commodo wisi sed ipsum. Donec quis ligula. Ut porttitor,
-nibh nec interdum fringilla, risus est nonummy nibh, at rutrum
-massa odio molestie dolor. Morbi metus. Nulla nec velit vitae elit
-nonummy lobortis. Phasellus facilisis, urna ac viverra tristique,
-tellus turpis commodo dui, id pharetra erat nibh et mi. Mauris
-iaculis nisl sit amet tellus molestie fringilla.
-
-Duis ligula lectus, condimentum in, lacinia eget, pellentesque sit
-amet, mi. Aliquam convallis euismod arcu. Aliquam erat volutpat.
-Fusce erat elit, congue eu, pretium in, congue non, neque.
-Vestibulum non massa eu nisl condimentum blandit. Integer pretium
-wisi id metus. Donec venenatis rutrum odio. Fusce porta. Curabitur
-a ipsum sit amet arcu semper posuere. Donec orci felis, auctor sit
-amet, semper non, suscipit vel, tellus. Sed eu ipsum nec mi egestas
-ultrices. Sed mauris. Aliquam purus. In hac habitasse platea
-dictumst. Phasellus ut urna. Etiam sit amet ligula ultrices massa
-iaculis suscipit. Integer odio lacus, interdum eget, tempor eu,
-aliquam nec, elit.
-
-\begin{table}
-\begin{center}
-\begin{tabular}{ccc}
-Level & Requirement\\
-1 & 0\\
-2 & 1000\\
-3 & 3000\\
-4 & 6000
-\end{tabular}
-\end{center}
-\caption{Requirements of XP for given level.}
-\end{table}
-Suspendisse lobortis. Donec ornare elit sit amet nibh. Mauris nec
-augue. Sed dignissim dictum mauris. Morbi tincidunt leo at est.
-Pellentesque habitant morbi tristique senectus et netus et
-malesuada fames ac turpis egestas. Etiam eu ante in sem dictum
-eleifend. Nunc eget odio. Sed vitae elit at justo tristique
-dapibus. Quisque sit amet urna at velit faucibus cursus.
-Morbi metus. Nulla nec velit vitae elit
-nonummy lobortis. Phasellus facilisis, urna ac viverra tristique,
-tellus turpis commodo dui, id pharetra erat nibh et mi. Mauris
-iaculis nisl sit amet tellus molestie fringilla. Sed nonummy mollis
-dui. Sed fermentum suscipit metus. Curabitur quis wisi. In
-condimentum pellentesque ante. Integer in odio. Curabitur arcu
-metus, convallis eu, euismod in, ultricies nec, felis. Suspendisse
-hendrerit ipsum ac neque pretium consectetuer. Vestibulum et diam
-vitae nulla faucibus mattis.
-
-\chapter{Contemporary Pseudo-Random Texts}
-
-Nullam neque felis, accumsan a cursus at, fringilla a massa.
-Praesent suscipit purus eget enim volutpat, bibendum imperdiet odio convallis.
-Vivamus congue ultricies dapibus.
-Sed eu pretium mauris, consequat scelerisque ex.
-Vestibulum ut sem venenatis, pharetra orci eu, tempus metus.
-Ut auctor porta tincidunt.
-Interdum et malesuada fames ac ante ipsum primis in faucibus.
-Sed eget risus fringilla, volutpat mauris non, cursus ex.
-Donec faucibus ante et nisl sodales, vel venenatis turpis rhoncus.
-Nunc eget congue elit.
-In placerat leo in libero hendrerit maximus.
-Vestibulum efficitur aliquet sapien blandit semper.
-Nulla est diam, fermentum a commodo id, ullamcorper vitae tellus.
-Aenean scelerisque fringilla velit id condimentum.
-Ut vel libero turpis.
-
-\begin{equation}
-P(X_n = x_n | X_{n - 1} = x_{n - 1}, X_{n - 2} = x_{n - 2}, \dots, X_0 = x_0) = P(X_n = x_n | X_{n - 1} = x_{n - 1})
-\end{equation}
-
-Sed ornare ac mauris semper efficitur.
-Aenean id sollicitudin augue.
-Quisque at efficitur libero, at scelerisque arcu.
-Suspendisse consectetur luctus posuere.
-Cras vel nisl non neque feugiat tempor.
-Aliquam dapibus, mi eget fringilla rutrum, ligula nulla mattis sem, in pharetra augue neque quis ipsum.
-Cras bibendum quis lacus eget interdum.
-Pellentesque facilisis auctor magna, vitae mattis metus fringilla in.
-Ut tempor vel tellus sed ultricies.
-Donec accumsan mi ac fermentum cursus.
-Donec congue, arcu et scelerisque ornare, velit leo placerat augue, in laoreet velit odio quis odio.
-Proin tempor gravida mauris.
-Nulla facilisi.
-Nam pellentesque porttitor enim non malesuada.
-
-Nullam lacus turpis, tempus et nisi non, dignissim tincidunt urna.
-Praesent vehicula dui urna, ac ullamcorper justo laoreet in.
-Pellentesque at arcu id risus ultricies consequat ut in dolor.
-Duis posuere, enim at ullamcorper facilisis, dolor erat auctor augue, vitae congue ipsum velit vitae nibh.
-Maecenas tincidunt magna luctus felis tempus blandit.
-Pellentesque mattis lorem lacus, pretium egestas tellus maximus in.
-Donec ligula felis, laoreet non tortor id, ultrices lobortis arcu.
-Proin in fermentum libero, vitae ultricies nibh.
-In non euismod sapien, ullamcorper luctus tellus.
-Donec nibh massa, malesuada a metus ac, molestie consectetur purus.
-Morbi elementum molestie libero, quis condimentum ipsum.
-
-Ut lobortis semper risus, non condimentum dui convallis ut.
-Nulla eget volutpat tellus.
-Vestibulum lobortis tincidunt massa eu rhoncus.
-Suspendisse luctus eu dui non vehicula.
-Vivamus elementum auctor felis, placerat maximus magna lobortis ut.
-Donec placerat sem a mi sagittis blandit.
-Maecenas pellentesque laoreet mauris, dictum viverra ante sodales sed.
-Nullam non ligula quis ante ultricies finibus non quis ex.
-Ut tempor vitae ipsum sed imperdiet.
-Fusce aliquam nisl sit amet nunc tempor convallis.
-Vivamus vehicula magna sit amet purus commodo, et tincidunt purus accumsan.
-Nulla velit dolor, lacinia nec scelerisque a, euismod a sapien.
 

### 2020-01-26 paper/main.tex
Add beginning of the Background & Theory

--- 2019-03-26 paper/main.tex
@@
-\documentclass[12pt]{book}
+\documentclass[thesismargins, english, thesislinespacing, twoside, openright, upjsfrontpage]{rnthesis}
+\usepackage[english]{babel}
+\usepackage[T1]{fontenc}
 \usepackage[utf8]{inputenc}
-\usepackage{graphicx}
-\usepackage[a4paper, margin=25mm]{geometry}
-\usepackage{fancyhdr}
-\usepackage{mathtools}
-\usepackage{hyperref}
-\usepackage{float}
-\pagestyle{fancy}
-\fancyhead{}
-\fancyhead[L]{Open digital signature software}
-\fancyhead[R]{Jakub Ďuraš}
-\fancyfoot{}
-\fancyfoot[C]{\thepage}
-\graphicspath{{./}}
+\usepackage{lmodern}
 
-\title{Open digital signature software}
+\usepackage{rnt-pic}
+\usepackage{rnt-thm}
+
+\usepackage{pdfpages} % inserting pdf pages
+
+\usepackage[hyphens]{url} % format and linebreak of URLs
+
+\title{Open electronic signature software}
 \author{Jakub Ďuraš}
+\typprace{bachelor's}
+\rok{2020}
+\odbor{Applied Informatics}
+\miesto{Košice}
+\podakovanie{
+  Thanks to the Planet!
+}
+\veduci{RNDr. Viliam Kačala}
+\pracovisko{Institute of Computer Science}
+\pdfzadanie{zadanie.pdf}
+
+\abstract{
+With the recent changes in the legal status of electronic signatures in many parts of the world, there is a need for easily accessible solutions intended as an alternative to handwritten signatures.
+This may be more necessary than ever since electronic communication is the preferred way of communication.
+This bachelor thesis aims to explore the principles and legal status of electronic signatures, review current electronic signature software, and explore possible obstacles the open-source community is facing to develop such specialized applications.
+We propose an open-source, cross-platform, and user-friendly software compliant with the eIDAS Regulation (Regulation No 910/2014).
+Our application should allow ordinary users to quickly sign and verify signatures of different types of documents and therefore easily use them in everyday life.
+}
+
+\keywords{electronic signature, electronic seal, qualified, open-source, XAdES, PAdES, CAdES, desktop software}
+
+\abstrakt{
+Abstrakt v SK jazyku bude pridaný neskôr ako voľný preklad z EN verzie (pozri ďalšiu stranu). Predíde sa tak zbytočnej práci keďže EN abstrakt sa môže časom meniť.
+}
+
+\klucoveslova{elektronický podpis, elektronická pečať, kvalifikovaný, open-source, XAdES, PAdES, CAdES, počítačový softvér}
+
+\bibliographystyle{alpha}
 
 \begin{document}
+\maketitle
+\newpage
+\tableofcontents
+\listoffigures
+\listoftables
+% zoznam značiek a skratiek. Pre tento koncept neexistuje
+% LaTeXovsky príkaz
 
-\maketitle
+\uvod
+
+Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
+Integer lacinia, nulla porta varius tempus, lacus metus blandit
+lorem, a rutrum justo wisi id sapien. Integer risus libero,
+feugiat eleifend, ornare ac, volutpat nec, sem. In facilisis,
+quam eu elementum aliquet, lorem quam euismod dui, aliquet
+laoreet purus ipsum ac quam.
+
+\chapter{Background and theory}
+
+In our work, we rely on knowledge in several different subjects that we go over in the following sections.
+Firstly, the current state of the law and recent changes relating to the use of electronic signatures.
+Secondly, on the well-established concepts in cryptography like asymmetric cryptography or hashing.
+Lastly, expertise from the field of software engineering - especially design, implementation, and maintenance of computer software.
+All in relation to the electronic signatures, and, where applicable, open-source and desktop computer software.
+
+\section{Law}
+
+Unless specified otherwise, we are considering law applicable locally in the Slovak Republic (SR).
+Its law is greatly influenced by the European Union (being its member since 2004) and to some extent by the rest of the world.
+Therefore, we can assume this is at least partially applicable outside of the SR.
+
+\subsection{Signatures}
+
+Signatures are essential part of the written legally binding documents like contracts.
+They are permanently affixed to the document and are supposed to uniquely identify the person and its deliberate, informed consent.
+As can be seen in the Slovak Civil Code, "A written legal act is valid if signed by the acting person;"\cite{1}.
+Law often explicitly requires signatures and further clarifies their expected use.
+For example, when selling an enterprise, "The contract requires written form and attested signatures of the seller and the buyer."\cite{2}.
+
+Signature is considered an "attested signature"\footnote{Translation from "osvedčený podpis" as used in the Slovak law.} if it is attested by the authorized third party.
+Such process, known as legalization, is regulated by the law and its purpose is to attest information which could form the basis for the exercise of rights or which could cause legal consequences\cite{3}.
+
+Historically, \todo{Comment on law and electronic signatures}
+
+\subsection{European Union Regulation eIDAS}
+
+With intention to stimulate digital growth, the European Union established regulation on electronic identification and trust services for electronic transactions in the European Single Market (eIDAS).
+It applies from 1st of July 2016 and, among other things, it regulates electronic signatures focusing on the:
+
+1. Electronic signatures in general.
+2. Qualified electronic signatures.
+3. Trust service.
+
+\todo{Explain and cite how electronic signatures can be used instead of hand-written one and qualified instead of attested}
+
+\subsection{Copyright and computer software}
+
+Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
+
+\section{Cryptography}
+
+-
+
+\section{Software engineering}
+
+-
+
+\chapter{Methodology}
+
+Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
+Integer lacinia, nulla porta varius tempus, lacus metus blandit
+lorem, a rutrum justo wisi id sapien. Integer risus libero,
+feugiat eleifend, ornare ac, volutpat nec, sem. In facilisis,
+quam eu elementum aliquet, lorem quam euismod dui, aliquet
+laoreet purus ipsum ac quam.
+
+\chapter{Suggested solution}
+
+Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
+Integer lacinia, nulla porta varius tempus, lacus metus blandit
+lorem, a rutrum justo wisi id sapien. Integer risus libero,
+feugiat eleifend, ornare ac, volutpat nec, sem. In facilisis,
+quam eu elementum aliquet, lorem quam euismod dui, aliquet
+laoreet purus ipsum ac quam.
+
+\chapter{Results}
+
+Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
+Integer lacinia, nulla porta varius tempus, lacus metus blandit
+lorem, a rutrum justo wisi id sapien. Integer risus libero,
+feugiat eleifend, ornare ac, volutpat nec, sem. In facilisis,
+quam eu elementum aliquet, lorem quam euismod dui, aliquet
+laoreet purus ipsum ac quam.
+
+Donec dolor arcu, posuere at, vehicula vitae, accumsan ut,
+lacus. Nulla tristique eros eu diam. Vivamus nec tortor vel
+ligula elementum lacinia. Curabitur euismod eros adipiscing
+ipsum. Donec sed quam at felis suscipit egestas. Morbi faucibus
+libero sit amet libero. Nullam laoreet ipsum eu eros. Donec in
+diam. Ut facilisis eros vel leo. Nunc vitae mauris. Donec leo
+erat, luctus porttitor, laoreet eget, facilisis non, erat.
+Integer nec elit.
+
+\zaver
+
+Ut lobortis semper risus, non condimentum dui convallis ut.
+Nulla eget volutpat tellus.
+Vestibulum lobortis tincidunt massa eu rhoncus.
+Suspendisse luctus eu dui non vehicula.
+Vivamus elementum auctor felis, placerat maximus magna lobortis ut.
+Donec placerat sem a mi sagittis blandit.
+Maecenas pellentesque laoreet mauris, dictum viverra ante sodales sed.
+Nullam non ligula quis ante ultricies finibus non quis ex.
+Ut tempor vitae ipsum sed imperdiet.
+Fusce aliquam nisl sit amet nunc tempor convallis.
+Vivamus vehicula magna sit amet purus commodo, et tincidunt purus accumsan.
+Nulla velit dolor, lacinia nec scelerisque a, euismod a sapien.
+%
+
+%\renewcommand{\bibname}{Zoznam použitej literatúry}
+\begin{thebibliography}{9}
+  % Príklady popisu dokumentov citácií podľa systému meno a dátum (Harvardský systém)
+  % ----
+  % Varianty zápisov autorov:
+  % [1] GUZANIN, Štefan, Robert SABOVČÍK a Pavol KAČMÁR. Priezviská vždy VEĽKÝMI PÍSMENAMI,
+  %   priezvisko prvého autora je vždy pred menom, druhý a ďalší autor majú zápis
+  %   Meno PRIEZVISKO
+  % [2] Neuvádzať rodné mená autorov.
+  % [3] Verzálky nie sú povinné, možno použiť aj iné indikatívnejšie označenie
+  %
+  % --- 
+  % 1. Knižná publikácia (monografia, učebnica, zborník ...)
+  %   1 autor
+
+  \bibitem{1}
+  \emph{§ 40 ods. 3 zákona č. 40/1964 Zb. občiansky zákonník}
+
+  \bibitem{2}
+  \emph{§ 476 ods. 2 zákona č. 513/1991 Zb. obchodný zákonník}
+
+  \bibitem{3}
+  \emph{§ 56 ods. 1 zákona č. 323/1992 Zb. notársky poriadok}
+
+  %\bibitem{2}
+  %\osoba{Beck, G.}, 2007. \emph{Zakázaná rétorika: 30 manipulatívních technik}. Preklad
+  %\osoba{Pomikálová, M.}. Praha: Grada Publishing. ISBN 978-80-247-1743-2.
+  %\bibitem{3}
+  %\osoba{Vojčík, P.}, 2010. \emph{Občianske právo hmotné II.} 3. prep. a dopl. vyd. Košice: UPJŠ v Košiciach. ISBN 978-80-7097-817-7.
+  %  2 autori
+  %\bibitem{4}
+  %\osoba{Šoltés, M.} a \osoba{Radoňák, J.}, 2013. \emph{Základné princípy laparoskopickej chirurgie.} Košice: UPJŠ v Košiciach. ISBN 978-80-8152-074-7.
+  % 3 autori
+  %\bibitem{5}
+  %\osoba{Guzanin, Š.}, \osoba{Sabovčík, R.} a \osoba{Kačmár, P.}, 2004.
+\emph{Selected Chapters of Plastic and Reconstructive Surgery: vysokoškolské učebné texty}.
+Košice: Univerzita Pavla Jozefa Šafárika v Košiciach, Lekárska fakulta.
+ISBN 80-7097-557-1.
+  % 4 a viac autorov
+  %\bibitem{6}
+  %\osoba{Nagyová, I.} et al. 2009. \emph{Measuring health and quality of life in the chronically ill}. Košice: Equilibria. ISBN 978-80-892-8446-7.
+  % Elektronická kniha
+  %\bibitem{7}
+  %\osoba{Speight, J.
+G.}, 2005.
+\emph{Lange's Handbook of Chemistry} [online].
+London: McGraw-Hill.
+[cit.
+2009.06.10.] ISBN 978-1-60119-261-5.
+Dostupné na: \url{http://www.knovel.com/web/portal/basic_search/display?_EXT_KNOVEL_DISPLAY_bookid=1347&_EXT_KNOVEL_DISPLAY_fromSearch=true&_EXT_KNOVEL_DISPLAY_searchType=basic}
+  % zborník
+  %\bibitem{8}
+  %\osoba{Bačkor, M.} a \osoba{Mihaličová, S.}, zost., 2013.
+\emph{Zborník príspevkov z konferencie 11.
+dni doktorandov experimentálnej biológie rastlín a 13.
+konferencie experimentálnej biológie rastlín} [online].
+Košice: Univerzita Pavla Jozefa Šafárika v Košiciach, Prírodovedecká fakulta [cit.
+2009-06-10].
+ISBN 9788081520327.
+Dostupné na: \url{
+  %  http://www.upjs.sk/public/media/5596/PF-Zbornik-prispevkov-konferencie-11-dni-doktorandov.pdf}
+  %
+  % 2. Časopis (ako celok)
+  %\bibitem{9}
+  %\emph{Thaiszia: Journal of Botany}. Košice: P.J.Safarik University, Botanic Garden, \mbox{1990--\ .} ISSN 1210-0420.
+  %\bibitem{10}
+  %\emph{Ikaros: elektronický časopis o informační bezpečnosti} [online], 2002. [Praha]: Ikaros. 1997--{} [cit. 2002-03-08]. Dostupné na: \url{http://www.ikaros.cz/}. ISSN 1212-5075.
+  % Jedno číslo časopisu
+  %\bibitem{11}
+  %\emph{CHIP: magazín informačních technologií}, 2013. Praha: Burda Praha, roč. 23, říjen. ISSN 1210-0684.
+  %  ------
+  %  3. Príspevok v knihe/zborníku
+  %  ------
+  %\bibitem{12}
+  %\osoba{Sabol, J.}, 2000. Jazyk ako ľudské posolstvo: (namiesto doslovu). In: \emph{O jazyku a štýle kriticky aj prakticky}. Prešov: Náuka, s. 149--159. ISBN 809676022X.
+  %\bibitem{13}
+  %\osoba{Tóthová, E.} a kol., 2013.
+A rare t(9,22,16)(q34,q11,q24) translocation in chronic myeloid leukemia for which imatinib mesylate was effective: a case report.
+In: \emph{XXVII.
+Olomoucké hematologické dny s mezinárodní účastí, 12.--14.5.2013, Olomouc: sborník abstrakt}.
+Olomouc: Univerzita Palackého v Olomouci, s.
+75--76.
+ISBN 9788024434803.
+  %  ------
+  % 4. Článok v časopise
+  %  ------
+  %\bibitem{14}
+  %\osoba{Beňačka, J.} et al., 2009.
+A better cosine approximate solution to pendulum equation.
+In: \emph{International Journal of Mathematical Education in Science and Technology}.
+Vol.
+40, no.
+2, p.
+206--215.
+ISSN 0020-739X.
+  %\bibitem{15}
+  %\osoba{Dubayová, T.} et al., 2010.
+The impact of the intensity of fear on patient's delay regarding health care seeking behavior: a systematic review vyhľadaní zdravotníckej starostlivosti.
+In: \emph{International Journal of Public Health}.
+Vol.
+55, no.
+5, p.
+459--468.
+ISSN 1661-8556.
+  %\bibitem{16}
+  %\osoba{Steinerová, J.}, 2000. Princípy formovania vzdelania v informačnej vede. In: \emph{Pedagogická revue}. Roč. 2, č. 3, s. 8--16. ISSN 1335-1982.
+  %\bibitem{17}
+  %\osoba{Hoggan, D.}, 2002.
+Challenges, Strategies, and Tools for Research Scientists.
+In: \emph{Electronic Journal of Academic and Special Librarianship} [online].
+Vol.
+3, no.
+3 [cit.
+2013-01-10].
+ISSN 1525-321X.
+Dostupné na: \url{http://southernlibrarianship.icaap.org/content/v03n03/Hoggan_d01.htm}
+  %\bibitem{18}
+  %\osoba{Srbecká, Gabriela}, 2010.
+Rozvoj kompetencí studentů ve vzdělávání.
+In: \emph{Inflow: information journal} [online].
+Roč.
+3, č.
+7 [cit.
+2013-08-06].
+ISSN 1802-9736.
+Dostupné na: \url{http://www.inflow.cz/rozvoj-kompetenci-studentu-ve-vzdelavani}
+  %  ------
+  % 5. Príspevok v zborníku na CD-ROM
+  %  ------
+  %\bibitem{19}
+  %\osoba{Zemánek, P.}, 2001.
+The machines for ``green works'' in vineyards and their economical evaluation.
+In: \emph{9th International Conference: proceedings.
+Vol.
+2.
+Fruit Growing and viticulture} [CD-ROM].
+Lednice: Mendel University of Agriculture and Forestry, p.
+262--268.
+ISBN 80-7157-524-0.
+  %  ------
+  % 6. Záverečné a kvalifikačné práce
+  %  ------
+  %\bibitem{20}
+  %\osoba{Mikulášiková, M.}, 1999. \emph{Didaktické pomôcky pre praktickú výučbu na hodinách výtvarnej výchovy pre 2. stupeň základných škôl}: diplomová práca. Nitra: UKF.
+  %\bibitem{21}
+  %\osoba{Urdzík, P.}, 2007. \emph{Predikcia intrauterinnej rastovej retardácie a preeklampsie pomocou biochemických a ultrazvukových markerov}: dizertačná práca. Košice: UPJŠ v Košiciach.
+  %  ------
+  % 7. Výskumné správy
+  %  ------
+  %\bibitem{22}
+  %\osoba{Baumgartner, J.} a kol., 1998. \emph{Ochrana a udržiavanie genofondu zvierat, šľachtenie zvierat}: výskumná správa. Nitra: VÚŽV.
+  %  ------
+  % 8. Normy
+  %  ------
+  %\bibitem{23}
+  %STN ISO 690: 2012. \emph{Informácie a dokumentácia. Návod na tvorbu bibliografických odkazov na informačné pramene a ich citovanie}.
+  %  ------
+  % 9. Mapa
+  %  ------
+  %\bibitem{24}
+  %VKÚ, 2003. \emph{Košice: mapa okolia}. [1:15000]. 3. vyd. Harmanec: VKÚ. ISBN 80-8042-223-0.
+  %  ------
+  % 10. Zákon
+  %  ------
+  %\bibitem{25}
+  %\emph{Zákon č.131/2002 Zb. o vysokých školách a o zmene a doplnení niektorých zákonov.}
+  %\bibitem{26}
+  %\emph{Zákon č. 313/2001 o verejnej službe.}
+
+\end{thebibliography}
 

### 2020-01-27 paper/main.tex
Complete Law section in Background & Theory

@@
 
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
+Will be added at the end. Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
 Integer lacinia, nulla porta varius tempus, lacus metus blandit
@@
 Unless specified otherwise, we are considering law applicable locally in the Slovak Republic (SR).
-Its law is greatly influenced by the European Union (being its member since 2004) and to some extent by the rest of the world.
+Its law is greatly influenced by the European Union (EU), being its member since 2004, and to some extent by the rest of the world.
 Therefore, we can assume this is at least partially applicable outside of the SR.
@@
 They are permanently affixed to the document and are supposed to uniquely identify the person and its deliberate, informed consent.
-As can be seen in the Slovak Civil Code, "A written legal act is valid if signed by the acting person;"\cite{1}.
+As can be seen in the Slovak Civil Code, "A written legal act is valid if signed by the acting person;" \cite{1}.
 Law often explicitly requires signatures and further clarifies their expected use.
-For example, when selling an enterprise, "The contract requires written form and attested signatures of the seller and the buyer."\cite{2}.
+For example, when selling an enterprise, "The contract requires written form and attested signatures of the seller and the buyer." \cite{2}.
 
 Signature is considered an "attested signature"\footnote{Translation from "osvedčený podpis" as used in the Slovak law.} if it is attested by the authorized third party.
-Such process, known as legalization, is regulated by the law and its purpose is to attest information which could form the basis for the exercise of rights or which could cause legal consequences\cite{3}.
+Such process, known as legalization, is regulated by the law and its purpose is to attest information which could form the basis for the exercise of rights or which could cause legal consequences \cite{3}.
 
-Historically, \todo{Comment on law and electronic signatures}
+With regards to electronic signatures, law within the EU used to differ, with law applicable in the SR being now repealed "Zákon č. 215/2002 Z.z.".
 
-\subsection{European Union Regulation eIDAS}
+\subsection{EU Regulation eIDAS}
 
-With intention to stimulate digital growth, the European Union established regulation on electronic identification and trust services for electronic transactions in the European Single Market (eIDAS).
-It applies from 1st of July 2016 and, among other things, it regulates electronic signatures focusing on the:
+With intention to stimulate digital growth by building trust, the EU established regulation on electronic identification and trust services for electronic transactions in the internal market (eIDAS).
+It applies from 1st of July 2016, replaces local law, and regulates, among other things, electronic signatures and its more specific variants.
 
-1. Electronic signatures in general.
-2. Qualified electronic signatures.
-3. Trust service.
+\obrazok{./figures/electronic-advanced-qualified}{The relation between electronic, advanced, and qualified signatures.}
 
-\todo{Explain and cite how electronic signatures can be used instead of hand-written one and qualified instead of attested}
+In general, electronic signature can be represented in different ways (e.g.
+as an image, text or other data attached to the document) and they can not be denied in legal proceedings just because they are in the electronic form \cite{4}.
 
-\subsection{Copyright and computer software}
+Advanced electronic signatures are a subset of electronic signatures that have to uniquely link and identify the signature author, be created in a way that is possible only by them and any changes to the signed document is detectable \cite{5}.
+From the technical point of view, standards PAdES, XAdES and, CAdES specified by the ETSI comply with these requirements.
 
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
+Qualified electronic signatures (QES) are a subset of advanced electronic signatures with more specific requirements that can be found in the annex 1 of the eIDAS regulation\footnote{https://eur-lex.europa.eu/legal-content/EN/TXT/HTML/?uri=CELEX:32014R0910\#d1e32-111-1}.
+In more practical terms, such advanced electronic signatures are created with a qualified device using a qualified certificate and a qualified trust service.
+Qualified in this case means authorized for such use by the legal authorities.
+
+Furthemore, a qualified electronic signature has the legal effect of a handwritten signature \cite{4} and attested handwritten signatures in the SR.
+
+Mentioned terminology is used when considering natural person. Legal entities are able to use "electronic seals", which are, from the technical point of view and our purposes, identical.
+
+Overall, electronic signatures are supposed to be, within the EU, interoperable and transparent alternative to hand-written signatures.
+
+\subsection{Software copyright}
+
+Computer programs and associated materials are historically protected under copyright as literary works, and protected authors should be able to authorize or prohibit certain acts \cite{6}.
+
+Such right can be exercised, as far as we are concerned, to license the software under either proprietary or open-source license.
+Proprietary meaning under exclusive legal right of the author, typically also confidential and distributed as a paid product.
+Open-source meaning having its source code freely available, typically also distributed for free.
+
+Motivation to license the software under an open-source license can differ, and so do such licenses.
+While a proprietary license is usually made specifically for that entity and its interests, open-source licenses tend to be reused between different authors.
+This means consumers of the software can quickly recognize their rights and responsibilities if they decide to use, modify, or distribute the software.
+In general, we can classify the open-source licenses into two categories: permissive and copyleft.
+
+Copyleft license, in general, requires the user to publish the modified work under the compatible (free) license.
+Meaning forcing them to extend the rights they have received onto others and, in turn, somehow hindering use in software licensed under a different license and potential commercialization.
+Popular example of such license is the GNU General Public License (GPL) and its derivatives like GNU Lesser GPL (LGPL) or GNU Affero GPL (AGPL).
+
+Permissive license, on the other hand, does not have such requirement in place and is more suitable for potential commercialization.
+They typically allow commercial use, modifications, distribution or sublicensing as long as the author is not held liable and the original copyright and license is distributed with the software.
+Often used permissive licenses are MIT License, Apache License or BSD License.
+
+It is not uncommon to see hybrid licensing - licensing under more than one license with one being open-source and one proprietary.
+
+\obrazok{./figures/license-comparison}{Classification and comparison of software licensing models.}
 
@@
 
--
+Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
+Integer lacinia, nulla porta varius tempus, lacus metus blandit
+lorem, a rutrum justo wisi id sapien. Integer risus libero,
+feugiat eleifend, ornare ac, volutpat nec, sem. In facilisis,
+quam eu elementum aliquet, lorem quam euismod dui, aliquet
+laoreet purus ipsum ac quam.
+
+\subsection{Asymmetric cryptography}
+
+Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
+Integer lacinia, nulla porta varius tempus, lacus metus blandit
+lorem, a rutrum justo wisi id sapien. Integer risus libero,
+feugiat eleifend, ornare ac, volutpat nec, sem. In facilisis,
+quam eu elementum aliquet, lorem quam euismod dui, aliquet
+laoreet purus ipsum ac quam.
+
+\subsection{Hashing}
+
+Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
+Integer lacinia, nulla porta varius tempus, lacus metus blandit
+lorem, a rutrum justo wisi id sapien. Integer risus libero,
+feugiat eleifend, ornare ac, volutpat nec, sem. In facilisis,
+quam eu elementum aliquet, lorem quam euismod dui, aliquet
+laoreet purus ipsum ac quam.
 
@@
 
--
+Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
+Integer lacinia, nulla porta varius tempus, lacus metus blandit
+lorem, a rutrum justo wisi id sapien. Integer risus libero,
+feugiat eleifend, ornare ac, volutpat nec, sem. In facilisis,
+quam eu elementum aliquet, lorem quam euismod dui, aliquet
+laoreet purus ipsum ac quam.
 
-\chapter{Methodology}
+\subsection{Design}
+
+Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
+Integer lacinia, nulla porta varius tempus, lacus metus blandit
+lorem, a rutrum justo wisi id sapien. Integer risus libero,
+feugiat eleifend, ornare ac, volutpat nec, sem. In facilisis,
+quam eu elementum aliquet, lorem quam euismod dui, aliquet
+laoreet purus ipsum ac quam.
+
+\subsection{Implementation}
+
+Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
+Integer lacinia, nulla porta varius tempus, lacus metus blandit
+lorem, a rutrum justo wisi id sapien. Integer risus libero,
+feugiat eleifend, ornare ac, volutpat nec, sem. In facilisis,
+quam eu elementum aliquet, lorem quam euismod dui, aliquet
+laoreet purus ipsum ac quam.
+
+\subsection{Maintenance}
+
+Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
+Integer lacinia, nulla porta varius tempus, lacus metus blandit
+lorem, a rutrum justo wisi id sapien. Integer risus libero,
+feugiat eleifend, ornare ac, volutpat nec, sem. In facilisis,
+quam eu elementum aliquet, lorem quam euismod dui, aliquet
+laoreet purus ipsum ac quam.
+
+\chapter{Goals and analysis}
 
@@
   \emph{§ 56 ods. 1 zákona č. 323/1992 Zb. notársky poriadok}
+
+  \bibitem{4}
+  \emph{Article 25, Regulation (EU) No 910/2014}
+
+  \bibitem{5}
+  \emph{Article 26, Regulation (EU) No 910/2014}
+
+  \bibitem{6}
+  \emph{Council Directive 91/250/EEC of 14 May 1991 on the legal protection of computer programs}
 
@@
   %  ------
-  % 9. Mapa
-  %  ------
-  %\bibitem{24}
-  %VKÚ, 2003. \emph{Košice: mapa okolia}. [1:15000]. 3. vyd. Harmanec: VKÚ. ISBN 80-8042-223-0.
   %  ------

### 2020-01-27 paper/main.tex
Update Law subsection, Cryptography outline

@@
 Advanced electronic signatures are a subset of electronic signatures that have to uniquely link and identify the signature author, be created in a way that is possible only by them and any changes to the signed document is detectable \cite{5}.
-From the technical point of view, standards PAdES, XAdES and, CAdES specified by the ETSI comply with these requirements.
+From the technical point of view, standards PAdES, XAdES, and CAdES specified by the ETSI comply with these requirements.
 
@@
 Proprietary meaning under exclusive legal right of the author, typically also confidential and distributed as a paid product.
-Open-source meaning having its source code freely available, typically also distributed for free.
+Open-source meaning having its source code freely available, typically also distributed for free and without any liability (the software is provided "as is").
 
@@
 Copyleft license, in general, requires the user to publish the modified work under the compatible (free) license.
-Meaning forcing them to extend the rights they have received onto others and, in turn, somehow hindering use in software licensed under a different license and potential commercialization.
+Meaning forcing them to extend the rights they have received onto others and, in turn, somehow hindering use in software licensed under a different license.
 Popular example of such license is the GNU General Public License (GPL) and its derivatives like GNU Lesser GPL (LGPL) or GNU Affero GPL (AGPL).
@@
 
-It is not uncommon to see hybrid licensing - licensing under more than one license with one being open-source and one proprietary.
+It is not uncommon to see hybrid licensing - licensing under more than one license, with one being typically open-source and one proprietary.
+In this scenario, users can choose which license they want to use based on their needs.
 
@@
 
+As we have already mentioned, software licenses can influence whether and how it can be used as a part of different software.
+Both when we are using someone else's software and when someone else is using our software.
+Generally speaking, the more open the license is, the higher the probability it will not stop adoption for legal reasons.
+Various possible combinations can be seen in Figure 3, where greyed out options mean such combination is not possible unless a separate licensing agreement is reached with the copyright owner.
+
+\obrazok{./figures/license-compatibility}{Schematic representation of license directionality.\cite{7}}
+
+\newpage
+
 \section{Cryptography}
 
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-Integer lacinia, nulla porta varius tempus, lacus metus blandit
-lorem, a rutrum justo wisi id sapien. Integer risus libero,
-feugiat eleifend, ornare ac, volutpat nec, sem. In facilisis,
-quam eu elementum aliquet, lorem quam euismod dui, aliquet
-laoreet purus ipsum ac quam.
+Previously mentioned standards PAdES, XAdES, and CAdES define how asymmetric cryptography, hashing, or timestamping should be used to comply with the standards.
+We go over a few of these topics and why they are able to satisfy the requirements.
 
-\subsection{Asymmetric cryptography}
+\subsection{Asymmetric encryption}
 
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-Integer lacinia, nulla porta varius tempus, lacus metus blandit
-lorem, a rutrum justo wisi id sapien. Integer risus libero,
-feugiat eleifend, ornare ac, volutpat nec, sem. In facilisis,
-quam eu elementum aliquet, lorem quam euismod dui, aliquet
-laoreet purus ipsum ac quam.
+TODO: How it works.
+
+TODO: How the private key is stored on the HSM, PKCS11 (be created in a way that is possible only by them).
+
+TODO: Third party completing the trust chain (uniquely link and identify).
 
@@
 
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-Integer lacinia, nulla porta varius tempus, lacus metus blandit
-lorem, a rutrum justo wisi id sapien. Integer risus libero,
-feugiat eleifend, ornare ac, volutpat nec, sem. In facilisis,
-quam eu elementum aliquet, lorem quam euismod dui, aliquet
-laoreet purus ipsum ac quam.
+TODO: How it works.
+
+TODO: How it is used on the documents (guaranted that changes are detected).
+
+\subsection{Timestamping}
+
+TODO: How it works.
+
+TODO: How it is used on the documents (parallels with the hand-written signature and date on the contract).
+
+TODO: Long term validation.
+
+\subsection{PAdES, XAdES, and CAdES}
+
+TODO: Quick explanation of each.
+
+TODO: How they can be used on a different file formats (table).
 
@@
 
+  %  ------
+  % Law
+  %  ------
   \bibitem{1}
@@
   \emph{Council Directive 91/250/EEC of 14 May 1991 on the legal protection of computer programs}
+
+  %  ------
+  % Article
+  %  ------
+  \bibitem{7}
+  \osoba{Morin, A.}, et al.
+2012.
+A Quick Guide to Software Licensing for the Scientist-Programmer.
+In: \emph{PLoS computational biology} [online].
+Volume 8, issue 7 [cit.
+2020-01-27].
+e1002598.
+Available at: \url{https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1002598}
 

### 2020-01-28 paper/main.tex
Edit bibliography
Update footnote link

@@
 
-\title{Open electronic signature software}
+\title{Replacing hand-written signatures with open electronic signature software}
 \author{Jakub Ďuraš}
@@
 
-Qualified electronic signatures (QES) are a subset of advanced electronic signatures with more specific requirements that can be found in the annex 1 of the eIDAS regulation\footnote{https://eur-lex.europa.eu/legal-content/EN/TXT/HTML/?uri=CELEX:32014R0910\#d1e32-111-1}.
+Qualified electronic signatures (QES) are a subset of advanced electronic signatures with more specific requirements that can be found in the annex 1 of the eIDAS regulation\footnote{Available at https://eur-lex.europa.eu/legal-content/EN/TXT/HTML/?uri=CELEX:32014R0910\#d1e32-111-1}.
 In more practical terms, such advanced electronic signatures are created with a qualified device using a qualified certificate and a qualified trust service.

### 2020-01-29 articles/legal basis.tex
Add legal basis and motivation articles

\documentclass[a4paper,twocolumn]{article}
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{graphicx}

\renewcommand{\baselinestretch}{1.3}

\title{Legal basis for replacing handwritten signatures with open electronic signature software}
\author{Jakub Ďuraš, P. J. Safarik University}

\begin{document}

{\setstretch{1.0}
    \maketitle
}

\section*{Introduction}

Signing of documents is something we get used as we get older.
It is usually expected we will sign a real paper document.
How easy it is if we want to sign some documents without meeting with another party in person?
Do we print the documents from their electronic form only because we need to sign them? There is an alternative in the form of electronic signatures.
On what legal ground can we do so and what if we also want to make it a part of open-source movement?

Unless specified otherwise, we are considering law applicable locally in the Slovak Republic (SR).
Its law is greatly influenced by the European Union (EU), being its member since 2004, and to some extent by the rest of the world.
Therefore, we can assume this is at least partially applicable outside of the SR.

\section{Signatures}

Essential part of the written legally binding documents, they are permanently affixed to the document and are supposed to uniquely identify the person and its deliberate, informed consent.
As can be seen in the Slovak Civil Code, "A written legal act is valid if signed by the acting person;" \cite{1}.
Law often explicitly requires signatures and further clarifies their expected use.
For example, when selling an enterprise, the contract is required in written form and attested signatures of the seller and the buyer \cite{2}.

Signature is considered an "attested signature"\footnote{Translation from "osvedčený podpis" as used in the Slovak law.} if it is verified by the authorized third party.
Such process, known as legalization, is regulated by the law and its purpose is to form the basis for the exercise of legal rights \cite{3}.

With regards to electronic signatures, law within the EU used to differ, with one applicable in the SR being now repealed "Zákon č. 215/2002 Z.z.".

\section{EU Regulation eIDAS}

With the intent to stimulate digital growth by building trust, the EU established regulation on electronic identification and trust services for electronic transactions in the internal market (eIDAS).
It applies from 1st of July 2016, caused the local law to be repealed, and regulates, among other things, electronic signatures and its more advanced forms.

\begin{figure}[ht]
    \includegraphics[width=\linewidth]{../paper/figures/electronic-advanced-qualified}
    \caption{The relation between electronic, advanced, and qualified signatures.}
    \centering
\end{figure}

In general, electronic signature can be represented in different ways (e.g. as an image, text or other data attached to the document) and it can not be denied in legal proceedings just because it is in the electronic form \cite{4}.

Advanced electronic signatures are a subset of electronic signatures that have to uniquely link and identify the signature author, be created in a way that is possible only by them, and any changes to the signed document have to be detectable \cite{5}.
From the technical point of view, standards PAdES, XAdES, and CAdES specified by the ETSI comply with these requirements.

Qualified electronic signatures (QES) are a subset of advanced electronic signatures with more specific requirements that can be found in the Annex 1 of the eIDAS regulation\footnote{Available at https://eur-lex.europa.eu/legal-content/EN/TXT/HTML/?uri=CELEX:32014R0910\#d1e32-111-1}.
In more practical terms, such advanced electronic signatures are created with a qualified device, certificate, and using a qualified trust service. Qualified in this case means authorized for such use by the legal authorities that maintain publicly accessible, transparent list.

The qualified electronic signature has the legal effect of the handwritten signature \cite{4} or attested handwritten signature in the SR specifically.

Such terminology is used only when considering natural person. Legal entities are able to use "electronic seals", which are, from the technical point of view and for our purposes, almost identical.

\section{Software copyright}

Computer programs and associated materials shoule be protected under copyright as literary works, and protected authors should be able to authorize or prohibit certain acts \cite{6}.

Such right can be exercised, as far as we are concerned, to license the software under either proprietary or open-source license.
Proprietary meaning under the exclusive legal right of the author with source code being typically also confidential and distributed as a paid product.
Open-source meaning having its source code freely available, usually also distributed for free, and without any liability (the software is provided "as is").

Motivations to license the software under an open-source license can vary, and so do such licenses.
While a proprietary license is usually made specifically for that entity and its interests, open-source licenses tend to be reused between different authors.
This means consumers of the software can quickly recognize their rights and responsibilities if they decide to use, modify, or distribute the software.
In general, we can classify the open-source licenses into two categories: permissive and copyleft.

Copyleft license, in general, requires the user to publish the modified work under the compatible (free) license.
By doing that, it forces users to extend the rights they have received onto others and, in turn, somehow limit potential use in software licensed under a different license.
Popular example of license like that is the GNU General Public License (GPL) and its derivatives like GNU Lesser GPL (LGPL) or GNU Affero GPL (AGPL).

\begin{figure*}[ht!]
    \includegraphics[width=\linewidth]{../paper/figures/license-comparison}
    \caption{Classification and comparison of software licensing models.}
    \centering
\end{figure*}

Permissive license, on the other hand, does not have such requirement in place and is more suitable for potential commercialization.
They typically allow commercial use, modifications, distribution, or sublicensing as long as the author is not held liable, and the original copyright and license are distributed with the software.
Often used permissive licenses are MIT License, Apache License, or BSD License.

It is not uncommon to see hybrid licensing - licensing under more than one license, with one being typically open-source and one proprietary.
In this scenario, users can choose which license they want to use based on their needs.

As we have already mentioned, software licenses can influence whether and how it can be used as a part of different software.
Both when we are using someone else's software and when someone else is using our software.
Generally speaking, the more restrictive the license is, the higher the probability it will prevent adoption for legal reasons.

\begin{figure}[ht]
    \includegraphics[width=\linewidth]{../paper/figures/license-compatibility}
    \caption{Schematic representation of license directionality \cite{7}.}
    \centering
\end{figure}

Various possible combinations can be seen in Figure 3, where greyed out options mean such combination is not possible unless a separate licensing agreement is reached with the copyright owner.

\section*{Conclusion}

Overall, electronic signatures are supposed to be, within the EU, interoperable and transparent alternative to hand-written signatures.

Utilizing open-source software is possible, provided we are mindful of the license.
Likewise, choosing the proper license can help with adoption.

\onecolumn{
    \begin{thebibliography}{9}

    \bibitem{1}
    \emph{Zákon č. 40/1964 Zb. občiansky zákonník, § 40 ods. 3}

    \bibitem{2}
    \emph{Zákon č. 513/1991 Zb. obchodný zákonník, § 476 ods. 2}

    \bibitem{3}
    \emph{Zákon č. 323/1992 Zb. notársky poriadok, § 56 ods. 1}

    \bibitem{4}
    \emph{Regulation (EU) No 910/2014, Article 25}

    \bibitem{5}
    \emph{Regulation (EU) No 910/2014, Article 26}

    \bibitem{6}
    \emph{Directive 2009/24/EC of the European Parliament and of the Council of 23 April 2009 on the legal protection of computer programs}

    %  ------
    % Article
    %  ------
    \bibitem{7}
    MORIN, A., et al. 2012. A Quick Guide to Software Licensing for the Scientist-Programmer. In: \emph{PLoS computational biology} [online]. Volume 8, issue 7 [cit. 2020-01-27]. Available at: https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1002598

    \end{thebibliography}
}

\end{document}

### 2020-01-29 articles/motivation.tex
Add legal basis and motivation articles

\documentclass{article}

\begin{document}

\title{Motivation for choosing a reseach topic}
\author{Jakub Ďuraš}

\maketitle

We are signing documents to verify we are the ones who read the document and agree with its content. It is expected we will sign a real paper document. How easy is it if we want to sign some documents without meeting with another party in person? Do we print the documents from their electronic form only because we need to sign them? There is an alternative in the form of electronic signatures, and there are several reasons we have chosen to concentrate on them.

Firstly, recent changes in the legal status of electronic signatures define their legal status in many parts of the world. For example, in the EU, [eIDAS regulation](https://eur-lex.europa.eu/legal-content/EN/TXT/?uri=uriserv%3AOJ.L_.2014.257.01.0073.01.ENG) establishes the principle that an "electronic signature should not be denied legal effect on the grounds that it is in an electronic form" and a "qualified electronic signature should have the equivalent legal effect of a handwritten signature".

Furthermore, we could not find any specialized, open-source and cross-platform software that would allow for easy document signing using electronic, and, especially, qualified electronic (is defined in the [eIDAS regulation](https://eur-lex.europa.eu/legal-content/EN/TXT/?uri=uriserv%3AOJ.L_.2014.257.01.0073.01.ENG)) signature.

Lastly, current software development frameworks allow efficient cross-platform desktop application development enabling us to develop desktop software for all platforms faster than ever. There are libraries available that can simplify the implementation of the signing. Example of that is the "[DSS: Digital signature service](https://ec.europa.eu/cefdigital/wiki/display/CEFDIGITAL/Services+eSignature)" offered by the [CEF Digital](https://ec.europa.eu/cefdigital/wiki/display/CEFDIGITAL) that should facilitate "creation, extension and validation of advanced electronic signatures." This library is licensed under an open source license that allows usage in our software.

This bachelor thesis, therefore, aims to explore the principles and legal status of electronic signatures, review current electronic signature software, and explore possible obstacles the open-source community is facing to develop such specialized applications. We propose an open-source, cross-platform, and user-friendly software compliant with the eIDAS Regulation (Regulation No 910/2014). Our application should allow ordinary users to quickly sign and verify signatures of different types of documents and therefore easily use them in everyday life.

\end{document}

### 2020-01-29 articles/legal basis.tex
Update legal basis article

@@
 
-Signing of documents is something we get used as we get older.
-It is usually expected we will sign a real paper document.
+We are usually used to signing of paper documents.
 How easy it is if we want to sign some documents without meeting with another party in person?
 Do we print the documents from their electronic form only because we need to sign them? There is an alternative in the form of electronic signatures.
-On what legal ground can we do so and what if we also want to make it a part of open-source movement?
+We would like to explore on what legal ground we can do so and what if we also want to make it a part of the open-source movement.
 
 Unless specified otherwise, we are considering law applicable locally in the Slovak Republic (SR).
-Its law is greatly influenced by the European Union (EU), being its member since 2004, and to some extent by the rest of the world.
-Therefore, we can assume this is at least partially applicable outside of the SR.
+Its law is greatly influenced by the European Union (EU), being its member since 2004, and therefore, we can assume this is at least partially applicable outside of the SR.
 
@@
 
-Essential part of the written legally binding documents, they are permanently affixed to the document and are supposed to uniquely identify the person and its deliberate, informed consent.
+Essential part of the written legally binding documents, signatures are permanently affixed to the document and are supposed to uniquely identify the person and its deliberate, informed consent.
 As can be seen in the Slovak Civil Code, "A written legal act is valid if signed by the acting person;" \cite{1}.
 Law often explicitly requires signatures and further clarifies their expected use.
-For example, when selling an enterprise, the contract is required in written form and attested signatures of the seller and the buyer \cite{2}.
+For example, when selling an enterprise, the contract is required in written form, with attested signatures of the seller and the buyer \cite{2}.
 

### 2020-04-21 paper/main.tex
Add cryptography and outline TODOs to paper

--- 2020-01-27 paper/main.tex
@@
 
-\title{Open electronic signature software}
+\usepackage{tikz}
+\usetikzlibrary{datavisualization}
+\usetikzlibrary{datavisualization.formats.functions}
+
+\title{Replacing handwritten signatures with open electronic signature software}
 \author{Jakub Ďuraš}
@@
 In our work, we rely on knowledge in several different subjects that we go over in the following sections.
+TODO: Sounds very meh for the very first sentece of the thesis.
 Firstly, the current state of the law and recent changes relating to the use of electronic signatures.
@@
 All in relation to the electronic signatures, and, where applicable, open-source and desktop computer software.
+
+TODO: Think about adding section on handwritten signatures
 
@@
 They are permanently affixed to the document and are supposed to uniquely identify the person and its deliberate, informed consent.
-As can be seen in the Slovak Civil Code, "A written legal act is valid if signed by the acting person;" \cite{1}.
+As can be seen in the Slovak Civil Code, "A written legal act is valid if signed by the acting person;" \cite{civilcode}.
 Law often explicitly requires signatures and further clarifies their expected use.
-For example, when selling an enterprise, "The contract requires written form and attested signatures of the seller and the buyer." \cite{2}.
+For example, when selling an enterprise, "The contract requires written form and attested signatures of the seller and the buyer." \cite{civilcode}.
 
 Signature is considered an "attested signature"\footnote{Translation from "osvedčený podpis" as used in the Slovak law.} if it is attested by the authorized third party.
-Such process, known as legalization, is regulated by the law and its purpose is to attest information which could form the basis for the exercise of rights or which could cause legal consequences \cite{3}.
+Such process, known as legalization, is regulated by the law and its purpose is to attest information which could form the basis for the exercise of rights or which could cause legal consequences \cite{notarylaw}.
 
-With regards to electronic signatures, law within the EU used to differ, with law applicable in the SR being now repealed "Zákon č. 215/2002 Z.z.".
+With regards to electronic signatures, law within the EU used to differ, with law applicable in the SR being now repealed "Act No. 215/2002 Coll.".
 
-\subsection{EU Regulation eIDAS}
+\subsection{EU Regulation eIDAS} \label{eidas}
 
@@
 
-\obrazok{./figures/electronic-advanced-qualified}{The relation between electronic, advanced, and qualified signatures.}
+\obrazok{./figures/electronic-advanced-qualified}{The relation between electronic, advanced, and qualified signatures.}{signature-comparison}
 
 In general, electronic signature can be represented in different ways (e.g.
-as an image, text or other data attached to the document) and they can not be denied in legal proceedings just because they are in the electronic form \cite{4}.
+as an image, text or other data attached to the document) and they \textbf{can not be denied in legal proceedings} just because they are in the electronic form \cite{eidas}.
 
-Advanced electronic signatures are a subset of electronic signatures that have to uniquely link and identify the signature author, be created in a way that is possible only by them and any changes to the signed document is detectable \cite{5}.
-From the technical point of view, standards PAdES, XAdES, and CAdES specified by the ETSI comply with these requirements.
+Advanced electronic signatures are a subset of electronic signatures that have to \textbf{uniquely link and identify} the signature author, be created in a way that is \textbf{possible only by them} and any \textbf{changes to the signed document have to be detectable} \cite{eidas}.
+From the technical point of view, standards PAdES, XAdES, and CAdES specified by the European Telecommunications Standards Institute (ETSI) comply with these requirements.
+We explain why this is the case in Chapter \ref{cryptography}.
 
-Qualified electronic signatures (QES) are a subset of advanced electronic signatures with more specific requirements that can be found in the annex 1 of the eIDAS regulation\footnote{https://eur-lex.europa.eu/legal-content/EN/TXT/HTML/?uri=CELEX:32014R0910\#d1e32-111-1}.
+Qualified electronic signatures (QES) are a subset of advanced electronic signatures with more specific requirements that can be found in the annex 1 of the eIDAS regulation\footnote{Available at https://eur-lex.europa.eu/legal-content/EN/TXT/HTML/?uri=CELEX:32014R0910\#d1e32-111-1.}.
 In more practical terms, such advanced electronic signatures are created with a qualified device using a qualified certificate and a qualified trust service.
@@
 
-Furthemore, a qualified electronic signature has the legal effect of a handwritten signature \cite{4} and attested handwritten signatures in the SR.
+Furthemore, a qualified electronic signature has the legal effect of a handwritten signature \cite{eidas} and attested handwritten signatures in the SR.
 
@@
 
-Computer programs and associated materials are historically protected under copyright as literary works, and protected authors should be able to authorize or prohibit certain acts \cite{6}.
+Computer programs and associated materials are historically protected under copyright as literary works, and protected authors should be able to authorize or prohibit certain acts \cite{eeccopyright}.
 
@@
 
-\obrazok{./figures/license-comparison}{Classification and comparison of software licensing models.}
+\obrazok{./figures/license-comparison}{Classification and comparison of software licensing models.}{license-comparison}
 
@@
 Generally speaking, the more open the license is, the higher the probability it will not stop adoption for legal reasons.
-Various possible combinations can be seen in Figure 3, where greyed out options mean such combination is not possible unless a separate licensing agreement is reached with the copyright owner.
+Various possible combinations can be seen in Figure \ref{fig:license-compatibility}, where greyed out options mean such combination is not possible unless a separate licensing agreement is reached with the copyright owner.
 
-\obrazok{./figures/license-compatibility}{Schematic representation of license directionality.\cite{7}}
+\obrazok{./figures/license-compatibility}{Schematic representation of license directionality.\cite{licensing}}{license-compatibility}
 
-\newpage
-
-\section{Cryptography}
+\section{Cryptography} \label{cryptography}
 
@@
 
-\subsection{Asymmetric encryption}
+\subsection{Asymmetric cryptography}
 
-TODO: How it works.
+There are several issues that make symmetric encryption unsuitable.
+Let us imagine we are creating a key for each pair of users in our network - imaginary country Isle of Alice and Bob - for symmetric encryption and decryption of documents shared between each Alice and Bob.
 
-TODO: How the private key is stored on the HSM, PKCS11 (be created in a way that is possible only by them).
+\textbf{Secure distribution of keys} seems to be feasible when we are adding a new user to our network.
+We require each user to personally come and pick up physical device that contains pregenerated keys.
+Changes made after that, however, would mean we either need to require periodic personal visit to update the saved keys or transmission of these keys over some medium which can be potentialy unsafe.
+Any new members of the network are unable to participate untill other Alices and Bobs get their updates.
 
-TODO: Third party completing the trust chain (uniquely link and identify).
+Even if we think we have established a reasonably secure way to do that and we are willing to wait, we realize the \textbf{number of keys} is quickly increasing with each new member.
+With \textit{n} users, the number of keys is
+
+$$\frac{n \cdot (n - 1)}{2}$$
+
+which means that even our little country with a population of only 70 000 requires roughly 2.5 billion keys.
+However, since our signature system needs to be portable, we will need to count also with a population of other EU countries, and the number of keys quickly explodes.
+
+\begin{figure}[!htb]
+  \centering
+
+  \begin{tikzpicture}
+    \datavisualization [scientific axes=clean, visualize as smooth line, all axes={ticks={major={at={}}}}, y axis={label={number of keys}}, x axis={label={number of users}}]
+    data [format=function] {
+      var x : interval [2:125];
+      func y = \value x*\value x;
+    };
+  \end{tikzpicture}
+
+  \centerline{\parbox{13cm}{\sl \caption{Cubic growth of the number of keys.}}}
+\end{figure}
+
+\textbf{Either party can also lie} about not signing the document.
+Since both have the key that can be used for encryption, they can sign any document on behalf of the other party, and there is no way to disprove it if we look at it as a third party.
+
+These issues can be addressed with asymmetric cryptography.
+In general, we assign a pair of keys to each person, one of them used for encryption (public key), another used for decryption (private key).
+One party can encrypt, and the other is the only one that can decrypt.
+
+Digital signatures are a specific public-key algorithms.
+The signer is the only one who poses private key used for signing, and others can use the public key for verifying of signatures as can be seen in Figure \ref{fig:pubkey-scheme} where \(x\) is the "document", and \(s\) is the signature that is sent with the document.
+Alice can then sign the document \textit{including} Bob's signature.
+
+\obrazok{./figures/digital-signature-scheme}{Principle of digital signatures which involves signing and verifying a message (Paar 2009). TODO: I don't have rights to this image!!!}{pubkey-scheme}
+
+Practical, widely used implementation of this idea is the RSA signature scheme that relies on the integer factorization problem (difficulty of factoring a product of two large primes) published in 1978 \cite{rsa}.
+
+In Slovakia, the private key is usually handed over to the user personally, saved on the ID card.
+This ID card (used together with a card reader) provides an API for safe access, management, and cryptoprocessing.
+Such device is in general called a Hardware Security Module (HSM), and the API relevant for this thesis is PKCS \#11 \footnote{PKCS \#11 specification available at \url{http://docs.oasis-open.org/pkcs11/pkcs11-base/v2.40/os/pkcs11-base-v2.40-os.html}.}.
+
+We are left with the problem of public key distribution - even though public-key schemes do not require a \textit{secure channel}, they require \textit{authenticated channels} for the distribution of the public keys \cite{cryptotxtbook}.
+
+Common solution is the use of \textit{certificates}. Certificates are essentially digital signatures with metadata that can be used to establish validity of the received public key before it is used.
+Distribution of such certificates is the responsibility of a Certification Authority (CA) - a third party that all users in the network trust.
+In the case of the Slovak ID cards, it is Disig - SVK eID ACA issued by the National Security Authority (NBÚ in Slovak).
+Such certificates therefore reliably uniquely link and identify the public key owners as can be seen in Figure \ref{fig:certificate-scheme}\footnote{Created by unknown author, licensed under the CC BY-SA 3.0 license.}.
+
+\obrazok{./figures/certificate-scheme}{Application of digital signature with certificate. TODO: Modify to add CA, remove hashing}{certificate-scheme}
+
+Commonly used standard for public key certificates is X.509\footnote{Specification available under RFC5280 at \url{https://tools.ietf.org/html/rfc5280}.}.
 
@@
 
-TODO: How it works.
+It is often useful to map data of arbitrary size to fixed-length (typically short) set of bits.
+This process is commonly called \textit{hashing}, function that can be used for it \textit{hash function}, and output of such function \textit{hash}.
 
-TODO: How it is used on the documents (guaranted that changes are detected).
+In cryptography, we are considering only hash functions that can also satisfy these additional requirements:
+
+\begin{enumerate}
+  \item Deterministic - same input data are always mapped to the same hash.
+  \item One-way - it should be practically impossible to retrieve input data for a given hash.
+  \item Resistant to collisions - it should be practically impossible to find two inputs with the same hash.
+  \item Even small change to the input should lead to significantly different hash - so-called \textit{avalanche effect}.
+\end{enumerate}
+
+In the context of digital documents, we can use such hash function to generate a hash for the document.
+This hash can represent a document of any size and is used during the creation of a digital signature instad of document itself.
+Any further changes to the document will, therefore, lead to a significantly different hash, which means the digital signature will no longer be valid, and these changes are detected.
+For this purpose, only hash functions that are reasonably fast can be used, so that we can utilize them on documents of variable size.
+
+TODO: Add scheme of document > hash - valid signature, change in document > different hash - invalid signature.
+
+An example of commonly used hash function family is Secure Hash Algorithms (SHA), with hash functions relevant to this thesis being mainly SHA-2\footnote{SHA-2 specification available in FIPS PUB 180-4 at \url{https://nvlpubs.nist.gov/nistpubs/FIPS/NIST.FIPS.180-4.pdf}.}.
 
@@
 
-TODO: How it works.
+A common part of the paper signed documents is the date when they were signed.
+Their digital counterpart is PKI-based \textit{trusted timestamping} - a process of digitally signing date and time of document creation or modification.
+No one should be able to tamper with such timestamp - not even the author of the document.
+Additionally, we should not be able to choose arbitrary date and time ourselves, but it should represent the actual time at the signing.
 
-TODO: How it is used on the documents (parallels with the hand-written signature and date on the contract).
+To achieve this, a trusted third party - Timestamping Authority (TSA) - is the one that will sign the hash of the document together with the current date and time.
+This process can be seen in Figure \ref{fig:trusted-timestamping}\footnote{Schema created by Bart Van den Bosch, licensed under the CC BY-SA 2.0 BE license.}.
+Verifying can be by decrypting this information using the public key and comparing the hashes. Any attempt at tampering is therefore detected.
 
-TODO: Long term validation.
+\obrazok{./figures/trusted-timestamping}{Setting a timestamp via a trusted third party.}{trusted-timestamping}
+
+The commonly used standard for the PKI-based timestamping is part of the X.509\footnote{Specification available under RFC3161 at \url{https://www.ietf.org/rfc/rfc3161.txt}.}.
 
@@
 
-TODO: Quick explanation of each.
+Extensions for PDF files, XML files, and CMS signed data for signing using advanced electronic signatures (as explained in Chapter \ref{eidas}), are PAdES, XAdES, and CAdES, respectively.
+They are defined in the respective specifications by the ETSI \footnote{Specification for PAdES is available at \url{https://www.etsi.org/deliver/etsi_ts/102700_102799/10277801/01.01.01_60/ts_10277801v010101p.pdf}} \footnote{Specification for XAdES is available at \url{https://uri.etsi.org/01903/v1.3.2/ts_101903v010302p.pdf}} \footnote{Specification for CAdES is available at \url{https://www.etsi.org/deliver/etsi_ts/101700_101799/101733/02.01.01_60/ts_101733v020101p.pdf}}.
 
-TODO: How they can be used on a different file formats (table).
+When using PAdES, signatures are always part of the PDF file itself.
+This differs from XAdES, where the signature can be either part of the XML or provided as a separate file, or CAdES where signature is always in the form of binary data and it is up to the software.
+
+All of them support multiple signatures applied in succession.
+Implementation of the PAdES is rather portable since any compatible PDF file reader or editor will work with the signatures in a similar way - signature created in one should be easily verifiable in any other.
+XAdES and CAdES, however, are usually tied to specific implementation or standard that is built on top of them.
+
+\begin{table}[h!]
+  \begin{tabular}{ |p{2cm}||p{2.2cm}|p{2cm}|p{3cm}|p{3.9cm}| }
+    \hline
+    Standard & Usable with & Format & Multi-signature & Appearance \\
+    \hline
+    PAdES & PDF                  & embedded         & yes, sequential    & supported            \\
+    XAdES & XML, any             & XML              & yes                & depends on the usage \\
+    CAdES & any                  & binary           & yes                & depends on the usage \\
+    \hline
+  \end{tabular}
+
+  \caption{Overview of standards PAdES, XAdES, and CAdES}
+  \label{table:1}
+\end{table}
+
+The ability to retain signed documents for a prolonged period of time is covered in these standards under optional Long-Term Validation (LTV) that allows archiving of documents for many years, possibly decades.
+If enabled, a trustworthy timestamp is required as validity is verified for the time the signature (and therefore timestamp) was created.
+The expiration time is then limited by the validity of the timestamp certificate.
 
 \section{Software engineering}
+
+TODO: Give definition to the software engineering, what aspects we are going over, our focus on the engineering relevant for the open source.
+
+\subsection{Design}
+
+TODO: System modeling
+
+TODO: Architectural design
+
+TODO: Design patterns
+
+TODO: Dependability and Security
+
+\subsection{Implementation}
+
+TODO: Requirements engineering
+
+TODO: Software reuse
+
+TODO: Agile software development
+
+\subsection{Testing and Releasing}
+
+TODO: Quality management
+
+TODO: Automated tests (pyramid, V-model)
+
+TODO: Building, Release management
+
+\subsection{Maintenance}
+
+TODO: Evolution processes
+
+TODO: OSS Sustainability
+
+\chapter{Goals}
+
+\section{Software}
+
+TODO: Basis can be formed on the requirements article https://github.com/durasj/octosign/wiki/Initial-Requirements
+
+TODO: Basic software requirements and more specific software requirements.
+
+TODO: Enginnering parts - architecture, testing, maintainability.
+
+TODO: Backends
+
+\section{Information}
+
+TODO: Website - multilingual, editable by the community
+TODO: Global status - world map
 
@@
 
-\subsection{Design}
+\chapter{Analysis}
+
+\section{Review of existing software}
+
+TODO: Explain how they were picked, list them with info on the considered trait.
+
+TODO: Maybe add a simple comparison table? Consider: - format support, commercial, localization, eIDAS, web, required setup, advanced features
+
+\subsection{Desktop application JSignPdf}
+
+\subsection{Web application zep.disig.sk}
+
+\subsection{PDF viewer Adobe Acrobat Reader DC}
+
+\subsection{Commercial desktop application D.PDF Signer}
+
+\section{Modularity}
+
+TODO: Rough overview of different ways modular architecture can be achieved. Focus on modularity via multiple independant applications.
+
+\subsection{Communication}
+
+TODO: Communication in multiple independant applications - IPC, Sockets, Pipes, STDIO.
+
+\subsection{Bundling}
+
+\subsection{Communicating and bundling with GPL software}
+
+\chapter{Results}
+
+\section{Desktop software}
+
+TODO: Structure shoul roughly follow goals
 
@@
 
-\subsection{Implementation}
+\subsection{Comparison with existing software}
 
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-Integer lacinia, nulla porta varius tempus, lacus metus blandit
-lorem, a rutrum justo wisi id sapien. Integer risus libero,
-feugiat eleifend, ornare ac, volutpat nec, sem. In facilisis,
-quam eu elementum aliquet, lorem quam euismod dui, aliquet
-laoreet purus ipsum ac quam.
+TODO: Reference table from analysis, add new column
 
-\subsection{Maintenance}
+\section{Website}
 
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-Integer lacinia, nulla porta varius tempus, lacus metus blandit
-lorem, a rutrum justo wisi id sapien. Integer risus libero,
-feugiat eleifend, ornare ac, volutpat nec, sem. In facilisis,
-quam eu elementum aliquet, lorem quam euismod dui, aliquet
-laoreet purus ipsum ac quam.
-
-\chapter{Goals and analysis}
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-Integer lacinia, nulla porta varius tempus, lacus metus blandit
-lorem, a rutrum justo wisi id sapien. Integer risus libero,
-feugiat eleifend, ornare ac, volutpat nec, sem. In facilisis,
-quam eu elementum aliquet, lorem quam euismod dui, aliquet
-laoreet purus ipsum ac quam.
-
-\chapter{Suggested solution}
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-Integer lacinia, nulla porta varius tempus, lacus metus blandit
-lorem, a rutrum justo wisi id sapien. Integer risus libero,
-feugiat eleifend, ornare ac, volutpat nec, sem. In facilisis,
-quam eu elementum aliquet, lorem quam euismod dui, aliquet
-laoreet purus ipsum ac quam.
-
-\chapter{Results}
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-Integer lacinia, nulla porta varius tempus, lacus metus blandit
-lorem, a rutrum justo wisi id sapien. Integer risus libero,
-feugiat eleifend, ornare ac, volutpat nec, sem. In facilisis,
-quam eu elementum aliquet, lorem quam euismod dui, aliquet
-laoreet purus ipsum ac quam.
+TODO: Structure shoul roughly follow goals
 
@@
   %  ------
-  \bibitem{1}
-  \emph{§ 40 ods. 3 zákona č. 40/1964 Zb. občiansky zákonník}
+  \bibitem{civilcode}
+  \emph{Act No. 40/1964 Coll. Civil Code}
 
-  \bibitem{2}
-  \emph{§ 476 ods. 2 zákona č. 513/1991 Zb. obchodný zákonník}
+  \bibitem{notarylaw}
+  \emph{Act No. 323/1992 Coll. Notary Law}
 
-  \bibitem{3}
-  \emph{§ 56 ods. 1 zákona č. 323/1992 Zb. notársky poriadok}
+  \bibitem{eidas}
+  \emph{REGULATION (EU) No 910/2014 OF THE EUROPEAN PARLIAMENT AND OF THE COUNCIL, OJ L 257, 28.8.2014, p. 73–114}
 
-  \bibitem{4}
-  \emph{Article 25, Regulation (EU) No 910/2014}
-
-  \bibitem{5}
-  \emph{Article 26, Regulation (EU) No 910/2014}
-
-  \bibitem{6}
-  \emph{Council Directive 91/250/EEC of 14 May 1991 on the legal protection of computer programs}
+  \bibitem{eeccopyright}
+  \emph{Council Directive 91/250/EEC of 14 May 1991 on the legal protection of computer programs, OJ L 122, 17.5.1991, p. 42–46}
 
@@
   %  ------
-  \bibitem{7}
+  \bibitem{licensing}
   \osoba{Morin, A.}, et al.
@@
 2020-01-27].
-e1002598.
 Available at: \url{https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1002598}
+
+  \bibitem{cryptotxtbook}
+  \osoba{PAAR, Christof} and \osoba{Pelzl, Jan}, 2009. \emph{Understanding Cryptography - A Textbook for Students and Practitioner}. Berlin: Springer. ISBN 978-3-642-04100-6.
+
+  \bibitem{cryptojoy}
+  \osoba{ROSULEK, Mike}.
+\emph{The Joy of Cryptography} [online].
+Oregon: School of Electrical Engineering \& Computer Science, Corvallis, Oregon, USA [cit.
+2020-04-18].
+Available at: \url{http://web.engr.oregonstate.edu/rosulekm/crypto/crypto.pdf}
+
+  \bibitem{rsa}
+  \osoba{RIVEST, R.
+L.}, \osoba{SHAMIR, A.}, and \osoba{ADLEMAN, L.}, 1978.
+\emph{A method for obtaining digital signatures and public-key cryptosystems}.
+In: \emph{Communications of the ACM, 21(2):120–126, February 1978.}
+
+  \bibitem{law}
+  \osoba{MASON, Stephen}, 2017.
+\emph{Electronic Signatures in Law: Fourth Edition}.
+London: University of London.
+ISBN 978-1-911507-01-7.
+Available at: \url{https://humanities-digital-library.org/index.php/hdl/catalog/view/electronicsignatures/1/86-1}
+
+  \bibitem{sommerville}
+  \osoba{SOMMERVILLE, Ian}, 2016. \emph{Software Engineering, 10th Edition}. Harlow: Pearson Education. ISBN 978-0-13-394303-0.
+
+  \bibitem{oss}
+  \osoba{FOGEL, Karl}, 2019. \emph{Producing Open Source Software} [online]. TODO: TODO. [cit. 2020-04-21]. Available at: \url{https://producingoss.com/en/index.html}
+
+  \bibitem{osslicensing}
+  \osoba{LAURENT, Andrew}, 2008. \emph{Understanding Open Source and Free Software Licensing}. Sebastopol: O’Reilly Media. ISBN 978-0596005818.
+
+  \bibitem{eloquentjs}
+  \osoba{HAVERBEKE, Marijn}, 2018. \emph{Eloquent JavaScript, 3rd Edition}. San Francisco: No Starch Press. ISBN 978-1593279509.
+
+  \bibitem{ydkjs}
+  \osoba{SIMPSON, Kyle}, 2020. \emph{You Don't Know JS Yet (book series) - 2nd Edition} [online]. Available at: \url{https://github.com/getify/You-Dont-Know-JS}
 

### 2020-05-07 paper/main.tex
Add final draft for the SVK

@@
-\documentclass[thesismargins, english, thesislinespacing, twoside, openright, upjsfrontpage]{rnthesis}
+\documentclass[thesismargins, english, thesislinespacing, onelinechapterstyle, upjsfrontpage]{rnthesis}
 \usepackage[english]{babel}
@@
 
+\usepackage{xparse}
+% Rotation: \rot[<angle>][<width>]{<stuff>}
+\NewDocumentCommand{\rot}{O{60} O{1em} m}{\makebox[#2][l]{\rotatebox{#1}{#3}}}%
+
 \title{Replacing handwritten signatures with open electronic signature software}
 \author{Jakub Ďuraš}
+\iffalse % Not part of the version for ŠVK
 \typprace{bachelor's}
+\fi
+\inytypprace{Student Scientific Conference}
 \rok{2020}
@@
 \miesto{Košice}
-\podakovanie{
-  Thanks to the Planet!
-}
 \veduci{RNDr. Viliam Kačala}
 \pracovisko{Institute of Computer Science}
+\iffalse % Not part of the version for ŠVK
+\podakovanie{
+  I would like to thank my supervisor Viliam Kačala for his continuous feedback and patience with my question.
+
+  I would also like to acknowledge help received from Pavol Sokol and Laura Rózenfeldová on legal questions, and access to various resources from Gabriel Semanišin.
+
+  Finally, I would like to thank my family and friends who contributed in various ways to my thesis.
+}
+\fi
+
+\iffalse % Not part of the version for ŠVK
 \pdfzadanie{zadanie.pdf}
+\fi
 
 \abstract{
-With the recent changes in the legal status of electronic signatures in many parts of the world, there is a need for easily accessible solutions intended as an alternative to handwritten signatures.
-This may be more necessary than ever since electronic communication is the preferred way of communication.
-This bachelor thesis aims to explore the principles and legal status of electronic signatures, review current electronic signature software, and explore possible obstacles the open-source community is facing to develop such specialized applications.
-We propose an open-source, cross-platform, and user-friendly software compliant with the eIDAS Regulation (Regulation No 910/2014).
-Our application should allow ordinary users to quickly sign and verify signatures of different types of documents and therefore easily use them in everyday life.
+With the recent changes in the legal status of electronic signatures in the world, there is an opportunity to expand the use of electronic signatures as an alternative to handwritten signatures outside the usual areas.
+Electronic signatures are becoming increasingly important in the age of electronic communication when handwritten signatures are unpractical due to the need to print, personally exchange, or send signed documents.
+We devote our attention to the exploration of legal preconditions relevant for the EU, technological preconditions, review of the existing software, and proposal of our own software platform.
+The result is open, user-friendly software for Windows, Mac OS X, and Linux.
+Using the application, it is possible, following the eIDAS Regulation, to sign and verify the signature of any document with the legal effect of an attested signature.
+Our solution supports ID cards of the Slovak Republic with the eID chip without any modifications or settings, which enables easy use outside the usual use in web applications of the state.
+The application is not tied only to the Slovak eID but is ready for further use with various devices in different countries and as a platform for other types of signatures.
+Another part is a website that, in addition to the possibility of downloading software and brief documentation, provides general information about electronic signatures so that the project is as accessible to the public as possible.
+We concentrate on simple extensibility of the program via language translations, documentation, functionality, and general information.
 }
@@
 \abstrakt{
-Abstrakt v SK jazyku bude pridaný neskôr ako voľný preklad z EN verzie (pozri ďalšiu stranu). Predíde sa tak zbytočnej práci keďže EN abstrakt sa môže časom meniť.
+So zmenami v právnom statuse elektronických podpisov vo svete sa vytvára príležitosť rozšíriť elektronické podpisovanie ako náhradu pre vlastnoručné podpisy mimo obvyklých oblastí.
+Elektronické podpisy sú dôležité čoraz viac v dobe elektronickej komunikácie kedy sú vlastnoručné podpisy nepraktické z dôvodu nutnosti tlače, osobného kontaktu alebo zasielania podpísaných dokumentov.
+V práci sa venujeme právnym predpokladom relevantným pre EÚ, technologickým predpokladom, prehľadu v súčasnosti dostupného softvéru a návrhu vlastnej softvérovej platformy.
+Výsledkom je otvorený, používateľsky prívetivý softvér pre platformy Windows, Mac OS X a Linux.
+Pomocou aplikácie je možné, v súlade s nariadením eIDAS, podpísať a overiť podpis ľubovoľného dokumentu, ktorý má rovnaký právny účinok ako osvedčený podpis.
+Naše riešenie podporuje občianske preukazy Slovenskej Republiky s čipom eID bez akýchkoľvek úprav alebo nastavení, čím sa umožní ich jednoduché využitie mimo obvyklého použitia vo webových aplikáciách štátu.
+Aplikácia nie je viazaná iba na slovenské eID, ale je pripravená na ďalšie použitie s rôznymi zariadeniami v rôznych štátoch a ako platforma pre iné druhy podpisov.
+Súčasťou je aj webová stránka, ktorá okrem možnosti stiahnutia softvéru a stručnej dokumentácie poskytuje všeobecné informácie o elektronických podpisoch aby bol projekt čo najviac prístupný verejnosti.
+Dôraz kladieme na jednoduchú rozšíriteľnosť programu vo forme jazykových prekladov, dokumentácie, funkcionality a všeobecných informácií.
 }
@@
 \maketitle
+
 \newpage
+
 \tableofcontents
+\iffalse % Not part of the version for ŠVK
 \listoffigures
 \listoftables
-% zoznam značiek a skratiek. Pre tento koncept neexistuje
-% LaTeXovsky príkaz
+\fi
 
@@
 
-Will be added at the end. Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-Integer lacinia, nulla porta varius tempus, lacus metus blandit
-lorem, a rutrum justo wisi id sapien. Integer risus libero,
-feugiat eleifend, ornare ac, volutpat nec, sem. In facilisis,
-quam eu elementum aliquet, lorem quam euismod dui, aliquet
-laoreet purus ipsum ac quam.
+Electronic signatures have been around for a while.
+Their practical usability, however, is influenced by their legal status and available infrastructure.
+With the recent changes in the relevant law around the world and the European Union specifically, they see increased adoption.
 
-\chapter{Background and theory}
+While, specifically in the European Union, they seem to be relatively extensively used in communication with the government, they do not seem to be widely adopted in an exchange between individuals or smaller organizations.
+The problem is handwritten signatures are often unpractical for electronic communication due to the need to print, personally exchange, or send signed documents.
 
-In our work, we rely on knowledge in several different subjects that we go over in the following sections.
-TODO: Sounds very meh for the very first sentece of the thesis.
-Firstly, the current state of the law and recent changes relating to the use of electronic signatures.
-Secondly, on the well-established concepts in cryptography like asymmetric cryptography or hashing.
-Lastly, expertise from the field of software engineering - especially design, implementation, and maintenance of computer software.
-All in relation to the electronic signatures, and, where applicable, open-source and desktop computer software.
+Our goal is to explore an accessible way to use electronic signatures.
+In the beginning, we go over the preconditions that make it a viable alternative to the handwritten signatures.
+We cover the legal preconditions relevant for the European Union and Slovakia in particular.
+Also, we explain technical preconditions focusing on the cryptographic concepts like asymmetric cryptography, timestamping, hashing, and standards that comply with the law within the European Union.
+Secondly, we do a review of the existing software.
+We go quickly over several applications that are meant for basic use.
+We compare them using several criteria to find out the strengths and weaknesses of the current implementations and identify a potential space for improvements.
+Finally, we devote our attention to our own implementation.
+We explain the proposed desktop software and go into details on software engineering aspects.
+Specifically, modularity, architecture, usability, localization, testing, and automation.
+In the end, we compare it with the existing software and identify possible future improvements.
+Another result we focus on is a supporting website and how it helps with the use of the software.
+All in a way open for extensibility by the community.
 
+\chapter{Preconditions and theory}
+
+There are several preconditions that allow us to consider electronic signatures as an alternative.
+Firstly, the current state of the law and recent changes relating to the use of electronic signatures, secondly, well-established concepts in cryptography like \pojem{asymmetric cryptography} or \pojem{hashing}, and lastly, expertise from the field of software engineering - design, implementation, and maintenance of computer software.
+We will concetrate on the first two in relation to the electronic signatures, and, where applicable, open-source desktop computer software.
+
+\iffalse % Not part of the version for ŠVK
 TODO: Think about adding section on handwritten signatures
+\fi
 
@@
 
-Unless specified otherwise, we are considering law applicable locally in the Slovak Republic (SR).
-Its law is greatly influenced by the European Union (EU), being its member since 2004, and to some extent by the rest of the world.
-Therefore, we can assume this is at least partially applicable outside of the SR.
+We are considering law applicable locally in the Slovak Republic (SR).
+Its law is greatly influenced by the European Union (EU), being its member since 2004, and by the rest of the world.
+We can assume this is at least partially applicable outside of the SR as well.
 
@@
 
-Signatures are essential part of the written legally binding documents like contracts.
+Signatures are an essential part of the written legally binding documents like contracts.
 They are permanently affixed to the document and are supposed to uniquely identify the person and its deliberate, informed consent.
-As can be seen in the Slovak Civil Code, "A written legal act is valid if signed by the acting person;" \cite{civilcode}.
-Law often explicitly requires signatures and further clarifies their expected use.
-For example, when selling an enterprise, "The contract requires written form and attested signatures of the seller and the buyer." \cite{civilcode}.
+As can be seen in the Slovak Civil Code, § 40, a written legal act is valid if signed by the acting person \cite{civilcode}.
+Our law often explicitly requires signatures and further clarifies their expected use.
+For example, when selling an enterprise, looking at the Commercial Code, § 476, the contract requires written form and attested signatures of the seller and the buyer \cite{commercialcode}.
 
-Signature is considered an "attested signature"\footnote{Translation from "osvedčený podpis" as used in the Slovak law.} if it is attested by the authorized third party.
-Such process, known as legalization, is regulated by the law and its purpose is to attest information which could form the basis for the exercise of rights or which could cause legal consequences \cite{notarylaw}.
+Signature is considered an "attested signature"\footnote{Translation from "osvedčený podpis" as used in the Slovak law.} if it is verified by an authorized third party.
+This process is referred to as legalization in Notary Law, § 56, and its purpose is to attest information that could form the basis for the exercise of rights or which could cause legal consequences \cite{notarylaw}.
 
-With regards to electronic signatures, law within the EU used to differ, with law applicable in the SR being now repealed "Act No. 215/2002 Coll.".
+With regards to electronic signatures, law within the EU used to differ, with the law applicable in the SR being now-repealed \pojem{Act No.
+215/2002 Coll.}.
+This situation changed with the EU Regulation eIDAS.
 
@@
 
-Advanced electronic signatures are a subset of electronic signatures that have to \textbf{uniquely link and identify} the signature author, be created in a way that is \textbf{possible only by them} and any \textbf{changes to the signed document have to be detectable} \cite{eidas}.
-From the technical point of view, standards PAdES, XAdES, and CAdES specified by the European Telecommunications Standards Institute (ETSI) comply with these requirements.
-We explain why this is the case in Chapter \ref{cryptography}.
+\pojem{Advanced electronic signatures} are a subset of electronic signatures that have to \textbf{uniquely link and identify} the signature author, be created in a way that is \textbf{possible only by them} and any \textbf{changes to the signed document have to be detectable} \cite{eidas}.
+From the technical point of view, standards \pojem{PAdES}, \pojem{XAdES}, and \pojem{CAdES} specified by the European Telecommunications Standards Institute (ETSI) comply with these requirements.
+We explain why this is the case in Section \ref{cryptography}.
 
-Qualified electronic signatures (QES) are a subset of advanced electronic signatures with more specific requirements that can be found in the annex 1 of the eIDAS regulation\footnote{Available at https://eur-lex.europa.eu/legal-content/EN/TXT/HTML/?uri=CELEX:32014R0910\#d1e32-111-1.}.
-In more practical terms, such advanced electronic signatures are created with a qualified device using a qualified certificate and a qualified trust service.
-Qualified in this case means authorized for such use by the legal authorities.
+\pojem{Qualified electronic signatures} (QES) are a subset of \pojem{advanced electronic signatures} with more specific requirements that can be found in annex 1 of the eIDAS regulation\footnote{Available at https://eur-lex.europa.eu/legal-content/EN/TXT/HTML/?uri=CELEX:32014R0910\#d1e32-111-1.}.
+In more practical terms, such \pojem{advanced electronic signatures} are created with a qualified device using a qualified certificate and a qualified trust service.
+Qualified, in this case, means authorized for such use by the legal authorities.
+These devices, \pojem{certificates}, and trust services are explicitly listed by the EU. In many cases indirectly by listing the organization (and its items) from a specific country.
 
-Furthemore, a qualified electronic signature has the legal effect of a handwritten signature \cite{eidas} and attested handwritten signatures in the SR.
+Furthermore, a qualified electronic signature has the legal effect of a handwritten signature \cite{eidas} and attested handwritten signatures in the SR.
 
-Mentioned terminology is used when considering natural person. Legal entities are able to use "electronic seals", which are, from the technical point of view and our purposes, identical.
+Mentioned terminology is used when considering a natural person. Legal entities are able to use \pojem{electronic seals}, which are, from the technical point of view and for our purposes, identical.
 
-Overall, electronic signatures are supposed to be, within the EU, interoperable and transparent alternative to hand-written signatures.
+Overall, electronic signatures are supposed to be, within the EU, interoperable and transparent alternative to handwritten signatures.
 
-\subsection{Software copyright}
+\subsection{Software copyright} \label{copyright}
 
@@
 
-Such right can be exercised, as far as we are concerned, to license the software under either proprietary or open-source license.
-Proprietary meaning under exclusive legal right of the author, typically also confidential and distributed as a paid product.
-Open-source meaning having its source code freely available, typically also distributed for free and without any liability (the software is provided "as is").
+Such right can be exercised to license the software under either \pojem{proprietary} or \pojem{open-source} license.
+\pojem{Proprietary} meaning under the exclusive legal right of the author, typically also \textbf{confidential and distributed as a paid product}.
+\pojem{Open-source} meaning having its source code \textbf{freely available}, typically also \textbf{distributed for free} and without any liability (the software is provided "as is").
 
-Motivation to license the software under an open-source license can differ, and so do such licenses.
-While a proprietary license is usually made specifically for that entity and its interests, open-source licenses tend to be reused between different authors.
-This means consumers of the software can quickly recognize their rights and responsibilities if they decide to use, modify, or distribute the software.
-In general, we can classify the open-source licenses into two categories: permissive and copyleft.
+Motivation to license the software under an \pojem{open-source} license can differ, and so do such licenses.
+While a \pojem{proprietary} license is usually made specifically for that entity and its interests, \pojem{open-source} licenses tend to be reused between different authors.
+This means that consumers of the software can quickly recognize their rights and responsibilities if they decide to use, modify, or distribute the software.
+In general, we can classify the \pojem{open-source} licenses into two categories: \pojem{permissive} and \pojem{copyleft}.
 
-Copyleft license, in general, requires the user to publish the modified work under the compatible (free) license.
-Meaning forcing them to extend the rights they have received onto others and, in turn, somehow hindering use in software licensed under a different license.
-Popular example of such license is the GNU General Public License (GPL) and its derivatives like GNU Lesser GPL (LGPL) or GNU Affero GPL (AGPL).
+\pojem{Copyleft} license, in general, requires the user to publish the modified work under the compatible (free) license.
+As a result, it forces them to extend the rights they have received onto others and, in turn, somehow hinder usability in software licensed under a different license.
+A popular example of such license is the GNU General Public License(GPL) and its derivatives like GNU Lesser GPL (LGPL) or GNU Affero GPL (AGPL).
 
-Permissive license, on the other hand, does not have such requirement in place and is more suitable for potential commercialization.
-They typically allow commercial use, modifications, distribution or sublicensing as long as the author is not held liable and the original copyright and license is distributed with the software.
-Often used permissive licenses are MIT License, Apache License or BSD License.
+\pojem{Permissive} licenses, on the other hand, do not have such a requirement in place and are more suitable for potential commercialization.
+They typically allow commercial use, modifications, distribution or sublicensing as long as the author is not held liable, and the original copyright and license are distributed with the software.
+Often used permissive licenses are MIT License, Apache License, or BSD License.
 
-It is not uncommon to see hybrid licensing - licensing under more than one license, with one being typically open-source and one proprietary.
+It is not uncommon to see \pojem{hybrid licensing} - licensing under more than one license, with one being typically \pojem{open-source} and one \pojem{proprietary}.
 In this scenario, users can choose which license they want to use based on their needs.
@@
 As we have already mentioned, software licenses can influence whether and how it can be used as a part of different software.
-Both when we are using someone else's software and when someone else is using our software.
-Generally speaking, the more open the license is, the higher the probability it will not stop adoption for legal reasons.
+Both when we are the ones using someone else's software and when someone else is using our software.
+Generally speaking, the more open the license is, the higher the probability it will not prevent adoption for legal reasons.
 Various possible combinations can be seen in Figure \ref{fig:license-compatibility}, where greyed out options mean such combination is not possible unless a separate licensing agreement is reached with the copyright owner.
 
-\obrazok{./figures/license-compatibility}{Schematic representation of license directionality.\cite{licensing}}{license-compatibility}
+\obrazok{./figures/license-compatibility}{Schematic representation of license directionality. \osoba{Morin 2012} \cite{licensing}}{license-compatibility}
 
@@
 
-Previously mentioned standards PAdES, XAdES, and CAdES define how asymmetric cryptography, hashing, or timestamping should be used to comply with the standards.
-We go over a few of these topics and why they are able to satisfy the requirements.
+Previously mentioned standards \pojem{PAdES}, \pojem{XAdES}, and \pojem{CAdES} define how \pojem{asymmetric cryptography}, \pojem{hashing}, or \pojem{timestamping} should be used to comply with the standards.
+We go over a these topics and why they are able to satisfy the requirements for QES mentioned in the Section \ref{eidas}.
 
@@
 There are several issues that make symmetric encryption unsuitable.
-Let us imagine we are creating a key for each pair of users in our network - imaginary country Isle of Alice and Bob - for symmetric encryption and decryption of documents shared between each Alice and Bob.
+Let us imagine we are creating a key for each pair of users in our network - imaginary country Isle of Alices and Bobs - for symmetric encryption and decryption of documents shared between each Alice and Bob.
 
 \textbf{Secure distribution of keys} seems to be feasible when we are adding a new user to our network.
-We require each user to personally come and pick up physical device that contains pregenerated keys.
-Changes made after that, however, would mean we either need to require periodic personal visit to update the saved keys or transmission of these keys over some medium which can be potentialy unsafe.
-Any new members of the network are unable to participate untill other Alices and Bobs get their updates.
+We require each user to personally come and pick up a physical device that contains pre-generated keys.
+Changes made after that, however, would mean we either need to require a periodic personal visit to update the saved keys or transmission of these keys over some medium, which can be potentially unsafe.
+Any new members of the network are unable to participate until other Alices and Bobs get their updates.
 
@@
 
-which means that even our little country with a population of only 70 000 requires roughly 2.5 billion keys.
-However, since our signature system needs to be portable, we will need to count also with a population of other EU countries, and the number of keys quickly explodes.
+which means that even our little imaginary country with a population of only 70 000 requires roughly 2.5 billion keys.
+However, since our signature system needs to be portable, we will need to count also with a population of other countries, and so the number of keys grows very quickly.
 
+\iffalse % Not part of the version for ŠVK
 \begin{figure}[!htb]
@@
 \end{figure}
+\fi
 
@@
 
-These issues can be addressed with asymmetric cryptography.
-In general, we assign a pair of keys to each person, one of them used for encryption (public key), another used for decryption (private key).
+These issues can be addressed with \pojem{asymmetric cryptography}.
+In general, we assign \textbf{a pair of keys} to each person, one of them used for encryption, known as a \pojem{public key}, another used for decryption, known as a \pojem{private key}.
 One party can encrypt, and the other is the only one that can decrypt.
 
-Digital signatures are a specific public-key algorithms.
-The signer is the only one who poses private key used for signing, and others can use the public key for verifying of signatures as can be seen in Figure \ref{fig:pubkey-scheme} where \(x\) is the "document", and \(s\) is the signature that is sent with the document.
-Alice can then sign the document \textit{including} Bob's signature.
+\pojem{Digital signatures} are specific \pojem{asymmetric cryptography} algorithms.
+The signer is the only one who possesses the \pojem{private key} used for signing, and others can use the \pojem{public key} for signature verification.
+We \textbf{sign} the document, including the other parties' signature.
 
+\iffalse % Not part of the version for ŠVK
 \obrazok{./figures/digital-signature-scheme}{Principle of digital signatures which involves signing and verifying a message (Paar 2009). TODO: I don't have rights to this image!!!}{pubkey-scheme}
+\fi
 
-Practical, widely used implementation of this idea is the RSA signature scheme that relies on the integer factorization problem (difficulty of factoring a product of two large primes) published in 1978 \cite{rsa}.
+Practical, widely used implementation of this idea is the \pojem{RSA} signature scheme that relies on the integer factorization problem (difficulty of factoring a product of two large primes) published in 1978 \cite{rsa}.
 
@@
 This ID card (used together with a card reader) provides an API for safe access, management, and cryptoprocessing.
-Such device is in general called a Hardware Security Module (HSM), and the API relevant for this thesis is PKCS \#11 \footnote{PKCS \#11 specification available at \url{http://docs.oasis-open.org/pkcs11/pkcs11-base/v2.40/os/pkcs11-base-v2.40-os.html}.}.
+Such a device is in general called a \pojem{Hardware Security Module} (HSM), and the API relevant for this thesis is \pojem{PKCS \#11} \footnote{PKCS \#11 specification available at \url{http://docs.oasis-open.org/pkcs11/pkcs11-base/v2.40/os/pkcs11-base-v2.40-os.html}.}.
 
-We are left with the problem of public key distribution - even though public-key schemes do not require a \textit{secure channel}, they require \textit{authenticated channels} for the distribution of the public keys \cite{cryptotxtbook}.
+We are left with the \textbf{problem of public key distribution} - even though \pojem{public-key schemes} do not require a \textit{secure channel}, they require \textit{authenticated channels} for the distribution of the public keys \cite{cryptotxtbook}.
 
-Common solution is the use of \textit{certificates}. Certificates are essentially digital signatures with metadata that can be used to establish validity of the received public key before it is used.
-Distribution of such certificates is the responsibility of a Certification Authority (CA) - a third party that all users in the network trust.
-In the case of the Slovak ID cards, it is Disig - SVK eID ACA issued by the National Security Authority (NBÚ in Slovak).
-Such certificates therefore reliably uniquely link and identify the public key owners as can be seen in Figure \ref{fig:certificate-scheme}\footnote{Created by unknown author, licensed under the CC BY-SA 3.0 license.}.
+A common solution is the use of \pojem{certificates}.
+\pojem{Certificates} are essentially \pojem{digital signatures} with metadata that can be used to establish the validity of the received \pojem{public key} before it is used.
+Distribution of such \pojem{certificates} is the responsibility of a \pojem{Certification Authority} (CA) - a third party that all users in the network trust.
+In the case of the Slovak ID cards, it is Disig - SVK eID Accredited CA, issued by the National Security Authority\footnote{Translation from "Národný bezpečnostný úrad".}.
+Such \pojem{certificates}, therefore, \textbf{reliably uniquely link and identify} the \pojem{public key} owners, as can be seen in Figure \ref{fig:certificate-scheme}\footnote{Created by an unknown author, licensed under the CC BY-SA 3.0 license.}.
+Commonly used standard for public key certificates is X.509\footnote{Specification available under RFC5280 at \url{https://tools.ietf.org/html/rfc5280}.}.
 
-\obrazok{./figures/certificate-scheme}{Application of digital signature with certificate. TODO: Modify to add CA, remove hashing}{certificate-scheme}
+\obrazok{./figures/certificate-scheme}{Application of digital signature with certificate.}{certificate-scheme}
 
-Commonly used standard for public key certificates is X.509\footnote{Specification available under RFC5280 at \url{https://tools.ietf.org/html/rfc5280}.}.
+\iffalse % Not part of the version for ŠVK
+TODO: Mention EU LOTL
+\fi
 
@@
 
-It is often useful to map data of arbitrary size to fixed-length (typically short) set of bits.
-This process is commonly called \textit{hashing}, function that can be used for it \textit{hash function}, and output of such function \textit{hash}.
+It is often useful to map data of arbitrary size to a fixed-length (typically short) set of bits.
+This process is commonly called \pojem{hashing}, a function that can be used for it \pojem{hash function}, and output of such function \pojem{hash}.
 
-In cryptography, we are considering only hash functions that can also satisfy these additional requirements:
+In cryptography, we are considering only \pojem{hash functions} that can also satisfy these additional \textbf{requirements}:
 
 \begin{enumerate}
-  \item Deterministic - same input data are always mapped to the same hash.
-  \item One-way - it should be practically impossible to retrieve input data for a given hash.
-  \item Resistant to collisions - it should be practically impossible to find two inputs with the same hash.
-  \item Even small change to the input should lead to significantly different hash - so-called \textit{avalanche effect}.
+  \item \textbf{Deterministic} - same input data are always mapped to the same \pojem{hash}.
+  \item \textbf{One-way} - it should be practically impossible to retrieve input data for a given \pojem{hash}.
+  \item \textbf{Resistant to collisions} - it should be practically impossible to find two inputs with the same \pojem{hash}.
+  \item Even small change to the input should lead to significantly different \pojem{hash} - so-called \textbf{avalanche effect}.
 \end{enumerate}
 
-In the context of digital documents, we can use such hash function to generate a hash for the document.
-This hash can represent a document of any size and is used during the creation of a digital signature instad of document itself.
-Any further changes to the document will, therefore, lead to a significantly different hash, which means the digital signature will no longer be valid, and these changes are detected.
-For this purpose, only hash functions that are reasonably fast can be used, so that we can utilize them on documents of variable size.
+In the context of \pojem{digital documents}, we can use such \pojem{hash function} to generate a \pojem{hash} for the document.
+This \pojem{hash} can represent a document of any size and is used during the creation of a \pojem{digital signature} instead of the document itself.
+Any further changes to the document will, therefore, lead to a significantly different \pojem{hash}, which means the \pojem{digital signature} will no longer be valid, and these \textbf{changes are detected}.
+For this purpose, only \pojem{hash functions} that are reasonably fast can be used, so that we can utilize them on documents of variable size.
 
+\iffalse % Not part of the version for ŠVK
 TODO: Add scheme of document > hash - valid signature, change in document > different hash - invalid signature.
+\fi
 
-An example of commonly used hash function family is Secure Hash Algorithms (SHA), with hash functions relevant to this thesis being mainly SHA-2\footnote{SHA-2 specification available in FIPS PUB 180-4 at \url{https://nvlpubs.nist.gov/nistpubs/FIPS/NIST.FIPS.180-4.pdf}.}.
+An example of commonly used \pojem{hash function} family is \pojem{Secure Hash Algorithms} (SHA), with \pojem{hash functions} relevant to this thesis being mainly SHA-2\footnote{SHA-2 specification available in FIPS PUB 180-4 at \url{https://nvlpubs.nist.gov/nistpubs/FIPS/NIST.FIPS.180-4.pdf}.}.
 
@@
 A common part of the paper signed documents is the date when they were signed.
-Their digital counterpart is PKI-based \textit{trusted timestamping} - a process of digitally signing date and time of document creation or modification.
-No one should be able to tamper with such timestamp - not even the author of the document.
+Their digital counterpart is PKI-based \pojem{trusted timestamping} - a process of \pojem{digitally signing} date and time of document creation or modification.
+No one should be able to tamper with such \pojem{timestamp} - not even the author of the document.
 Additionally, we should not be able to choose arbitrary date and time ourselves, but it should represent the actual time at the signing.
 
-To achieve this, a trusted third party - Timestamping Authority (TSA) - is the one that will sign the hash of the document together with the current date and time.
+To achieve this, a trusted third party - \pojem{Timestamping Authority} (TSA) - is the one that will sign the \pojem{hash} of the document together with the current date and time.
 This process can be seen in Figure \ref{fig:trusted-timestamping}\footnote{Schema created by Bart Van den Bosch, licensed under the CC BY-SA 2.0 BE license.}.
-Verifying can be by decrypting this information using the public key and comparing the hashes. Any attempt at tampering is therefore detected.
+Verifying can be done by decrypting this information using the \pojem{public key} and comparing the \pojem{hashes}. Any attempt at \textbf{tampering is therefore detected}.
 
@@
 
-The commonly used standard for the PKI-based timestamping is part of the X.509\footnote{Specification available under RFC3161 at \url{https://www.ietf.org/rfc/rfc3161.txt}.}.
+The commonly used standard for the \pojem{Public Key Infrastructure} (PKI) based timestamping is part of the X.509\footnote{Specification available under RFC3161 at \url{https://www.ietf.org/rfc/rfc3161.txt}.}.
 
@@
 
-Extensions for PDF files, XML files, and CMS signed data for signing using advanced electronic signatures (as explained in Chapter \ref{eidas}), are PAdES, XAdES, and CAdES, respectively.
-They are defined in the respective specifications by the ETSI \footnote{Specification for PAdES is available at \url{https://www.etsi.org/deliver/etsi_ts/102700_102799/10277801/01.01.01_60/ts_10277801v010101p.pdf}} \footnote{Specification for XAdES is available at \url{https://uri.etsi.org/01903/v1.3.2/ts_101903v010302p.pdf}} \footnote{Specification for CAdES is available at \url{https://www.etsi.org/deliver/etsi_ts/101700_101799/101733/02.01.01_60/ts_101733v020101p.pdf}}.
+Extensions for PDF files, XML files, and CMS signed data for signing using \pojem{advanced electronic signatures} (as explained in Section \ref{eidas}), are \pojem{PAdES}, \pojem{XAdES}, and \pojem{CAdES}, respectively.
+They are defined in the respective specifications by the ETSI in \cite{pades}, \cite{xades}, and \cite{cades}.
 
 When using PAdES, signatures are always part of the PDF file itself.
-This differs from XAdES, where the signature can be either part of the XML or provided as a separate file, or CAdES where signature is always in the form of binary data and it is up to the software.
+This differs from XAdES, where the signature can be either part of the XML or provided as a separate file, or CAdES where the signature is always in the form of binary data, and it is up to the software.
 
 All of them support multiple signatures applied in succession.
+\iffalse % Not part of the version for ŠVK
 Implementation of the PAdES is rather portable since any compatible PDF file reader or editor will work with the signatures in a similar way - signature created in one should be easily verifiable in any other.
 XAdES and CAdES, however, are usually tied to specific implementation or standard that is built on top of them.
+\fi
 
@@
 
-The ability to retain signed documents for a prolonged period of time is covered in these standards under optional Long-Term Validation (LTV) that allows archiving of documents for many years, possibly decades.
+The ability to keep signed documents valid for a prolonged period of time is covered in these standards under optional \pojem{Long-Term Validation} (LTV) that allows \textbf{archiving} of documents for many years, possibly decades.
 If enabled, a trustworthy timestamp is required as validity is verified for the time the signature (and therefore timestamp) was created.
@@
 
+\iffalse % Not part of the version for ŠVK
 \section{Software engineering}
+
+We will consider Ian Sommerville's definition of software engineeing - "an engineering discipline that is concerned with all aspects of software production from initial conception to operation and maintenance" \cite{sommerville}.
+Aspects we will focus on are design, implementation, testing and releasing, and maintenance.
+Small
 
@@
 TODO: OSS Sustainability
+\fi
 
+\iffalse % Not part of the version for ŠVK
 \chapter{Goals}
@@
 TODO: Global status - world map
+\fi
 
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-Integer lacinia, nulla porta varius tempus, lacus metus blandit
-lorem, a rutrum justo wisi id sapien. Integer risus libero,
-feugiat eleifend, ornare ac, volutpat nec, sem. In facilisis,
-quam eu elementum aliquet, lorem quam euismod dui, aliquet
-laoreet purus ipsum ac quam.
+\chapter{Review of existing software} \label{review}
 
-\chapter{Analysis}
+We focus only on simple applications made mostly for the signing of the documents.
+During our review, we have used several PDF files that we have tried to sign using QES available on the SR eID.
+We have tried to include software of all types in our review, preferring the most popular, locally available ones, since those should be prepared for the use with the SR eID.
+Excluded are all complex applications with advanced features and high price tag.
+We have picked one open-source desktop application, one web application - Software as a Service (SaaS), one PDF viewer, and one simple paid desktop application. Criteria we focus on are:
 
-\section{Review of existing software}
+\begin{enumerate}
+  \item \textbf{Licensing} - is it freely available - both in terms of money and source code - best case is open-source; worst case is paid proprietary software.
+  \item \textbf{Cross-platform availability} - is it available for users of all major desktop platforms - best case is availability on Windows, Mac OS X, and Linux; worst case is availability only on one platform.
+  \item \textbf{Localization} - is it translated and does it display data like dates in local format - best case is availability at least in English and Slovak language; worst case is availability only in one language with no support for localization of dates, etc..
+  \item \textbf{File support} - does it support different file formats and sizes - best case is support for all file formats with sensible size up to 50 MB; worst case is availability only for one file format or strict limits on the file size.
+  \item \textbf{Easy setup} - does it require non-trivial setup to use the SR eID - best case is no setup required; worst case is possibly demanding setup or troubleshooting.
+  \item \textbf{Flexibility} - how much logic can be customized - best case is various options for signature format, signature policy, used algorithms or timestamping configuration; worst case is none or almost none customization.
+  \item \textbf{Built-in verification} - is the verification of signatures available in the same software - best case is ability to verify various signature types; worst case is no support for signature verification.
+  \item \textbf{Extensibility} - is it extensible by the community - best case is it is prepared for easy extensibility in terms of functionality and localization; worst case is no extensibility.
+\end{enumerate}
 
-TODO: Explain how they were picked, list them with info on the considered trait.
+\section{Desktop application JSignPdf}
 
-TODO: Maybe add a simple comparison table? Consider: - format support, commercial, localization, eIDAS, web, required setup, advanced features
+JSignPdf is an open-source Java application that adds digital signatures to PDF documents. It can be used as a standalone application or as an Add-On in OpenOffice. We focus only on standalone use.
 
-\subsection{Desktop application JSignPdf}
+\obrazok{./figures/screenshot-jsignpdf}{Screenshot of the JSignPdf version 1.6.4}{screenshot-jsignpdf}
 
-\subsection{Web application zep.disig.sk}
+The application consists only of one window, which makes it very simple at first sight.
+The very first input - "Keystore type" - can be harder to understand for the basic user since it contains over a dozen options in the form of abbreviations like "CASEEXACTJKS" or "PKCS12-DEF-3DES-40RC2".
+The default value is "WINDOWS-MY", however, which is what is exactly required to use the SR eID.
+Application has a toggle for the "Advanced view" that is off by default, but it is necessary to turn it on if we want to choose the correct key.
+On some of the computers, we were not able to see the key from the eID. On others, it took a while and several pressing of a button to load them.
+The correct key was not the first, preselected key in the form of "Name Surname", but the second "Name Surname (1)".
+With everything set up correctly, the application was successfully able to attach a QES.
+Even though the application looks pretty simple, it is still pretty flexible with its options regarding the used algorithms, TSA, OCSP, or CRL.
+That being said, it is available only in English, and we have also noticed issues with the diacritics that you can see in Figure \ref{fig:screenshot-jsignpdf} - "Jakub Ďuraš" is displayed as "Jakub Ïura".
 
-\subsection{PDF viewer Adobe Acrobat Reader DC}
+\section{Web application zep.disig.sk}
 
-\subsection{Commercial desktop application D.PDF Signer}
+Web application zep.disig.sk is a pro-bono SaaS from company Disig a.s. for user-friendly document signing and validation utilizing the currently bundled software with the SR eID.
 
+\obrazok{./figures/screenshot-zepdisigsk}{Screenshot of the zep.disig.sk}{screenshot-zepdisigsk}
+
+The process to sign or validate is very simple and straightforward.
+There are, however, several potential problems.
+Since this is a web service, documents have to be uploaded.
+Any confidential or personal data are, therefore, shared with the company.
+Additionally, we have no way of knowing "how things work" in the background.
+File support is pretty extensive, but there is a strict restriction on file size at 4 MB - scanned documents or simply larger files will be a problem.
+The website also provides valuable information, and the company seems to be open to questions from users - all completely for free.
+
+\section{PDF viewer Adobe Acrobat Reader DC}
+
+One of the most popular PDF viewers from the company Adobe - Acrobat Reader DC - can also verify and add signatures to PDF files.
+
+\obrazok{./figures/screenshot-adobeacrobat}{Screenshot of the Adobe Acrobat Reader DC}{screenshot-adobeacrobat}
+
+Using the QES is slightly harder to find in the Adobe Acrobat Reader.
+It is not available under "Signing", but "More Tools" where it is called "Certificates".
+Setup of the certificate for use with the SR eID can be a hit or miss.
+On one computer, the application would not include a certificate from the SR eID, and we had to go through a process of adding the PKCS \#11 library ourselves.
+That meant we had to go through several steps and windows and disable "protected mode".
+The Setup of this application should be done for the user ideally by the IT personnel.
+On another computer, we were simply able to choose from the list of the available certificates.
+That being said, everything is correctly localized, and being one of the most used applications of its kind, it is very easy to search for help in situations like this.
+Outside of that, verification of signatures is part of the application and is very responsive and flexible.
+The application is proprietary and can not be extended or verified.
+It is also available only on the Windows and Mac OS X.
+
+\section{Commercial desktop application D.PDF Signer}
+
+Application D.PDF Signer is a paid proprietary desktop application.
+
+\obrazok{./figures/screenshot-dpdfsigner}{Screenshot of the D.PDF Signer}{screenshot-dpdfsigner}
+
+Just as the zep.digis.sk, D.PDF signer utilizes the currently bundled software with the SR eID.
+It is straightforward and flexible.
+Downsides are lack of support for platforms other than Windows, localization, and built-in signature verification.
+
+\section{Overview}
+
+We have summarized our findings in Table \ref{table:software}.
+Value \textit{X} means best-case scenario, no value worst-case scenario, and value \textit{/} somewhere in between - partially or depends.
+
+\begin{table}[h!]
+  \begin{tabular}{ p{3cm}|p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm} }
+                      & \rot{licensing} & \rot{cross-platform} & \rot{localization} & \rot{file suppport} & \rot{easy setup} & \rot{flexibility} & \rot{built-in verification} & \rot{extensibility} \\
+    \hline
+    JSignPdf          & X               & /                   &                    &                      &                  & X                 &                             & / \\
+    zep.disig.sk      & /               & X                   & X                  & /                    & X                &                   & X                           &   \\
+    Acrobat DC        & /               & /                   & X                  &                      & /                & X                 & X                           &   \\
+    D.PDF Signer      &                 &                     & /                  &                      & X                & /                 &                             &   \\
+  \end{tabular}
+
+  \caption{Comparison of existing software}
+  \label{table:software}
+\end{table}
+
+\iffalse % Not part of the version for ŠVK
 \section{Modularity}
@@
 \subsection{Communicating and bundling with GPL software}
+\fi
 
@@
 
-TODO: Structure shoul roughly follow goals
+We have developed an open-source, cross-platform desktop software with focus on the modularity, architecture, usability, localization, automated testing with overall automation, and extensibility by the community.
+The source code is available at \url{https://github.com/durasj/octosign}, \url{https://github.com/durasj/octosign-dss}, and \url{https://github.com/durasj/octosign-image}.
 
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-Integer lacinia, nulla porta varius tempus, lacus metus blandit
-lorem, a rutrum justo wisi id sapien. Integer risus libero,
-feugiat eleifend, ornare ac, volutpat nec, sem. In facilisis,
-quam eu elementum aliquet, lorem quam euismod dui, aliquet
-laoreet purus ipsum ac quam.
+\subsection{Modularity}
+
+Although we focus on the use with the SR eID, we want the application to be extensible by the community and ready for experimentation with new ways of signing.
+That is why we have designed our application to be a fully modular platform.
+
+The main module is an Electron\footnote{Electron is a framework allowing the development of desktop software using web technologies.} application written in the TypeScript.
+This module features simple UI and abstraction around modules for signing that we call \pojem{backends}.
+It handles all of the communication with the user for the backend and tries to simplify development of the backends as much as possible.
+
+Signing modules - \pojem{backends} - are separate CLI applications written in any programming language.
+They are responsible for the actual document manipulation - signing and verification.
+
+The main module loads all available backends at runtime and verifies their availability, options, and supported file types.
+Once the user decides to verify or sign a file, they are executed as CLI applications.
+
+Communication between the main module and backends is in plaintext over the standard streams - standard input, output, and error.
+To make sure we have correctly documented and easily accessible way for backend development, we have prepared a Backend specification\footnote{Backend specification is available at \url{https://github.com/durasj/octosign/wiki/Backend-specification}.} that specifies what is required of the backend, what is optionally supported, and the format of the inputs, outputs, and error handling.
+
+\iffalse % Not part of the version for ŠVK
+TODO: Go a bit more over the backend specification here
+\fi
+
+\obrazok{./figures/modularity}{Modules of our application and their communication}{modularity}
+
+There are several advantages to this approach.
+Firstly, we can write backends in any language that suits us the most.
+For example, our backend that allows signing documents with the signatures compliant with the EU eIDAS directive is written in Java because it is based on a library from EU - DSS - written in Java.
+We expect most of the languages have APIs for the consumption of the standard streams with excellent portability across different desktop platforms.
+Secondly, since we are working on separate applications, they can be written by different people under a different license.
+That also means we can easily swap the main module or any of the backends for use in different environments or different bundles.
+Thirdly, together with the choice of using standard streams, it means that anyone can quickly start the development of a new backend.
+Writing to the standard output stream in CLI application is one of the first things developers usually learn - so-called "Hello World" applications.
+It usually should not involve studying any required third-party libraries, and authors can choose the language they are most familiar with.
+
+\subsection{Architecture of the main module} \label{architecture}
+
+Our main module is the most complex one.
+Electron applications have a main thread that is running a Node.js\footnote{Node.js is a JavaScript runtime for use outside of the browser with easier access to the OS APIs.} and a UI thread running the relevant part of the Chromium.
+Their communication is done via IPC and allows access to the Node.js APIs in the UI thread.
+Considering the JavaScript is executed at runtime, this, however, means we are exposing ourselves to potential attacks (mostly XSS\footnote{XSS means Cross-site scripting, see \url{https://owasp.org/www-community/attacks/xss/}.}).
+To reduce the attack surface, as Node.js allows easy access to many APIs like full access to the filesystem, we have fully separated the UI thread and main thread.
+
+The main thread is focused only on the communication with the backends and provides the only absolutely necessary functionality for the UI.
+They are connected by a very thin level of abstraction - a set of functions that are bound on the \textit{window} object during the initialization by the main thread and, when called from the UI thread, automatically communicate via IPC.
+All of the UI code and main thread code is, therefore, also fully portable and could also be used in a different setup - for example, in a web application or mobile application.
+We would be mostly required to only change the thin level of abstraction - implement the interface we have created on the \textit{window} object.
+Furthermore, it allows us to develop the UI in the browser with a mocked main thread and also allows for easy UI testing, which we go over in Section \ref{testing}.
+
+\subsection{Usability}
+
+Our main objective for the usability was to provide the end-users with an easy to use, straightforward UI, and to make as many decisions for them as possible.
+We have tried to achieve this by using as little controls as possible and by trying to have only one main control that we expect the user to interact with.
+
+\obrazok{./figures/screenshot-octosign}{Screenshot from the main screen of our application Octosign.}{screenshot-octosign}
+
+We hide all unnecessary information and choose a most likely default without asking users to choose from options that could confuse them, or would require them to study to make informed decisions.
+We assume this will mean the users will be more confident as there will be fewer steps where they were not sure if they have made the right choice.
+
+Since we are developing this application as a cross-platform desktop application available under a permissive open-source license, it should be usable by as many people as possible.
+
+\subsection{Localization}
+
+To make our application accessible to as many users as possible, we have also made the application translatable and try to display data like dates in the local format.
+
+When the application is started, we try to detect a locale of the user and set it as a default language.
+If we do not yet support the locale of the user, we default to the English language.
+At the time of writing, our application is fully translated in three languages: English (source language), Slovak, and Czech.
+
+The process of content translation is mostly automated.
+We start with the automated extraction of the source strings from the code of the main module and files provided by the backends.
+Next, we manually include the extracted source strings in the commit to our versioning control - we have retained manual control over this to prevent a noise that automated commits could make.
+On push to the hosted repository, a localization platform Transifex\footnote{Transifex is a SaaS localization platform free for OSS, see \url{https://www.transifex.com/}.} automatically pulls and process the updates.
+These phrases can then be translated by anybody using the UI of the Transifex platform to make it easily accessible.
+Once the changes are made on the Transifex platform, it automatically creates a pull request on our hosted repository with the changes made to that particular language.
+As we explain in detail in Section \ref{testing}, part of the build process is to incorporate these in the application.
+
+\subsection{Testing and automation} \label{testing}
+
+When approaching automated testing, we have decided to follow the opinion of the engineers at Google\footnote{Approach outlined in the article \textit{Just Say No to More End-to-End Tests} available at \url{https://testing.googleblog.com/2015/04/just-say-no-to-more-end-to-end-tests.html}.}.
+We write three types of tests: end-to-end, integration, and unit tests.
+The ratio of these types of tests should form a pyramid where the heavy bottom is formed by the unit tests, followed by a smaller set of component tests forming the middle of the pyramid, and a small set of end-to-end tests forming the tip of the pyramid.
+
+In our case, we have decided to write unit tests in the Jest\footnote{Jest is a testing framework and runner, see \url{https://jestjs.io/}.} testing both the UI and main thread code, component tests in Jest for the main thread and Cypress\footnote{Cypress is a testing framework and runner using the Chromium, see \url{https://www.cypress.io/}.} for the UI thread combined with the Percy\footnote{Percy is a SaaS for visual testing, see \url{https://percy.io/}.}, and end-to-end tests as Spectron tests which run and interact with the actual distributables (executable files).
+We are able to fully isolate each type of tests because of our architecture that allows easy mocking.
+We explain the architecture in Section \ref{architecture}.
+
+Our code is kept in a git repository hosted at Github and is a source for the Continuous Integration (CI) pipeline that, on each push, runs static analysis on the code (), mentioned automated tests, builds the bundle and uploads created distributables for all platforms back to the Github.
+Publishing of the new release is then done by saving the release in the Github.
 
@@
 
-TODO: Reference table from analysis, add new column
+\begin{table}[h!]
+  \begin{tabular}{ p{3cm}|p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm} }
+                      & \rot{licensing} & \rot{cross-platform} & \rot{localization} & \rot{file suppport} & \rot{easy setup} & \rot{flexibility} & \rot{built-in verification} & \rot{extensibility} \\
+    \hline
+    JSignPdf          & X               & /                   &                    &                      &                  & X                 &                             & / \\
+    zep.disig.sk      & /               & X                   & X                  & /                    & X                &                   & X                           &   \\
+    Acrobat DC        & /               & /                   & X                  &                      & /                & X                 & X                           &   \\
+    D.PDF Signer      &                 &                     & /                  &                      & X                & /                 &                             &   \\
+    \hline
+    Our - Octosign    & X               & X                   & X                  & X                    & /                &                   & X                           & X \\
+  \end{tabular}
+
+  \caption{Comparison of existing software with our software.}
+  \label{table:software-our}
+\end{table}
+
+Using our review from Chapter \ref{review}, we have compared our application in Table \ref{table:software}.
+Value \textit{X} means best-case scenario, no value worst-case scenario, and value \textit{/} somewhere in between - partially or depends.
+
+Our software is available under a permissive license (see Section \ref{copyright} for why this matters), supports all major desktop platforms, available in several locales, with support for practically all file types and sizes, with built-in verification and easy extensibility.
+
+In the future, we can improve our application to be easier to setup for more users, and be more flexibile when it comes to the source of the private key and certificate, or timestamping.
+That being said, we need to consider each configurable functionality we add as it increases complexity which could mean our software is harder to understand.
 
@@
 
-TODO: Structure shoul roughly follow goals
+Goal of the website is to both support the application in the form of Download page and Help page and provide a to provide a general information on electronic signatures.
+The website is available at \url{https:\\octosign.com} and source code at \url{https://github.com/durasj/octosign-website}.
+The website is, at the time of writing, available in English and Slovak language.
 
-Donec dolor arcu, posuere at, vehicula vitae, accumsan ut,
-lacus. Nulla tristique eros eu diam. Vivamus nec tortor vel
-ligula elementum lacinia. Curabitur euismod eros adipiscing
-ipsum. Donec sed quam at felis suscipit egestas. Morbi faucibus
-libero sit amet libero. Nullam laoreet ipsum eu eros. Donec in
-diam. Ut facilisis eros vel leo. Nunc vitae mauris. Donec leo
-erat, luctus porttitor, laoreet eget, facilisis non, erat.
-Integer nec elit.
+\obrazok{./figures/website-home}{Home page of the octosign.com website.}{website-home}
+
+\subsection{Downloads}
+
+Downloads are available for each major desktop platform in a format suitable for installation or integration in the operating system.
+For Windows, we are distributing an \textit{.exe} installer working on all supported 64-bit Windows versions.
+For Linux, a portable \pojem{AppImage} and a \textit{.deb} package for Debian-like distributions.
+For Mac OS X, a \textit{.dmg} disk image.
+
+All of the distributables originate from the Github as assets of a release, where they are uploaded by the CI (see Section \ref{testing}).
+Each release is tied to a tag on the master branch, where each commit has to be signed using a recognized PGP key.
+In turn, we have some level of confidence source code used for the distributables comes from one of the authors.
+
+\subsection{Help and resources}
+
+Our website provides a simple \textbf{help page} with different use cases that should allow users to start using the application quickly.
+It also contains a brief explanation of more advanced features of the application like settings.
+
+Another issue we try to solve with the website is a lack of understanding of electronic signatures.
+Some users do not know if they are a viable alternative, why is that the case, and how they should be used properly.
+We also plan to provide more detailed information for people who are curious about the specifics of the signatures used in the eIDAS scheme.
+
+These pages are written in a Markdown\footnote{Markdown is a lightweight markup language, see \url{https://commonmark.org/}.}, making it more accessible for editing by the community that does not need to understand specific technologies used to display the content on the website.
 
@@
 
-Ut lobortis semper risus, non condimentum dui convallis ut.
-Nulla eget volutpat tellus.
-Vestibulum lobortis tincidunt massa eu rhoncus.
-Suspendisse luctus eu dui non vehicula.
-Vivamus elementum auctor felis, placerat maximus magna lobortis ut.
-Donec placerat sem a mi sagittis blandit.
-Maecenas pellentesque laoreet mauris, dictum viverra ante sodales sed.
-Nullam non ligula quis ante ultricies finibus non quis ex.
-Ut tempor vitae ipsum sed imperdiet.
-Fusce aliquam nisl sit amet nunc tempor convallis.
-Vivamus vehicula magna sit amet purus commodo, et tincidunt purus accumsan.
-Nulla velit dolor, lacinia nec scelerisque a, euismod a sapien.
-%
+Will be added later once the content is reviewed and it has a green light and we discuss what are the main points I should focus on here.
+
+Our goal is to explore an accessible way to use electronic signatures.
+In the beginning, we go over the preconditions that make it a viable alternative to the handwritten signatures.
+We cover the legal preconditions relevant for the European Union and Slovakia in particular.
+Also, we explain technical preconditions focusing on the cryptographic concepts like asymmetric cryptography, timestamping, hashing, and standards that comply with the law within the European Union.
+
+Secondly, we do a review of the existing software.
+We go quickly over several applications that are meant for basic use.
+We compare them using several criteria to find out the strengths and weaknesses of the current implementations and identify a potential space for improvements.
+
+Finally, we devote our attention to our own implementation.
+We explain the proposed desktop software and go into details on software engineering aspects.
+Specifically, modularity, architecture, usability, localization, testing, and automation.
+In the end, we compare it with the existing software and identify possible future improvements.
+
+Another result we focus on is a supporting website and how it helps with the use of the software.
+All in a way open for extensibility by the community.
 
@@
 
+  \bibitem{commercialcode}
+  \emph{Act No. 513/1991 Coll. Commercial code}
+
   \bibitem{notarylaw}
@@
   %  ------
+  \bibitem{law}
+  \osoba{MASON, Stephen}, 2017.
+\emph{Electronic Signatures in Law: Fourth Edition}.
+London: University of London.
+ISBN 978-1-911507-01-7.
+Available at: \url{https://humanities-digital-library.org/index.php/hdl/catalog/view/electronicsignatures/1/86-1}
+
   \bibitem{licensing}
@@
 Available at: \url{https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1002598}
+
+  \bibitem{osslicensing}
+  \osoba{LAURENT, Andrew}, 2008. \emph{Understanding Open Source and Free Software Licensing}. Sebastopol: O’Reilly Media. ISBN 978-0596005818.
 
@@
 
-  \bibitem{law}
-  \osoba{MASON, Stephen}, 2017.
-\emph{Electronic Signatures in Law: Fourth Edition}.
-London: University of London.
-ISBN 978-1-911507-01-7.
-Available at: \url{https://humanities-digital-library.org/index.php/hdl/catalog/view/electronicsignatures/1/86-1}
+  \bibitem{pades}
+  ETSI TS 102 778-1 V1.1.1 2007-07.
+\emph{PAdES Overview - a framework document for PAdES}.
+Sophia Antipolis: European Telecommunications Standards Institute.
+Available at: \url{https://www.etsi.org/deliver/etsi_ts/102700_102799/10277801/01.01.01_60/ts_10277801v010101p.pdf}
+
+  \bibitem{xades}
+  ETSI TS 101 903 V1.4.2 2010-12.
+\emph{XML Advanced Electronic Signatures (XAdES)}.
+Sophia Antipolis: European Telecommunications Standards Institute.
+Available at: \url{https://www.etsi.org/deliver/etsi_ts/101900_101999/101903/01.04.02_60/ts_101903v010402p.pdf}
+
+  \bibitem{cades}
+  ETSI TS 101 733 V2.2.1 2013-04.
+\emph{CMS Advanced Electronic Signatures (CAdES)}.
+Sophia Antipolis: European Telecommunications Standards Institute.
+Available at: \url{https://www.etsi.org/deliver/etsi_ts/101700_101799/101733/02.02.01_60/ts_101733v020201p.pdf}
 
@@
   \bibitem{oss}
-  \osoba{FOGEL, Karl}, 2019. \emph{Producing Open Source Software} [online]. TODO: TODO. [cit. 2020-04-21]. Available at: \url{https://producingoss.com/en/index.html}
-
-  \bibitem{osslicensing}
-  \osoba{LAURENT, Andrew}, 2008. \emph{Understanding Open Source and Free Software Licensing}. Sebastopol: O’Reilly Media. ISBN 978-0596005818.
+  \osoba{FOGEL, Karl}, 2019. \emph{Producing Open Source Software} [online]. [cit. 2020-04-21]. Available at: \url{https://producingoss.com/en/index.html}
 

### 2020-05-07 paper/main.tex
Add fixes for paper feedback

@@
 \usepackage[hyphens]{url} % format and linebreak of URLs
+
+\usepackage{amssymb}
+\newcommand\doublecheck{\checkmark\kern-0.6em\checkmark}
 
@@
 \abstrakt{
-So zmenami v právnom statuse elektronických podpisov vo svete sa vytvára príležitosť rozšíriť elektronické podpisovanie ako náhradu pre vlastnoručné podpisy mimo obvyklých oblastí.
-Elektronické podpisy sú dôležité čoraz viac v dobe elektronickej komunikácie kedy sú vlastnoručné podpisy nepraktické z dôvodu nutnosti tlače, osobného kontaktu alebo zasielania podpísaných dokumentov.
+So zmenami v právnom statuse elektronických podpisov vo svete sa vytvára príležitosť rozšíriť elektronické podpisovanie ako plnohodnotnú náhradu pre vlastnoručné podpisy.
+Elektronické podpisy sú dôležité čoraz viac v dobe elektronickej komunikácie, kedy sú vlastnoručné podpisy nepraktické z dôvodu nutnosti tlače, osobného kontaktu alebo zasielania podpísaných dokumentov.
 V práci sa venujeme právnym predpokladom relevantným pre EÚ, technologickým predpokladom, prehľadu v súčasnosti dostupného softvéru a návrhu vlastnej softvérovej platformy.
@@
 Aplikácia nie je viazaná iba na slovenské eID, ale je pripravená na ďalšie použitie s rôznymi zariadeniami v rôznych štátoch a ako platforma pre iné druhy podpisov.
-Súčasťou je aj webová stránka, ktorá okrem možnosti stiahnutia softvéru a stručnej dokumentácie poskytuje všeobecné informácie o elektronických podpisoch aby bol projekt čo najviac prístupný verejnosti.
+Súčasťou je aj webová stránka, ktorá okrem možnosti stiahnutia softvéru a stručnej dokumentácie poskytuje všeobecné informácie o elektronických podpisoch, aby bol projekt čo najviac prístupný verejnosti.
 Dôraz kladieme na jednoduchú rozšíriteľnosť programu vo forme jazykových prekladov, dokumentácie, funkcionality a všeobecných informácií.
@@
 
-While, specifically in the European Union, they seem to be relatively extensively used in communication with the government, they do not seem to be widely adopted in an exchange between individuals or smaller organizations.
-The problem is handwritten signatures are often unpractical for electronic communication due to the need to print, personally exchange, or send signed documents.
+While, specifically in the European Union, electronic signatures seem to be relatively extensively used in communication with the government, they do not seem to be widely adopted in the exchange between individuals or smaller organizations.
+The problem is that handwritten signatures are often unpractical for electronic communication due to the need of physical presence or printing and sending of documents.
 
@@
 Also, we explain technical preconditions focusing on the cryptographic concepts like asymmetric cryptography, timestamping, hashing, and standards that comply with the law within the European Union.
-Secondly, we do a review of the existing software.
-We go quickly over several applications that are meant for basic use.
-We compare them using several criteria to find out the strengths and weaknesses of the current implementations and identify a potential space for improvements.
+Secondly, we do a review of the existing software, where we briefly cover several applications that are meant for basic use.
+We compare them using several criteria to find out the strengths and weaknesses of the current implementations and identify a potential gap.
 Finally, we devote our attention to our own implementation.
@@
 Another result we focus on is a supporting website and how it helps with the use of the software.
-All in a way open for extensibility by the community.
+All in a way that is open for extensibility by the community.
 
@@
 Firstly, the current state of the law and recent changes relating to the use of electronic signatures, secondly, well-established concepts in cryptography like \pojem{asymmetric cryptography} or \pojem{hashing}, and lastly, expertise from the field of software engineering - design, implementation, and maintenance of computer software.
-We will concetrate on the first two in relation to the electronic signatures, and, where applicable, open-source desktop computer software.
+We will concentrate on the first two in relation to the electronic signatures and where applicable open-source desktop computer software.
 
@@
 In general, electronic signature can be represented in different ways (e.g.
-as an image, text or other data attached to the document) and they \textbf{can not be denied in legal proceedings} just because they are in the electronic form \cite{eidas}.
+as an image, text or other data attached to the document) and they can not be denied in legal proceedings just because they are in the electronic form \cite{eidas}.
 
-\pojem{Advanced electronic signatures} are a subset of electronic signatures that have to \textbf{uniquely link and identify} the signature author, be created in a way that is \textbf{possible only by them} and any \textbf{changes to the signed document have to be detectable} \cite{eidas}.
+\pojem{Advanced electronic signatures} are a subset of electronic signatures that have to uniquely link and identify the signature author, be created in a way that is possible only by them and any changes to the signed document have to be detectable \cite{eidas}.
 From the technical point of view, standards \pojem{PAdES}, \pojem{XAdES}, and \pojem{CAdES} specified by the European Telecommunications Standards Institute (ETSI) comply with these requirements.
@@
 \pojem{Qualified electronic signatures} (QES) are a subset of \pojem{advanced electronic signatures} with more specific requirements that can be found in annex 1 of the eIDAS regulation\footnote{Available at https://eur-lex.europa.eu/legal-content/EN/TXT/HTML/?uri=CELEX:32014R0910\#d1e32-111-1.}.
-In more practical terms, such \pojem{advanced electronic signatures} are created with a qualified device using a qualified certificate and a qualified trust service.
+In more practical terms, such advanced electronic signatures are created with a qualified device using a qualified certificate and a qualified trust service.
 Qualified, in this case, means authorized for such use by the legal authorities.
-These devices, \pojem{certificates}, and trust services are explicitly listed by the EU. In many cases indirectly by listing the organization (and its items) from a specific country.
+These devices, certificates, and trust services are explicitly listed by the EU. In many cases indirectly by listing the organization (and its items) from a specific country.
 
@@
 
-Mentioned terminology is used when considering a natural person. Legal entities are able to use \pojem{electronic seals}, which are, from the technical point of view and for our purposes, identical.
+Mentioned terminology is used when considering a natural person.
+Legal entities are able to use \pojem{electronic seals}, which are, from the technical point of view and for our purposes, identical to electronic signatures.
 
@@
 Such right can be exercised to license the software under either \pojem{proprietary} or \pojem{open-source} license.
-\pojem{Proprietary} meaning under the exclusive legal right of the author, typically also \textbf{confidential and distributed as a paid product}.
-\pojem{Open-source} meaning having its source code \textbf{freely available}, typically also \textbf{distributed for free} and without any liability (the software is provided "as is").
+\iffalse % Not part of the version for ŠVK
+\pojem{Proprietary} meaning under the exclusive legal right of the author, typically also confidential and distributed as a paid product.
+\pojem{Open-source} meaning having its source code freely available, typically also distributed for free and without any liability (the software is provided "as is").
 
@@
 This means that consumers of the software can quickly recognize their rights and responsibilities if they decide to use, modify, or distribute the software.
+\fi
 In general, we can classify the \pojem{open-source} licenses into two categories: \pojem{permissive} and \pojem{copyleft}.
@@
 As a result, it forces them to extend the rights they have received onto others and, in turn, somehow hinder usability in software licensed under a different license.
+\iffalse % Not part of the version for ŠVK
 A popular example of such license is the GNU General Public License(GPL) and its derivatives like GNU Lesser GPL (LGPL) or GNU Affero GPL (AGPL).
+\fi
 
@@
 They typically allow commercial use, modifications, distribution or sublicensing as long as the author is not held liable, and the original copyright and license are distributed with the software.
+\iffalse % Not part of the version for ŠVK
 Often used permissive licenses are MIT License, Apache License, or BSD License.
+\fi
 
@@
 
+\iffalse % Not part of the version for ŠVK
 As we have already mentioned, software licenses can influence whether and how it can be used as a part of different software.
 Both when we are the ones using someone else's software and when someone else is using our software.
+\fi
 Generally speaking, the more open the license is, the higher the probability it will not prevent adoption for legal reasons.
+We provide examples of categorized licenses in Figure \ref{fig:license-comparison}.
+
+\iffalse % Not part of the version for ŠVK
 Various possible combinations can be seen in Figure \ref{fig:license-compatibility}, where greyed out options mean such combination is not possible unless a separate licensing agreement is reached with the copyright owner.
@@
 \obrazok{./figures/license-compatibility}{Schematic representation of license directionality. \osoba{Morin 2012} \cite{licensing}}{license-compatibility}
+\fi
 
@@
 
-\textbf{Secure distribution of keys} seems to be feasible when we are adding a new user to our network.
+Secure distribution of keys seems to be feasible when we are adding a new user to our network.
 We require each user to personally come and pick up a physical device that contains pre-generated keys.
@@
 
-Even if we think we have established a reasonably secure way to do that and we are willing to wait, we realize the \textbf{number of keys} is quickly increasing with each new member.
+Even if we think we have established a reasonably secure way to do that and we are willing to wait, we realize the number of keys is quickly increasing with each new member.
 With \textit{n} users, the number of keys is
@@
 
-\textbf{Either party can also lie} about not signing the document.
+Either party can also lie about not signing the document.
 Since both have the key that can be used for encryption, they can sign any document on behalf of the other party, and there is no way to disprove it if we look at it as a third party.
@@
 These issues can be addressed with \pojem{asymmetric cryptography}.
-In general, we assign \textbf{a pair of keys} to each person, one of them used for encryption, known as a \pojem{public key}, another used for decryption, known as a \pojem{private key}.
+In general, we assign a pair of keys to each person, one of them used for encryption, known as a \pojem{public key}, another used for decryption, known as a \pojem{private key}.
 One party can encrypt, and the other is the only one that can decrypt.
@@
 The signer is the only one who possesses the \pojem{private key} used for signing, and others can use the \pojem{public key} for signature verification.
-We \textbf{sign} the document, including the other parties' signature.
+We sign the document, including the other parties' signature.
 
@@
 
-Practical, widely used implementation of this idea is the \pojem{RSA} signature scheme that relies on the integer factorization problem (difficulty of factoring a product of two large primes) published in 1978 \cite{rsa}.
+Practical, widely used implementation of this idea is the \pojem{RSA} signature scheme that relies on the integer factorization problem published in 1978 \cite{rsa}.
 
@@
 
-We are left with the \textbf{problem of public key distribution} - even though \pojem{public-key schemes} do not require a \textit{secure channel}, they require \textit{authenticated channels} for the distribution of the public keys \cite{cryptotxtbook}.
+We are left with the problem of public key distribution - even though \pojem{public-key schemes} do not require a \textit{secure channel}, they require \textit{authenticated channels} for the distribution of the public keys \cite{cryptotxtbook}.
 
@@
 In the case of the Slovak ID cards, it is Disig - SVK eID Accredited CA, issued by the National Security Authority\footnote{Translation from "Národný bezpečnostný úrad".}.
-Such \pojem{certificates}, therefore, \textbf{reliably uniquely link and identify} the \pojem{public key} owners, as can be seen in Figure \ref{fig:certificate-scheme}\footnote{Created by an unknown author, licensed under the CC BY-SA 3.0 license.}.
+Such \pojem{certificates}, therefore, reliably uniquely link and identify the \pojem{public key} owners, as can be seen in Figure \ref{fig:certificate-scheme}\footnote{Created by an unknown author, licensed under the CC BY-SA 3.0 license.}.
+Ability to do that is the first requirement for the advanced electronic signatures mentioned in Section \ref{eidas}.
 Commonly used standard for public key certificates is X.509\footnote{Specification available under RFC5280 at \url{https://tools.ietf.org/html/rfc5280}.}.
@@
 
-It is often useful to map data of arbitrary size to a fixed-length (typically short) set of bits.
-This process is commonly called \pojem{hashing}, a function that can be used for it \pojem{hash function}, and output of such function \pojem{hash}.
+It is often useful to map data of arbitrary size to a fixed-length and typically short set of bits called \pojem{hash}.
+Function that can be used for it is called \pojem{hash function}.
 
-In cryptography, we are considering only \pojem{hash functions} that can also satisfy these additional \textbf{requirements}:
+In cryptography, we are considering only \pojem{hash functions} that can also satisfy these additional requirements:
 
@@
   \item \textbf{Deterministic} - same input data are always mapped to the same \pojem{hash}.
-  \item \textbf{One-way} - it should be practically impossible to retrieve input data for a given \pojem{hash}.
-  \item \textbf{Resistant to collisions} - it should be practically impossible to find two inputs with the same \pojem{hash}.
-  \item Even small change to the input should lead to significantly different \pojem{hash} - so-called \textbf{avalanche effect}.
+  \item \textbf{One-way} - it should be infeasible to retrieve input data for a given \pojem{hash}.
+  \item \textbf{Resistant to collisions} - it should be infeasible to find two inputs with the same \pojem{hash}.
+  \item \textbf{Avalanche effect} - even small change to the input should lead to significantly different \pojem{hash}.
 \end{enumerate}
 
-In the context of \pojem{digital documents}, we can use such \pojem{hash function} to generate a \pojem{hash} for the document.
-This \pojem{hash} can represent a document of any size and is used during the creation of a \pojem{digital signature} instead of the document itself.
-Any further changes to the document will, therefore, lead to a significantly different \pojem{hash}, which means the \pojem{digital signature} will no longer be valid, and these \textbf{changes are detected}.
-For this purpose, only \pojem{hash functions} that are reasonably fast can be used, so that we can utilize them on documents of variable size.
+In the context of \pojem{digital documents}, we can use such hash function to generate a hash for the document.
+This hash can represent a document of any size and is used during the creation of a digital signature instead of the document itself.
+Any further changes to the document will, therefore, lead to a significantly different hash, which means the digital signature will no longer be valid, and these changes are detected.
+This ability is one of the requirements for the advanced electronic signatures from Section \ref{eidas}.
+For this purpose, only hash functions that are reasonably fast can be used, so that we can utilize them on documents of variable size.
 
@@
 
-An example of commonly used \pojem{hash function} family is \pojem{Secure Hash Algorithms} (SHA), with \pojem{hash functions} relevant to this thesis being mainly SHA-2\footnote{SHA-2 specification available in FIPS PUB 180-4 at \url{https://nvlpubs.nist.gov/nistpubs/FIPS/NIST.FIPS.180-4.pdf}.}.
+An example of commonly used hash function family is \pojem{Secure Hash Algorithms} (SHA), with hash functions relevant to this thesis being mainly SHA-2\footnote{SHA-2 specification available in FIPS PUB 180-4 at \url{https://nvlpubs.nist.gov/nistpubs/FIPS/NIST.FIPS.180-4.pdf}.}.
 
@@
 A common part of the paper signed documents is the date when they were signed.
-Their digital counterpart is PKI-based \pojem{trusted timestamping} - a process of \pojem{digitally signing} date and time of document creation or modification.
-No one should be able to tamper with such \pojem{timestamp} - not even the author of the document.
+Their digital counterpart is PKI-based \pojem{trusted timestamping} - a process of digitally signing date and time of document creation or modification.
+No one should be able to tamper with such timestamp - not even the author of the document.
 Additionally, we should not be able to choose arbitrary date and time ourselves, but it should represent the actual time at the signing.
@@
 This process can be seen in Figure \ref{fig:trusted-timestamping}\footnote{Schema created by Bart Van den Bosch, licensed under the CC BY-SA 2.0 BE license.}.
-Verifying can be done by decrypting this information using the \pojem{public key} and comparing the \pojem{hashes}. Any attempt at \textbf{tampering is therefore detected}.
+Verifying can be done by decrypting this information using the public key and comparing the hashes.
+Any attempt at tampering is therefore detected, which helps to cover requirements for the advanced electronic signatures from Section \ref{eidas}.
 
@@
 
-Extensions for PDF files, XML files, and CMS signed data for signing using \pojem{advanced electronic signatures} (as explained in Section \ref{eidas}), are \pojem{PAdES}, \pojem{XAdES}, and \pojem{CAdES}, respectively.
+Extensions for PDF files, XML files, and CMS signed data for signing using advanced electronic signatures, as explained in Section \ref{eidas}, are \pojem{PAdES}, \pojem{XAdES}, and \pojem{CAdES}, respectively.
 They are defined in the respective specifications by the ETSI in \cite{pades}, \cite{xades}, and \cite{cades}.
@@
 
-The ability to keep signed documents valid for a prolonged period of time is covered in these standards under optional \pojem{Long-Term Validation} (LTV) that allows \textbf{archiving} of documents for many years, possibly decades.
+The ability to keep signed documents valid for a prolonged period of time is covered in these standards under optional \pojem{Long-Term Validation} (LTV) that allows archiving of documents for many years, possibly decades.
 If enabled, a trustworthy timestamp is required as validity is verified for the time the signature (and therefore timestamp) was created.
@@
   \item \textbf{Licensing} - is it freely available - both in terms of money and source code - best case is open-source; worst case is paid proprietary software.
-  \item \textbf{Cross-platform availability} - is it available for users of all major desktop platforms - best case is availability on Windows, Mac OS X, and Linux; worst case is availability only on one platform.
+  \item \textbf{Cross-platform compatibility} - is it available for users of all major desktop platforms - best case is availability on Windows, Mac OS X, and Linux; worst case is availability only on one platform.
   \item \textbf{Localization} - is it translated and does it display data like dates in local format - best case is availability at least in English and Slovak language; worst case is availability only in one language with no support for localization of dates, etc..
@@
 The application consists only of one window, which makes it very simple at first sight.
-The very first input - "Keystore type" - can be harder to understand for the basic user since it contains over a dozen options in the form of abbreviations like "CASEEXACTJKS" or "PKCS12-DEF-3DES-40RC2".
-The default value is "WINDOWS-MY", however, which is what is exactly required to use the SR eID.
-Application has a toggle for the "Advanced view" that is off by default, but it is necessary to turn it on if we want to choose the correct key.
+The very first input - \textit{Keystore type} - can be harder to understand for the basic user since it contains over a dozen options in the form of abbreviations like \textit{CASEEXACTJKS} or \textit{PKCS12-DEF-3DES-40RC2}.
+The default value is \textit{WINDOWS-MY}, however, which is what is exactly required for the SR eID.
+Application has a toggle for the \textit{Advanced view}, which is off by default, but it is necessary to turn it on if we want to choose the correct key.
 On some of the computers, we were not able to see the key from the eID. On others, it took a while and several pressing of a button to load them.
-The correct key was not the first, preselected key in the form of "Name Surname", but the second "Name Surname (1)".
+The correct key was not the first, preselected key in the form of \textit{Name Surname}, but the second \textit{Name Surname (1)}.
 With everything set up correctly, the application was successfully able to attach a QES.
 Even though the application looks pretty simple, it is still pretty flexible with its options regarding the used algorithms, TSA, OCSP, or CRL.
-That being said, it is available only in English, and we have also noticed issues with the diacritics that you can see in Figure \ref{fig:screenshot-jsignpdf} - "Jakub Ďuraš" is displayed as "Jakub Ïura".
+That being said, it is available only in English, and we have also noticed issues with the diacritics as you can see in Figure \ref{fig:screenshot-jsignpdf} - "Jakub Ďuraš" is displayed as "Jakub Ïura".
 
@@
 
-\obrazok{./figures/screenshot-adobeacrobat}{Screenshot of the Adobe Acrobat Reader DC}{screenshot-adobeacrobat}
+\obrazok{./figures/screenshot-adobeacrobat}{Screenshot of the Adobe Acrobat Reader DC.}{screenshot-adobeacrobat}
 
 Using the QES is slightly harder to find in the Adobe Acrobat Reader.
-It is not available under "Signing", but "More Tools" where it is called "Certificates".
+It is not available under \textit{Signing}, but \textit{More Tools} where it is called \textit{Certificates}.
 Setup of the certificate for use with the SR eID can be a hit or miss.
 On one computer, the application would not include a certificate from the SR eID, and we had to go through a process of adding the PKCS \#11 library ourselves.
-That meant we had to go through several steps and windows and disable "protected mode".
-The Setup of this application should be done for the user ideally by the IT personnel.
+That meant we had to go through several steps and windows and disable the \textit{protected mode} option.
+The setup of this application should be done for the user ideally by the IT personnel.
 On another computer, we were simply able to choose from the list of the available certificates.
@@
 The application is proprietary and can not be extended or verified.
-It is also available only on the Windows and Mac OS X.
+It is available only on the Windows and Mac OS X.
 
@@
 We have summarized our findings in Table \ref{table:software}.
-Value \textit{X} means best-case scenario, no value worst-case scenario, and value \textit{/} somewhere in between - partially or depends.
+Value \doublecheck means best-case scenario, no value worst-case scenario, and value \checkmark somewhere in between - partially or depends.
 
 \begin{table}[h!]
-  \begin{tabular}{ p{3cm}|p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm} }
+  \begin{tabular}{ p{2.75cm}|p{0.85cm}p{0.85cm}p{0.85cm}p{0.85cm}p{0.85cm}p{0.85cm}p{0.85cm}p{0.85cm} }
                       & \rot{licensing} & \rot{cross-platform} & \rot{localization} & \rot{file suppport} & \rot{easy setup} & \rot{flexibility} & \rot{built-in verification} & \rot{extensibility} \\
     \hline
-    JSignPdf          & X               & /                   &                    &                      &                  & X                 &                             & / \\
-    zep.disig.sk      & /               & X                   & X                  & /                    & X                &                   & X                           &   \\
-    Acrobat DC        & /               & /                   & X                  &                      & /                & X                 & X                           &   \\
-    D.PDF Signer      &                 &                     & /                  &                      & X                & /                 &                             &   \\
+    JSignPdf          & \doublecheck    & \checkmark           &                    &                     &                  & \doublecheck      &                             & \checkmark \\
+    zep.disig.sk      & \checkmark      & \doublecheck         & \doublecheck       & \checkmark          & \doublecheck     &                   & \doublecheck                &            \\
+    Acrobat DC        & \checkmark      & \checkmark           & \doublecheck       &                     & \checkmark       & \doublecheck      & \doublecheck                &            \\
+    D.PDF Signer      &                 &                      & \checkmark         &                     & \doublecheck     & \checkmark        &                             &            \\
   \end{tabular}
@@
 
+\subsection{Usability and localization}
+
+Our main objective for the usability was to provide the end-users with an easy to use, straightforward UI, and to make as many decisions for them as possible.
+We have tried to achieve this by using as little controls as possible and by trying to have only one main control that we expect the user to interact with.
+
+\obrazok{./figures/screenshot-octosign}{Screenshot from the main screen of our application Octosign.}{screenshot-octosign}
+
+We hide all unnecessary information and choose a most likely default without asking users to choose from options that could confuse them, or would require them to study to make informed decisions.
+We assume this will mean the users will be more confident as there will be fewer steps where they were not sure if they have made the right choice.
+
+Since we are developing this application as a cross-platform desktop application available under a permissive open-source license, it should be usable by as many people as possible.
+
+We have also made the application easily translatable and try to display data like dates in the local format.
+When the application is started, we try to detect the locale of the user and set it as a default language.
+If we do not yet support the locale of the user, we default to the English language.
+At the time of writing, our application is fully translated in three languages: English (source language), Slovak, and Czech.
+
+The process of content translation is mostly automated.
+We start with the automated extraction of the source strings from the code of the main module and files provided by the backends.
+Next, we manually include the extracted source strings in the commit to our versioning control - we have retained manual control over this to prevent a noise that automated commits could make.
+On push to the hosted repository, a localization platform Transifex\footnote{Transifex is a SaaS localization platform free for OSS, see \url{https://www.transifex.com/}.} automatically pulls and process the updates.
+These phrases can then be translated by anybody using the UI of the Transifex platform to make it easily accessible.
+Once the changes are made on the Transifex platform, it automatically creates a pull request on our hosted repository with the changes made to that particular language.
+As we explain in detail in Section \ref{testing}, part of the build process is to incorporate these in the application.
+
 \subsection{Modularity}
@@
 
-The main module is an Electron\footnote{Electron is a framework allowing the development of desktop software using web technologies.} application written in the TypeScript.
+The main module is an Electron\footnote{Electron is a framework allowing the development of desktop software using web technologies.} application written in the language TypeScript\footnote{TypeScript is statically typed superset of JavaScript that compiles to JavaScript.}.
 This module features simple UI and abstraction around modules for signing that we call \pojem{backends}.
@@
 
-Signing modules - \pojem{backends} - are separate CLI applications written in any programming language.
+Signing modules - backends - are separate CLI applications written in any programming language.
 They are responsible for the actual document manipulation - signing and verification.
@@
 Firstly, we can write backends in any language that suits us the most.
-For example, our backend that allows signing documents with the signatures compliant with the EU eIDAS directive is written in Java because it is based on a library from EU - DSS - written in Java.
+For example, our backend that allows signing documents with the signatures compliant with the EU eIDAS directive is written in Java because it is based on a library from EU - DSS - also written in Java.
 We expect most of the languages have APIs for the consumption of the standard streams with excellent portability across different desktop platforms.
@@
 Our main module is the most complex one.
-Electron applications have a main thread that is running a Node.js\footnote{Node.js is a JavaScript runtime for use outside of the browser with easier access to the OS APIs.} and a UI thread running the relevant part of the Chromium.
+Electron applications have a main thread that is running a Node.js\footnote{Node.js is a JavaScript runtime for use outside of the browser with easier access to the OS APIs.} and a UI thread running the relevant part of the Chromium web browser.
 Their communication is done via IPC and allows access to the Node.js APIs in the UI thread.
-Considering the JavaScript is executed at runtime, this, however, means we are exposing ourselves to potential attacks (mostly XSS\footnote{XSS means Cross-site scripting, see \url{https://owasp.org/www-community/attacks/xss/}.}).
+Considering the JavaScript is interpreted at runtime, we are exposing ourselves to potential attacks (mostly XSS\footnote{XSS means Cross-site scripting, see \url{https://owasp.org/www-community/attacks/xss/}.}).
 To reduce the attack surface, as Node.js allows easy access to many APIs like full access to the filesystem, we have fully separated the UI thread and main thread.
@@
 Furthermore, it allows us to develop the UI in the browser with a mocked main thread and also allows for easy UI testing, which we go over in Section \ref{testing}.
-
-\subsection{Usability}
-
-Our main objective for the usability was to provide the end-users with an easy to use, straightforward UI, and to make as many decisions for them as possible.
-We have tried to achieve this by using as little controls as possible and by trying to have only one main control that we expect the user to interact with.
-
-\obrazok{./figures/screenshot-octosign}{Screenshot from the main screen of our application Octosign.}{screenshot-octosign}
-
-We hide all unnecessary information and choose a most likely default without asking users to choose from options that could confuse them, or would require them to study to make informed decisions.
-We assume this will mean the users will be more confident as there will be fewer steps where they were not sure if they have made the right choice.
-
-Since we are developing this application as a cross-platform desktop application available under a permissive open-source license, it should be usable by as many people as possible.
-
-\subsection{Localization}
-
-To make our application accessible to as many users as possible, we have also made the application translatable and try to display data like dates in the local format.
-
-When the application is started, we try to detect a locale of the user and set it as a default language.
-If we do not yet support the locale of the user, we default to the English language.
-At the time of writing, our application is fully translated in three languages: English (source language), Slovak, and Czech.
-
-The process of content translation is mostly automated.
-We start with the automated extraction of the source strings from the code of the main module and files provided by the backends.
-Next, we manually include the extracted source strings in the commit to our versioning control - we have retained manual control over this to prevent a noise that automated commits could make.
-On push to the hosted repository, a localization platform Transifex\footnote{Transifex is a SaaS localization platform free for OSS, see \url{https://www.transifex.com/}.} automatically pulls and process the updates.
-These phrases can then be translated by anybody using the UI of the Transifex platform to make it easily accessible.
-Once the changes are made on the Transifex platform, it automatically creates a pull request on our hosted repository with the changes made to that particular language.
-As we explain in detail in Section \ref{testing}, part of the build process is to incorporate these in the application.
 
@@
 
-Our code is kept in a git repository hosted at Github and is a source for the Continuous Integration (CI) pipeline that, on each push, runs static analysis on the code (), mentioned automated tests, builds the bundle and uploads created distributables for all platforms back to the Github.
+Our code is kept in a git repository hosted at Github and is a source for the Continuous Integration (CI) pipeline that, on each push, runs static analysis on the code, mentioned automated tests, builds the bundle and uploads created distributables for all platforms back to the Github.
 Publishing of the new release is then done by saving the release in the Github.
@@
 \begin{table}[h!]
-  \begin{tabular}{ p{3cm}|p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm} }
+  \begin{tabular}{ p{2.75cm}|p{0.85cm}p{0.85cm}p{0.85cm}p{0.85cm}p{0.85cm}p{0.85cm}p{0.85cm}p{0.85cm} }
                       & \rot{licensing} & \rot{cross-platform} & \rot{localization} & \rot{file suppport} & \rot{easy setup} & \rot{flexibility} & \rot{built-in verification} & \rot{extensibility} \\
     \hline
-    JSignPdf          & X               & /                   &                    &                      &                  & X                 &                             & / \\
-    zep.disig.sk      & /               & X                   & X                  & /                    & X                &                   & X                           &   \\
-    Acrobat DC        & /               & /                   & X                  &                      & /                & X                 & X                           &   \\
-    D.PDF Signer      &                 &                     & /                  &                      & X                & /                 &                             &   \\
+    JSignPdf          & \doublecheck    & \checkmark           &                    &                     &                  & \doublecheck      &                             & \checkmark   \\
+    zep.disig.sk      & \checkmark      & \doublecheck         & \doublecheck       & \checkmark          & \doublecheck     &                   & \doublecheck                &              \\
+    Acrobat DC        & \checkmark      & \checkmark           & \doublecheck       &                     & \checkmark       & \doublecheck      & \doublecheck                &              \\
+    D.PDF Signer      &                 &                      & \checkmark         &                     & \doublecheck     & \checkmark        &                             &              \\
     \hline
-    Our - Octosign    & X               & X                   & X                  & X                    & /                &                   & X                           & X \\
+    Our - Octosign    & \doublecheck    & \doublecheck         & \doublecheck       & \doublecheck        & \checkmark       &                   & \doublecheck                & \doublecheck \\
   \end{tabular}
@@
 Using our review from Chapter \ref{review}, we have compared our application in Table \ref{table:software}.
-Value \textit{X} means best-case scenario, no value worst-case scenario, and value \textit{/} somewhere in between - partially or depends.
+Value \doublecheck means best-case scenario, no value worst-case scenario, and value \checkmark somewhere in between - partially or depends.
 
@@
 
-Our website provides a simple \textbf{help page} with different use cases that should allow users to start using the application quickly.
+Our website provides a simple help page with different use cases that should allow users to start using the application quickly.
 It also contains a brief explanation of more advanced features of the application like settings.
@@
 
-Will be added later once the content is reviewed and it has a green light and we discuss what are the main points I should focus on here.
+We verified that all legal and technical preconditions are met for electronic signatures and should be, therefore, a viable alternative to handwritten signatures.
 
-Our goal is to explore an accessible way to use electronic signatures.
-In the beginning, we go over the preconditions that make it a viable alternative to the handwritten signatures.
-We cover the legal preconditions relevant for the European Union and Slovakia in particular.
-Also, we explain technical preconditions focusing on the cryptographic concepts like asymmetric cryptography, timestamping, hashing, and standards that comply with the law within the European Union.
+We did a review of the existing software focusing on the licensing, cross-platform compatibility, localization, file support, ease of setup, flexibility, verification ability, and extensibility.
+We identified the positive and negative sides of the current implementations and found a gap in the software for signing using the qualified electronic signatures.
 
-Secondly, we do a review of the existing software.
-We go quickly over several applications that are meant for basic use.
-We compare them using several criteria to find out the strengths and weaknesses of the current implementations and identify a potential space for improvements.
-
-Finally, we devote our attention to our own implementation.
-We explain the proposed desktop software and go into details on software engineering aspects.
-Specifically, modularity, architecture, usability, localization, testing, and automation.
-In the end, we compare it with the existing software and identify possible future improvements.
-
-Another result we focus on is a supporting website and how it helps with the use of the software.
-All in a way open for extensibility by the community.
+We proposed a simple software platform that aims to be straightforward for the end-user by choosing the most likely defaults and reducing the amount of information the user has to interact with, increasing the potential confidence of the user.
+Also, we made our application easy to translate using a third-party web UI where anyone from the community can contribute with translations.
+Our application was translated into three languages: English, Slovak, and Czech.
+Our solution is fully modular, dividing the main module that contains the UI and modules performing the document manipulation.
+We implemented a module for signing using advanced electronic signatures that follow the European standards and should be fully compliant with the European law, and a module that allows signing using a scanned signature or drawn signature where we take into consideration the velocity.
+Our software is a platform on which others can quickly build new ways to sign documents complying with different laws or to base further research.
+From the software engineering perspective, we also covered testing and automation and how it can be used to correctly test and maintain Electron applications like ours in practice.
+We compared our application with the existing software and found that there is room for improvement for our application in flexibility and ease of setup.
+We also created a supporting website that provides users with downloads for all major desktop platforms, help, and general information on electronic signatures.
 

### 2020-05-07 paper/main.tex
Fix last issues before submission to SVK

@@
 \NewDocumentCommand{\rot}{O{60} O{1em} m}{\makebox[#2][l]{\rotatebox{#1}{#3}}}%
+
+\hyphenation{elektro-nic-kých}
+\hyphenation{na-ria-dením}
 
@@
 So zmenami v právnom statuse elektronických podpisov vo svete sa vytvára príležitosť rozšíriť elektronické podpisovanie ako plnohodnotnú náhradu pre vlastnoručné podpisy.
-Elektronické podpisy sú dôležité čoraz viac v dobe elektronickej komunikácie, kedy sú vlastnoručné podpisy nepraktické z dôvodu nutnosti tlače, osobného kontaktu alebo zasielania podpísaných dokumentov.
+Elektronické podpisy sú dôležité čoraz viac v dobe elektronickej komunikácie, kedy sú vlastnoručné podpisy nepraktické z dôvodu nutnosti tlače, osobného kontaktu alebo zasielania podpísaných dokumentov.\\
 V práci sa venujeme právnym predpokladom relevantným pre EÚ, technologickým predpokladom, prehľadu v súčasnosti dostupného softvéru a návrhu vlastnej softvérovej platformy.
@@
 There are several preconditions that allow us to consider electronic signatures as an alternative.
-Firstly, the current state of the law and recent changes relating to the use of electronic signatures, secondly, well-established concepts in cryptography like \pojem{asymmetric cryptography} or \pojem{hashing}, and lastly, expertise from the field of software engineering - design, implementation, and maintenance of computer software.
+Firstly, the current state of the law and recent changes relating to the use of electronic signatures, secondly, well-established concepts in cryptography like asymmetric cryptography or hashing, and lastly, expertise from the field of software engineering - design, implementation, and maintenance of computer software.
 We will concentrate on the first two in relation to the electronic signatures and where applicable open-source desktop computer software.
@@
 
-With regards to electronic signatures, law within the EU used to differ, with the law applicable in the SR being now-repealed \pojem{Act No.
+With regards to electronic signatures, law within the EU used to differ, with the law applicable in the SR being now-repealed \textit{Act No.
 215/2002 Coll.}.
@@
 
-\pojem{Qualified electronic signatures} (QES) are a subset of \pojem{advanced electronic signatures} with more specific requirements that can be found in annex 1 of the eIDAS regulation\footnote{Available at https://eur-lex.europa.eu/legal-content/EN/TXT/HTML/?uri=CELEX:32014R0910\#d1e32-111-1.}.
+\pojem{Qualified electronic signatures} (QES) are a subset of advanced electronic signatures with more specific requirements that can be found in annex 1 of the eIDAS regulation\footnote{Available at https://eur-lex.europa.eu/legal-content/EN/TXT/HTML/?uri=CELEX:32014R0910\#d1e32-111-1.}.
 In more practical terms, such advanced electronic signatures are created with a qualified device using a qualified certificate and a qualified trust service.
@@
 
-Such right can be exercised to license the software under either \pojem{proprietary} or \pojem{open-source} license.
+Such right can be exercised to license the software under either proprietary or open-source license.
 \iffalse % Not part of the version for ŠVK
-\pojem{Proprietary} meaning under the exclusive legal right of the author, typically also confidential and distributed as a paid product.
-\pojem{Open-source} meaning having its source code freely available, typically also distributed for free and without any liability (the software is provided "as is").
+Proprietary meaning under the exclusive legal right of the author, typically also confidential and distributed as a paid product.
+Open-source meaning having its source code freely available, typically also distributed for free and without any liability (the software is provided "as is").
 
-Motivation to license the software under an \pojem{open-source} license can differ, and so do such licenses.
-While a \pojem{proprietary} license is usually made specifically for that entity and its interests, \pojem{open-source} licenses tend to be reused between different authors.
+Motivation to license the software under an open-source license can differ, and so do such licenses.
+While a proprietary license is usually made specifically for that entity and its interests, open-source licenses tend to be reused between different authors.
 This means that consumers of the software can quickly recognize their rights and responsibilities if they decide to use, modify, or distribute the software.
 \fi
-In general, we can classify the \pojem{open-source} licenses into two categories: \pojem{permissive} and \pojem{copyleft}.
+In general, we can classify the open-source licenses into two categories: \pojem{permissive} and \pojem{copyleft}.
 
-\pojem{Copyleft} license, in general, requires the user to publish the modified work under the compatible (free) license.
+Copyleft license, in general, requires the user to publish the modified work under the compatible (free) license.
 As a result, it forces them to extend the rights they have received onto others and, in turn, somehow hinder usability in software licensed under a different license.
@@
 
-\pojem{Permissive} licenses, on the other hand, do not have such a requirement in place and are more suitable for potential commercialization.
+Permissive licenses, on the other hand, do not have such a requirement in place and are more suitable for potential commercialization.
 They typically allow commercial use, modifications, distribution or sublicensing as long as the author is not held liable, and the original copyright and license are distributed with the software.
@@
 
-It is not uncommon to see \pojem{hybrid licensing} - licensing under more than one license, with one being typically \pojem{open-source} and one \pojem{proprietary}.
+It is not uncommon to see hybrid licensing - licensing under more than one license, with one being typically open-source and one proprietary.
 In this scenario, users can choose which license they want to use based on their needs.
@@
 
-Previously mentioned standards \pojem{PAdES}, \pojem{XAdES}, and \pojem{CAdES} define how \pojem{asymmetric cryptography}, \pojem{hashing}, or \pojem{timestamping} should be used to comply with the standards.
+Previously mentioned standards PAdES, XAdES, and CAdES define how asymmetric cryptography, hashing, or timestamping should be used to comply with the standards.
 We go over a these topics and why they are able to satisfy the requirements for QES mentioned in the Section \ref{eidas}.
@@
 
-These issues can be addressed with \pojem{asymmetric cryptography}.
+These issues can be addressed with asymmetric cryptography.
 In general, we assign a pair of keys to each person, one of them used for encryption, known as a \pojem{public key}, another used for decryption, known as a \pojem{private key}.
@@
 
-\pojem{Digital signatures} are specific \pojem{asymmetric cryptography} algorithms.
-The signer is the only one who possesses the \pojem{private key} used for signing, and others can use the \pojem{public key} for signature verification.
+\pojem{Digital signatures} are specific asymmetric cryptography algorithms.
+The signer is the only one who possesses the private key used for signing, and others can use the public key for signature verification.
 We sign the document, including the other parties' signature.
@@
 
-We are left with the problem of public key distribution - even though \pojem{public-key schemes} do not require a \textit{secure channel}, they require \textit{authenticated channels} for the distribution of the public keys \cite{cryptotxtbook}.
+We are left with the problem of public key distribution - even though public-key schemes do not require a \textit{secure channel}, they require \textit{authenticated channels} for the distribution of the public keys \cite{cryptotxtbook}.
 
 A common solution is the use of \pojem{certificates}.
-\pojem{Certificates} are essentially \pojem{digital signatures} with metadata that can be used to establish the validity of the received \pojem{public key} before it is used.
-Distribution of such \pojem{certificates} is the responsibility of a \pojem{Certification Authority} (CA) - a third party that all users in the network trust.
+Certificates are essentially digital signatures with metadata that can be used to establish the validity of the received public key before it is used.
+Distribution of such certificates is the responsibility of a \pojem{Certification Authority} (CA) - a third party that all users in the network trust.
 In the case of the Slovak ID cards, it is Disig - SVK eID Accredited CA, issued by the National Security Authority\footnote{Translation from "Národný bezpečnostný úrad".}.
-Such \pojem{certificates}, therefore, reliably uniquely link and identify the \pojem{public key} owners, as can be seen in Figure \ref{fig:certificate-scheme}\footnote{Created by an unknown author, licensed under the CC BY-SA 3.0 license.}.
+Such certificates, therefore, reliably uniquely link and identify the public key owners, as can be seen in Figure \ref{fig:certificate-scheme}\footnote{Created by an unknown author, licensed under the CC BY-SA 3.0 license.}.
 Ability to do that is the first requirement for the advanced electronic signatures mentioned in Section \ref{eidas}.
@@
 
-\subsection{Hashing}
+\subsection{Hashing and timestamping}
 
@@
 
-In cryptography, we are considering only \pojem{hash functions} that can also satisfy these additional requirements:
+In cryptography, we are considering only hash functions that can also satisfy these additional requirements:
 
 \begin{enumerate}
-  \item \textbf{Deterministic} - same input data are always mapped to the same \pojem{hash}.
-  \item \textbf{One-way} - it should be infeasible to retrieve input data for a given \pojem{hash}.
-  \item \textbf{Resistant to collisions} - it should be infeasible to find two inputs with the same \pojem{hash}.
-  \item \textbf{Avalanche effect} - even small change to the input should lead to significantly different \pojem{hash}.
+  \item \textbf{Deterministic} - same input data are always mapped to the same hash.
+  \item \textbf{One-way} - it should be infeasible to retrieve input data for a given hash.
+  \item \textbf{Resistant to collisions} - it should be infeasible to find two inputs with the same hash.
+  \item \textbf{Avalanche effect} - even small change to the input should lead to significantly different hash.
 \end{enumerate}
 
-In the context of \pojem{digital documents}, we can use such hash function to generate a hash for the document.
+In the context of digital documents, we can use such hash function to generate a hash for the document.
 This hash can represent a document of any size and is used during the creation of a digital signature instead of the document itself.
@@
 
-\subsection{Timestamping}
-
 A common part of the paper signed documents is the date when they were signed.
@@
 
-To achieve this, a trusted third party - \pojem{Timestamping Authority} (TSA) - is the one that will sign the \pojem{hash} of the document together with the current date and time.
+To achieve this, a trusted third party - \pojem{Timestamping Authority} (TSA) - is the one that will sign the hash of the document together with the current date and time.
 This process can be seen in Figure \ref{fig:trusted-timestamping}\footnote{Schema created by Bart Van den Bosch, licensed under the CC BY-SA 2.0 BE license.}.
@@
 
-Extensions for PDF files, XML files, and CMS signed data for signing using advanced electronic signatures, as explained in Section \ref{eidas}, are \pojem{PAdES}, \pojem{XAdES}, and \pojem{CAdES}, respectively.
+Extensions for PDF files, XML files, and CMS signed data for signing using advanced electronic signatures, as explained in Section \ref{eidas}, are PAdES, XAdES, and CAdES, respectively.
 They are defined in the respective specifications by the ETSI in \cite{pades}, \cite{xades}, and \cite{cades}.
@@
 
-  \caption{Overview of standards PAdES, XAdES, and CAdES}
+  \caption{Overview of standards PAdES, XAdES, and CAdES.}
   \label{table:1}
@@
 
+We have summarized our findings in Table \ref{table:software}.
+Value \doublecheck means best-case scenario, no value worst-case scenario, and value \checkmark somewhere in between - partially or depends.
+
+\begin{table}[h!]
+  \begin{tabular}{ p{2.75cm}|p{0.85cm}p{0.85cm}p{0.85cm}p{0.85cm}p{0.85cm}p{0.85cm}p{0.85cm}p{0.85cm} }
+                      & \rot{licensing} & \rot{cross-platform} & \rot{localization} & \rot{file suppport} & \rot{easy setup} & \rot{flexibility} & \rot{built-in verification} & \rot{extensibility} \\
+    \hline
+    JSignPdf          & \doublecheck    & \checkmark           &                    &                     &                  & \doublecheck      &                             & \checkmark \\
+    zep.disig.sk      & \checkmark      & \doublecheck         & \doublecheck       & \checkmark          & \doublecheck     &                   & \doublecheck                &            \\
+    Acrobat DC        & \checkmark      & \checkmark           & \doublecheck       &                     & \checkmark       & \doublecheck      & \doublecheck                &            \\
+    D.PDF Signer      &                 &                      & \checkmark         &                     & \doublecheck     & \checkmark        &                             &            \\
+  \end{tabular}
+
+  \caption{Comparison of existing software.}
+  \label{table:software}
+\end{table}
+
 \section{Desktop application JSignPdf}
@@
 
-\obrazok{./figures/screenshot-jsignpdf}{Screenshot of the JSignPdf version 1.6.4}{screenshot-jsignpdf}
+\obrazok{./figures/screenshot-jsignpdf}{Screenshot of the JSignPdf version 1.6.4.}{screenshot-jsignpdf}
 
@@
 
-\obrazok{./figures/screenshot-zepdisigsk}{Screenshot of the zep.disig.sk}{screenshot-zepdisigsk}
+\obrazok{./figures/screenshot-zepdisigsk}{Screenshot of the zep.disig.sk.}{screenshot-zepdisigsk}
 
@@
 
-\obrazok{./figures/screenshot-dpdfsigner}{Screenshot of the D.PDF Signer}{screenshot-dpdfsigner}
+\obrazok{./figures/screenshot-dpdfsigner}{Screenshot of the D.PDF Signer.}{screenshot-dpdfsigner}
 
@@
 Downsides are lack of support for platforms other than Windows, localization, and built-in signature verification.
-
-\section{Overview}
-
-We have summarized our findings in Table \ref{table:software}.
-Value \doublecheck means best-case scenario, no value worst-case scenario, and value \checkmark somewhere in between - partially or depends.
-
-\begin{table}[h!]
-  \begin{tabular}{ p{2.75cm}|p{0.85cm}p{0.85cm}p{0.85cm}p{0.85cm}p{0.85cm}p{0.85cm}p{0.85cm}p{0.85cm} }
-                      & \rot{licensing} & \rot{cross-platform} & \rot{localization} & \rot{file suppport} & \rot{easy setup} & \rot{flexibility} & \rot{built-in verification} & \rot{extensibility} \\
-    \hline
-    JSignPdf          & \doublecheck    & \checkmark           &                    &                     &                  & \doublecheck      &                             & \checkmark \\
-    zep.disig.sk      & \checkmark      & \doublecheck         & \doublecheck       & \checkmark          & \doublecheck     &                   & \doublecheck                &            \\
-    Acrobat DC        & \checkmark      & \checkmark           & \doublecheck       &                     & \checkmark       & \doublecheck      & \doublecheck                &            \\
-    D.PDF Signer      &                 &                      & \checkmark         &                     & \doublecheck     & \checkmark        &                             &            \\
-  \end{tabular}
-
-  \caption{Comparison of existing software}
-  \label{table:software}
-\end{table}
 
@@
 
-\obrazok{./figures/modularity}{Modules of our application and their communication}{modularity}
+\obrazok{./figures/modularity}{Modules of our application and their communication.}{modularity}
 
@@
 
-\subsection{Downloads}
-
 Downloads are available for each major desktop platform in a format suitable for installation or integration in the operating system.
@@
 In turn, we have some level of confidence source code used for the distributables comes from one of the authors.
-
-\subsection{Help and resources}
 
@@
 
-We proposed a simple software platform that aims to be straightforward for the end-user by choosing the most likely defaults and reducing the amount of information the user has to interact with, increasing the potential confidence of the user.
+Therefore, we proposed a simple open-source software platform that aims to be straightforward for the end-user by choosing the most likely defaults and reducing the amount of information the user has to interact with, increasing the potential confidence of the user.
 Also, we made our application easy to translate using a third-party web UI where anyone from the community can contribute with translations.
 Our application was translated into three languages: English, Slovak, and Czech.
-Our solution is fully modular, dividing the main module that contains the UI and modules performing the document manipulation.
-We implemented a module for signing using advanced electronic signatures that follow the European standards and should be fully compliant with the European law, and a module that allows signing using a scanned signature or drawn signature where we take into consideration the velocity.
-Our software is a platform on which others can quickly build new ways to sign documents complying with different laws or to base further research.
-From the software engineering perspective, we also covered testing and automation and how it can be used to correctly test and maintain Electron applications like ours in practice.
-We compared our application with the existing software and found that there is room for improvement for our application in flexibility and ease of setup.
-We also created a supporting website that provides users with downloads for all major desktop platforms, help, and general information on electronic signatures.
+It is also is fully modular, dividing the main module that contains the UI and modules performing the document manipulation.
+We implemented a module for signing using advanced electronic signatures that follow the European standards and should be fully compliant with the European law, and a module that allows signing using a scanned signature or drawn signature.
+The proposed software is a platform on which others can quickly build new ways to sign documents complying with different laws or to base further research.
+From the software engineering perspective, the implementation is robust, and we covered testing and automation in connection to Electron applications like ours.
+On the other hand, there is some room for improvement in flexibility and ease of setup of our software.
+We also maintain a supporting website that provides download links for all major desktop platforms, a help page, and general information on electronic signatures.
 

### 2020-05-16 paper/main.tex
Remove extra new line in paper
Add progress - student scientific conference
Add changes to paper from svk to bachelor thesis

@@
 \hyphenation{na-ria-dením}
+\hyphenation{možno-sťami}
+\hyphenation{používateľ-sky}
+\hyphenation{akých-koľ-vek}
 
@@
 \author{Jakub Ďuraš}
-\iffalse % Not part of the version for ŠVK
 \typprace{bachelor's}
-\fi
-\inytypprace{Student Scientific Conference}
 \rok{2020}
@@
 \pracovisko{Institute of Computer Science}
-\iffalse % Not part of the version for ŠVK
 \podakovanie{
-  I would like to thank my supervisor Viliam Kačala for his continuous feedback and patience with my question.
-
-  I would also like to acknowledge help received from Pavol Sokol and Laura Rózenfeldová on legal questions, and access to various resources from Gabriel Semanišin.
-
+I want to thank my supervisor RNDr. Viliam Kačala for his continuous feedback and patience with my questions and late submissions.
+I would also like to acknowledge help received from RNDr. JUDr. Pavol Sokol, PhD. on legal questions, and access to resources from prof. RNDr. Gabriel Semanišin, PhD.
 Finally, I would like to thank my family and friends who contributed in various ways to my thesis.
 }
-\fi
 
-\iffalse % Not part of the version for ŠVK
-\pdfzadanie{zadanie.pdf}
-\fi
+\pdfzadanie{assignment.pdf}
 
@@
 With the recent changes in the legal status of electronic signatures in the world, there is an opportunity to expand the use of electronic signatures as an alternative to handwritten signatures outside the usual areas.
-Electronic signatures are becoming increasingly important in the age of electronic communication when handwritten signatures are unpractical due to the need to print, personally exchange, or send signed documents.
 We devote our attention to the exploration of legal preconditions relevant for the EU, technological preconditions, review of the existing software, and proposal of our own software platform.
-The result is open, user-friendly software for Windows, Mac OS X, and Linux.
+The result is an open software platform for operating systems Windows, Mac OS X, and Linux.
 Using the application, it is possible, following the eIDAS Regulation, to sign and verify the signature of any document with the legal effect of an attested signature.
-Our solution supports ID cards of the Slovak Republic with the eID chip without any modifications or settings, which enables easy use outside the usual use in web applications of the state.
+Our solution supports ID cards of the Slovak Republic with the eID chip without any modifications or settings, which enables use outside the web applications of the state.
 The application is not tied only to the Slovak eID but is ready for further use with various devices in different countries and as a platform for other types of signatures.
-Another part is a website that, in addition to the possibility of downloading software and brief documentation, provides general information about electronic signatures so that the project is as accessible to the public as possible.
-We concentrate on simple extensibility of the program via language translations, documentation, functionality, and general information.
+Another part is a website that, in addition to the possibility of downloading software and brief documentation, provides general information about electronic signatures.
+The emphasis is on simple extensibility of the program via language translations, documentation, functionality, and general information.
+The thesis can be further extended with novel ways of signing or by improving flexibility, which is not up to par with compared software.
 }
@@
 So zmenami v právnom statuse elektronických podpisov vo svete sa vytvára príležitosť rozšíriť elektronické podpisovanie ako plnohodnotnú náhradu pre vlastnoručné podpisy.
-Elektronické podpisy sú dôležité čoraz viac v dobe elektronickej komunikácie, kedy sú vlastnoručné podpisy nepraktické z dôvodu nutnosti tlače, osobného kontaktu alebo zasielania podpísaných dokumentov.\\
 V práci sa venujeme právnym predpokladom relevantným pre EÚ, technologickým predpokladom, prehľadu v súčasnosti dostupného softvéru a návrhu vlastnej softvérovej platformy.
-Výsledkom je otvorený, používateľsky prívetivý softvér pre platformy Windows, Mac OS X a Linux.
+Výsledkom je otvorená softvérová platforma pre operačné systémy Windows, Mac OS X a Linux.
 Pomocou aplikácie je možné, v súlade s nariadením eIDAS, podpísať a overiť podpis ľubovoľného dokumentu, ktorý má rovnaký právny účinok ako osvedčený podpis.
-Naše riešenie podporuje občianske preukazy Slovenskej Republiky s čipom eID bez akýchkoľvek úprav alebo nastavení, čím sa umožní ich jednoduché využitie mimo obvyklého použitia vo webových aplikáciách štátu.
+Naše riešenie podporuje občianske preukazy Slovenskej Republiky s čipom eID bez akýchkoľvek úprav alebo nastavení, čím sa umožňuje ich využitie mimo použitia vo webových aplikáciách štátu.
 Aplikácia nie je viazaná iba na slovenské eID, ale je pripravená na ďalšie použitie s rôznymi zariadeniami v rôznych štátoch a ako platforma pre iné druhy podpisov.
-Súčasťou je aj webová stránka, ktorá okrem možnosti stiahnutia softvéru a stručnej dokumentácie poskytuje všeobecné informácie o elektronických podpisoch, aby bol projekt čo najviac prístupný verejnosti.
-Dôraz kladieme na jednoduchú rozšíriteľnosť programu vo forme jazykových prekladov, dokumentácie, funkcionality a všeobecných informácií.
+Súčasťou je aj webová stránka, ktorá okrem možnosti stiahnutia softvéru a stručnej dokumentácie poskytuje všeobecné informácie o elektronických podpisoch.
+Dôraz je kladený na jednoduchú rozšíriteľnosť programu vo forme jazykových prekladov, dokumentácie, funkcionality a všeobecných informácií.
+Prácu je možné ďalej rožšíriť nepreskúmanými možnosťami podpisovania či zlepšením flexibility, ktorá nie je na úrovni niektorých zástupcov porovnávaného softvéru. 
 }
@@
 \tableofcontents
-\iffalse % Not part of the version for ŠVK
-\listoffigures
-\listoftables
-\fi
 
@@
 We explain the proposed desktop software and go into details on software engineering aspects.
-Specifically, modularity, architecture, usability, localization, testing, and automation.
+Specifically, we cover usability, localization, modularity, architecture, testing, and automation.
 In the end, we compare it with the existing software and identify possible future improvements.
@@
 
-\iffalse % Not part of the version for ŠVK
-TODO: Think about adding section on handwritten signatures
-\fi
-
 \section{Law}
 
-We are considering law applicable locally in the Slovak Republic (SR).
+Unless stated otherwise, we are considering law applicable locally in the Slovak Republic (SR).
 Its law is greatly influenced by the European Union (EU), being its member since 2004, and by the rest of the world.
 We can assume this is at least partially applicable outside of the SR as well.
+The \cite{law} shapes our overall view on the global legal status of signatures and chapters 1, 2, 3, 6, and 7 of \cite{osslicensing} our Chapter \ref{copyright}.
 
@@
 
-\obrazok{./figures/electronic-advanced-qualified}{The relation between electronic, advanced, and qualified signatures.}{signature-comparison}
+\obrazok{./figures/electronic-advanced-qualified}{Relation between electronic, advanced, and qualified signatures.}{signature-comparison}
 
@@
 
-\pojem{Qualified electronic signatures} (QES) are a subset of advanced electronic signatures with more specific requirements that can be found in annex 1 of the eIDAS regulation\footnote{Available at https://eur-lex.europa.eu/legal-content/EN/TXT/HTML/?uri=CELEX:32014R0910\#d1e32-111-1.}.
+\pojem{Qualified electronic signatures} (QES) are a subset of advanced electronic signatures with more specific requirements that can be found in annex 1 of the eIDAS regulation\footnote{Available at \url{https://eur-lex.europa.eu/legal-content/EN/TXT/HTML/?uri=CELEX:32014R0910\#d1e32-111-1}.}.
 In more practical terms, such advanced electronic signatures are created with a qualified device using a qualified certificate and a qualified trust service.
@@
 Such right can be exercised to license the software under either proprietary or open-source license.
-\iffalse % Not part of the version for ŠVK
 Proprietary meaning under the exclusive legal right of the author, typically also confidential and distributed as a paid product.
@@
 This means that consumers of the software can quickly recognize their rights and responsibilities if they decide to use, modify, or distribute the software.
-\fi
 In general, we can classify the open-source licenses into two categories: \pojem{permissive} and \pojem{copyleft}.
@@
 As a result, it forces them to extend the rights they have received onto others and, in turn, somehow hinder usability in software licensed under a different license.
-\iffalse % Not part of the version for ŠVK
-A popular example of such license is the GNU General Public License(GPL) and its derivatives like GNU Lesser GPL (LGPL) or GNU Affero GPL (AGPL).
-\fi
+A popular example of such license is the GNU General Public License (GPL) and its derivatives like GNU Lesser GPL (LGPL) or GNU Affero GPL (AGPL).
 
@@
 They typically allow commercial use, modifications, distribution or sublicensing as long as the author is not held liable, and the original copyright and license are distributed with the software.
-\iffalse % Not part of the version for ŠVK
 Often used permissive licenses are MIT License, Apache License, or BSD License.
-\fi
 
@@
 
-\iffalse % Not part of the version for ŠVK
 As we have already mentioned, software licenses can influence whether and how it can be used as a part of different software.
 Both when we are the ones using someone else's software and when someone else is using our software.
-\fi
 Generally speaking, the more open the license is, the higher the probability it will not prevent adoption for legal reasons.
@@
 
-\iffalse % Not part of the version for ŠVK
+\iffalse % I have no space for this :(
 Various possible combinations can be seen in Figure \ref{fig:license-compatibility}, where greyed out options mean such combination is not possible unless a separate licensing agreement is reached with the copyright owner.
@@
 Previously mentioned standards PAdES, XAdES, and CAdES define how asymmetric cryptography, hashing, or timestamping should be used to comply with the standards.
-We go over a these topics and why they are able to satisfy the requirements for QES mentioned in the Section \ref{eidas}.
+We go over these topics and why they are able to satisfy the requirements for QES mentioned in the Section \ref{eidas}.
+We draw our knowledge in chapters about cryptographic primitives from the \cite{cryptotxtbook} and \cite{cryptojoy}.
 
@@
 
-\iffalse % Not part of the version for ŠVK
-\begin{figure}[!htb]
-  \centering
-
-  \begin{tikzpicture}
-    \datavisualization [scientific axes=clean, visualize as smooth line, all axes={ticks={major={at={}}}}, y axis={label={number of keys}}, x axis={label={number of users}}]
-    data [format=function] {
-      var x : interval [2:125];
-      func y = \value x*\value x;
-    };
-  \end{tikzpicture}
-
-  \centerline{\parbox{13cm}{\sl \caption{Cubic growth of the number of keys.}}}
-\end{figure}
-\fi
-
 Either party can also lie about not signing the document.
@@
 
-\iffalse % Not part of the version for ŠVK
-\obrazok{./figures/digital-signature-scheme}{Principle of digital signatures which involves signing and verifying a message (Paar 2009). TODO: I don't have rights to this image!!!}{pubkey-scheme}
-\fi
-
 Practical, widely used implementation of this idea is the \pojem{RSA} signature scheme that relies on the integer factorization problem published in 1978 \cite{rsa}.
@@
 In Slovakia, the private key is usually handed over to the user personally, saved on the ID card.
-This ID card (used together with a card reader) provides an API for safe access, management, and cryptoprocessing.
+This ID card (used together with a card reader) provides an \pojem{Application Programming Inteface} (API) for safe access, management, and cryptoprocessing.
 Such a device is in general called a \pojem{Hardware Security Module} (HSM), and the API relevant for this thesis is \pojem{PKCS \#11} \footnote{PKCS \#11 specification available at \url{http://docs.oasis-open.org/pkcs11/pkcs11-base/v2.40/os/pkcs11-base-v2.40-os.html}.}.
@@
 
-\iffalse % Not part of the version for ŠVK
-TODO: Mention EU LOTL
-\fi
+Specific for the EU is online \pojem{List of The Lists} (LOTL) that contains a list with trusted service providers for each country.
+Each member state is obligated to maintain such list, as can be seen in Article 22 of \cite{eidas}.
+From the technical point of view, it is an XML document that is signed and can be processed automatically.
 
@@
 
-\iffalse % Not part of the version for ŠVK
-TODO: Add scheme of document > hash - valid signature, change in document > different hash - invalid signature.
-\fi
-
 An example of commonly used hash function family is \pojem{Secure Hash Algorithms} (SHA), with hash functions relevant to this thesis being mainly SHA-2\footnote{SHA-2 specification available in FIPS PUB 180-4 at \url{https://nvlpubs.nist.gov/nistpubs/FIPS/NIST.FIPS.180-4.pdf}.}.
@@
 
-\subsection{PAdES, XAdES, and CAdES}
+\subsection{PAdES, XAdES, and CAdES} \label{etsi}
 
-Extensions for PDF files, XML files, and CMS signed data for signing using advanced electronic signatures, as explained in Section \ref{eidas}, are PAdES, XAdES, and CAdES, respectively.
+Extensions for Portable Document Format (PDF) files, eXtensible Markup Language (XML) files, and Cryptographic Message Syntax (CMS) signed data for signing using advanced electronic signatures, as explained in Section \ref{eidas}, are PAdES, XAdES, and CAdES, respectively.
 They are defined in the respective specifications by the ETSI in \cite{pades}, \cite{xades}, and \cite{cades}.
@@
 All of them support multiple signatures applied in succession.
-\iffalse % Not part of the version for ŠVK
 Implementation of the PAdES is rather portable since any compatible PDF file reader or editor will work with the signatures in a similar way - signature created in one should be easily verifiable in any other.
 XAdES and CAdES, however, are usually tied to specific implementation or standard that is built on top of them.
-\fi
 
-\begin{table}[h!]
+\begin{table}[!ht]
   \begin{tabular}{ |p{2cm}||p{2.2cm}|p{2cm}|p{3cm}|p{3.9cm}| }
@@
 The expiration time is then limited by the validity of the timestamp certificate.
+Different profiles - levels - for different standards can be seen in Table \ref{table:2}.
 
-\iffalse % Not part of the version for ŠVK
+\begin{table}[!ht]
+  \centering
+  \begin{tabular}{|l||l|l|l|}
+    \hline
+    Level                     & XAdES     & CAdES     & PAdES     \\
+    \hline
+    Basic                     & XAdES-B   & CAdES-B   & AdES-B    \\
+    B with Timestamp          & XAdES-T   & CAdES-T   & PAdES-T   \\
+    T with Long Term Data     & XAdES-LT  & CAdES-LT  & PAdES-LT  \\
+    LT with Archive timestamp & XAdES-LTA & CAdES-LTA & PAdES-LTA \\
+    \hline
+  \end{tabular}
+
+  \caption{Different profiles - levels - for standards PAdES, XAdES, and CAdES.}
+  \label{table:2}
+\end{table}
+
+\iffalse % I have no space for this :(
 \section{Software engineering}
@@
 Aspects we will focus on are design, implementation, testing and releasing, and maintenance.
-Small
-
-TODO: Give definition to the software engineering, what aspects we are going over, our focus on the engineering relevant for the open source.
+We try to consider specifics of the open-source software as they are sometimes defined in Chapter 7 of \cite{oss}.
 
@@
 
-\iffalse % Not part of the version for ŠVK
-\chapter{Goals}
-
-\section{Software}
-
-TODO: Basis can be formed on the requirements article https://github.com/durasj/octosign/wiki/Initial-Requirements
-
-TODO: Basic software requirements and more specific software requirements.
-
-TODO: Enginnering parts - architecture, testing, maintainability.
-
-TODO: Backends
-
-\section{Information}
-
-TODO: Website - multilingual, editable by the community
-TODO: Global status - world map
-\fi
-
 \chapter{Review of existing software} \label{review}
@@
 Excluded are all complex applications with advanced features and high price tag.
-We have picked one open-source desktop application, one web application - Software as a Service (SaaS), one PDF viewer, and one simple paid desktop application. Criteria we focus on are:
+We have picked one open-source desktop application, one web application - \pojem{Software as a Service} (SaaS), one PDF viewer, one simple paid desktop application, and one commercial desktop application with free version.
+Criteria we focus on are:
 
@@
     D.PDF Signer      &                 &                      & \checkmark         &                     & \doublecheck     & \checkmark        &                             &            \\
+    Podpisuj.sk       & \checkmark      & \doublecheck         & \checkmark         & \doublecheck        & \doublecheck     & \checkmark        & \doublecheck                &            \\
   \end{tabular}
@@
 
-\section{Commercial desktop application D.PDF Signer}
+\section{Commercial application D.PDF Signer}
 
@@
 
-Just as the zep.digis.sk, D.PDF signer utilizes the currently bundled software with the SR eID.
+Just like the zep.digis.sk, D.PDF signer utilizes the currently bundled software with the SR eID.
 It is straightforward and flexible.
-Downsides are lack of support for platforms other than Windows, localization, and built-in signature verification.
+On the other thand, it lacks support for platforms other than Windows.
+Additionally, the application is only partially localized to English.
+It does not support file formats other than PDF and has no built-in signature verification.
 
-\iffalse % Not part of the version for ŠVK
-\section{Modularity}
+\section{Commercial desktop application Podpisuj.sk}
 
-TODO: Rough overview of different ways modular architecture can be achieved. Focus on modularity via multiple independant applications.
+Application Podpisuj.sk is a paid proprietary desktop application with a free version developed locally in Slovakia.
 
-\subsection{Communication}
+\obrazok{./figures/screenshot-podpisujsk}{Screenshot of the Podpisuj.sk.}{screenshot-podpisujsk}
 
-TODO: Communication in multiple independant applications - IPC, Sockets, Pipes, STDIO.
-
-\subsection{Bundling}
-
-\subsection{Communicating and bundling with GPL software}
-\fi
+The Podpisuj.sk can be downloaded for all three major desktop platforms.
+It does not rely on the software bundled with the SR eID.
+It is relatively easy to use and comes fully preconfigured so it can be used with the SR eID right away.
+Flexibility is not as great as with some other applications we have reviewed, and some options (like custom TSP) require one of the paid versions.
+Not everything is fully localized - especially the website, and any documentation we could find was available only in Slovak.
+It is not possible to extend it, but it is easy to use different tokens for signing.
 
@@
 We have developed an open-source, cross-platform desktop software with focus on the modularity, architecture, usability, localization, automated testing with overall automation, and extensibility by the community.
-The source code is available at \url{https://github.com/durasj/octosign}, \url{https://github.com/durasj/octosign-dss}, and \url{https://github.com/durasj/octosign-image}.
+Our knowledge on these topics comes from Ian Sommerville's \cite{sommerville}.
+We try to consider specifics of the open-source software as they are sometimes defined in Chapter 7 of \cite{oss}.
 
-\subsection{Usability and localization}
+\subsection{Usability}
 
@@
 
+\subsection{Localization}
+
 We have also made the application easily translatable and try to display data like dates in the local format.
@@
 At the time of writing, our application is fully translated in three languages: English (source language), Slovak, and Czech.
+Additional languages - German and Spanish - are actively being developed.
 
@@
 On push to the hosted repository, a localization platform Transifex\footnote{Transifex is a SaaS localization platform free for OSS, see \url{https://www.transifex.com/}.} automatically pulls and process the updates.
-These phrases can then be translated by anybody using the UI of the Transifex platform to make it easily accessible.
+These phrases can then be translated by anybody using the \textit{User Interface} (UI) of the Transifex platform to make it easily accessible.
 Once the changes are made on the Transifex platform, it automatically creates a pull request on our hosted repository with the changes made to that particular language.
@@
 
-Signing modules - backends - are separate CLI applications written in any programming language.
+Signing modules - backends - are separate \pojem{Command Line Interface} (CLI) applications written in any programming language.
 They are responsible for the actual document manipulation - signing and verification.
 
-The main module loads all available backends at runtime and verifies their availability, options, and supported file types.
+The main module loads all available backends at runtime.
 Once the user decides to verify or sign a file, they are executed as CLI applications.
 
-Communication between the main module and backends is in plaintext over the standard streams - standard input, output, and error.
-To make sure we have correctly documented and easily accessible way for backend development, we have prepared a Backend specification\footnote{Backend specification is available at \url{https://github.com/durasj/octosign/wiki/Backend-specification}.} that specifies what is required of the backend, what is optionally supported, and the format of the inputs, outputs, and error handling.
-
-\iffalse % Not part of the version for ŠVK
-TODO: Go a bit more over the backend specification here
-\fi
+Communication between the main module and backends is in plaintext over the standard streams (STDIO) - standard input (STDIN), output (STDOUT), and error (STDERR).
+To make sure we have correctly documented and easily accessible way for backend development, we have prepared a Backend specification that specifies what is required of the backend, what is optionally supported, and the format of the inputs, outputs, and error handling.
 
@@
 
+\subsection{Signing backends} \label{backends}
+
+Following our specification, each backend is required to provide a configuration file and follow the communication protocol.
+
+Configuration files are written in YAML\footnote{YAML is a human-readable data serialization language, see \url{https://yaml.org/}.} and contain basic information about the backend like name, description, repository, version, author, or license.
+It also contains two properties defining what command should be executed and with what parameters to execute the backend, and build the backend, respectively.
+
+Backend is always called with operation as the first argument, followed by an additional argument that differs between operations.
+The backend should be stateless - that means it should not be influenced by any state it would save like info about what files it signed.
+All communication is in UTF-8 encoded plaintext.
+It needs to implement operations \textit{meta}, \textit{sign}, and \textit{verify} that are run using the configured executable, for example, \textit{./backend-executable meta}, or, \textit{./backend-executable sign /home/path.pdf}.
+
+Operation meta is called to get info about its availability, options, and supported file types.
+This allows the backend to check the platform, available software of the user, etc. and make the decision based on that.
+
+Operation sign and verify are called to sign or verify the file with the absolute path to the file.
+The main module never sends anything over the STDIO unless it's asked to do so from the backend.
+That means the backend acts as a \textit{master} and the main module as a \textit{slave}.
+The UI can end the process if any unexpected situation arises and displays an error.
+The backend can during those two operations, at any moment: write error to the STDERR that is immediately displayed; prompt for additional information by writing to the STDOUT; get the current value of any option by writing to the STDOUT.
+
+Ending the process with the exit code 0 signals success.
+Some operations expect a result on the STDOUT before the process exits.
+If code other than 0 is used, the backend should always send an error message to the STDERR.
+
+Messages exchanged via the STDIN and STDOUT can be sometimes polluted by different libraries used on either part, so they are distinguished from the rest by using the following delimiter:
+
+\begin{verbatim}
+---TYPE---
+any information exchanged in the appropriate format
+---TYPE---
+\end{verbatim}
+
+\textit{TYPE} here is replaced by \textit{PROMPT}, \textit{GETOPTION}, or \textit{RESULT} based on the type of the message.
+Such messages should be written in one chunk and flushed. Communication on the STDERR is accepted and displayed as-is.
+
+Prompt is used to interactively ask for more information from the user.
+The backend can ask to get an absolute path to the output file, absolute path to opened file, text, password, image (drawn or picked), or position and width of the signature that will be placed in the UI by the user.
+
+Getting an option is used to ask for the current option value that can be set in the UI on the Settings screen.
+The option is automatically prefixed on the main module for each backend so it can not retrieve options from other backends to prevent collisions and security issues.
+
+Such communication protocol allows using any language that can simply work with the STDIO without a need for knowledge or third-party libraries to work with serialized data in formats like JSON.
+The backend does not have to use any regular expressions but always simply reads the whole line.
+It is supposed to be as simple as possible for newcomers.
+
+The current version of the full specification is available at \url{https://github.com/durasj/octosign/wiki/Backend-specification}.
+
+\subsubsection{DSS eIDAS backend}
+
+The backend that we call \pojem{Advanced Electronic Signature} (AES), is a Java CLI application that uses \pojem{Digital Signature Service} (DSS) framework. 
+The DSS is a project from the Connecting Europe Facility (CEF) Digital, a European Union fund for pan-European infrastructure investment.
+It is one of the \textit{}{Building Blocks}, as they are called on their website, which helps to create and verify electronic signatures in line with European standards.
+Its documentation, available in \cite{ecdss}, also provides general information about the used standards.
+
+Source of the private key is configurable, and we can use PKCS \#11 to utilize those in HSMs, PKCS \#12 to work with key stores in files, or Microsoft Crypto API that provides an abstraction on different sources.
+It supports all standards mentioned in Chapter \ref{etsi} in various configurations.
+Since we are trying to simplify it down for the user, when signing, we automatically choose to use PAdES when signing PDF, XAdES when signing XML with enveloped signature, and ASiC with CAdES for all other file formats.
+This could be further extended in the future, so the user can manually choose the preferred standard, container type, or signature policy.
+
+Signature verification, however, works with all different combinations that are detected automatically.
+During the verification, we construct a trusted list of root certificates from the LOTL.
+Different lists for different countries are hosted individually on each countries servers.
+For better performance and reliability when there is an outage on those servers, we proxy requests through CloudFlare\footnote{CloudFlare is a CDN provider, see \url{https://www.cloudflare.com/}.} \pojem{Content Delivery Network} (CDN) to distribute cached files that are periodically downloaded and stored on the Amazon Web Services\footnote{Amazon Web Services is a cloud provider with various services, see \url{https://aws.amazon.com/what-is-aws/}.} (AWS) infrastructure using the AWS Lambda and AWS S3.
+
+We also support adding a visible signature for PDFs.
+This helps users to recognize it was digitally signed without a need to mention it in the document itself and provides information on who or when signed it.
+Labels are translated using the current language of the application as we assume this will be the language the receiver will most likely speak.
+
+The backend currently supports configurable TSP URL, so any TSP following the RFC 3161 can be utilized as long as it does not require authentication.
+Authentication is usually done using a name and password or using a digital signature but is currently not supported.
+This could hinder use with qualified TSPs that usually require authentication.
+
+One issue we were facing when working with the DSS framework is the assumption that it is often implemented on the server-side as a service.
+This meant that some APIs are meant to be used in an environment where they are started once, and then they are available or can be run periodically several times a day.
+This is obviously not the case with our architecture since we always run it only when some operation is required, and the application is stopped again.
+We assume this issue can come up during the development of other backends as well and may require a certain level of flexibility from the developer to tackle them.
+
+\subsubsection{Image backend}
+
+To provide users with the ability to sign using a simple image saved on the computer, which is probably a scanned signature, or signature drawn directly in our application, we created what we call \textit{Simple image signature} backend.
+It is useful when users do not have the resources required to sign the document using the QES, or in situations when they would rather not do so.
+
+It is written in Go and utilizes a library for manipulation of PDF files.
+Its purpose is just to ask the main module for the signature file and its position.
+Therefore, this backend utilizes prompt functionality of the communication protocol with options image and position.
+
+Option to draw the signature in the application is using a splining algorithm with a variable stroke that calculates the time it takes to travel between two points.
+Such signatures mimic one of the features of the handwritten signatures from \cite{verification} that are used to verify the authenticity of handwritten signatures.
+That being said, we are not aware of any proof that this approach has any practical benefits outside of aesthetics.
+This approach was outlined on the Square Engineering Blog\footnote{Smoother Signatures article available archived at \url{https://web.archive.org/web/20150211220342/http://corner.squareup.com/2012/07/smoother-signatures.html}.}.
+There is a JavaScript implementation for the browser called Signature Pad\footnote{Signature Pad library is available at \url{https://github.com/szimek/signature_pad}.}.
+
 \subsection{Architecture of the main module} \label{architecture}
@@
 Our main module is the most complex one.
-Electron applications have a main thread that is running a Node.js\footnote{Node.js is a JavaScript runtime for use outside of the browser with easier access to the OS APIs.} and a UI thread running the relevant part of the Chromium web browser.
-Their communication is done via IPC and allows access to the Node.js APIs in the UI thread.
+Electron applications have a main thread that is running a Node.js\footnote{Node.js is a JavaScript runtime for use outside of the browser with easier access to the OS APIs.} and a UI thread running the relevant part of the Chromium\footnote{Chromium is a popular open-source browser project, see \url{https://www.chromium.org/Home}.}.
+Their communication is done via \pojem{Inter-process communication} (IPC) and allows access to the Node.js APIs in the UI thread.
 Considering the JavaScript is interpreted at runtime, we are exposing ourselves to potential attacks (mostly XSS\footnote{XSS means Cross-site scripting, see \url{https://owasp.org/www-community/attacks/xss/}.}).
 To reduce the attack surface, as Node.js allows easy access to many APIs like full access to the filesystem, we have fully separated the UI thread and main thread.
+Our work was influenced by knowledge from \cite{eloquentjs} and \cite{ydkjs}.
 
@@
 
-When approaching automated testing, we have decided to follow the opinion of the engineers at Google\footnote{Approach outlined in the article \textit{Just Say No to More End-to-End Tests} available at \url{https://testing.googleblog.com/2015/04/just-say-no-to-more-end-to-end-tests.html}.}.
-We write three types of tests: end-to-end, integration, and unit tests.
+When approaching automated testing, we have decided to implement the testing pyramid as defined by the Mike Cohn\footnote{Mike Cohn came up with this concept in his book \textit{Succeeding with Agile}.}.
+One of the implementations of this idea was done by the Ham Vocke in the article \textit{The Practical Test Pyramid}\footnote{Available on the website of Martin Fowler at \url{https://martinfowler.com/articles/practical-test-pyramid.html}.}.
+It is implemented at a scale at Google, which explains their positive experience with it\footnote{Approach outlined in the article \textit{Just Say No to More End-to-End Tests} available at \url{https://testing.googleblog.com/2015/04/just-say-no-to-more-end-to-end-tests.html}.}.
+Following this approach, we write three types of tests: end-to-end, integration, and unit tests.
 The ratio of these types of tests should form a pyramid where the heavy bottom is formed by the unit tests, followed by a smaller set of component tests forming the middle of the pyramid, and a small set of end-to-end tests forming the tip of the pyramid.
+The motivation behind this approach is to write more tests that are faster and break less often (as they are more isolated) as opposed to tests that are slower and break more often.
 
@@
 
-Our code is kept in a git repository hosted at Github and is a source for the Continuous Integration (CI) pipeline that, on each push, runs static analysis on the code, mentioned automated tests, builds the bundle and uploads created distributables for all platforms back to the Github.
+Our code is kept in a git repository hosted at Github and is a source for the Continuous Integration (CI) and Continuous Delivery (CD) pipeline.
+For the CI/CD, we are using Azure Pipelines\footnote{Cloud service from Microsoft that is part of the Azure DevOps, see \url{https://docs.microsoft.com/en-us/azure/devops/pipelines/get-started/what-is-azure-pipelines?view=azure-devops}.} specifically, that are free for OSS.
+On each push, apart from setting up the environment, we perform static analysis on the code checking the type safety using TypeScript, code style using the ESLint and Prettier, code quality using the CodeClimate.com, and check for potential vulnerabilities using the Snyk.io.
+As for the mentioned automated tests, we run the unit and components tests and monitor the coverage using the Codecov.io.
+In the case coverage would be significantly negatively impacted, the Pull Request (PR) is blocked from being merged.
+Backends are referenced in the same repository by using the git submodules, which allows us to simplify the development and build process.
+Build process build individual backends, the main thread, UI thread, and API code, and uploads bundled distributable files for all platforms back to the Github.
+The whole process runs for each platform - Windows, Linux, and Mac - separately; therefore, there is no need for cross-compiling and running of platform-dependent tools in virtualized environments.
 Publishing of the new release is then done by saving the release in the Github.
+
+We had to overcome issues we have faced with the performance of the Github relases - mainly apparent bandwidth restrictions.
+We have deployed an AWS lambda that is called when a new release is published on Github and sends a message to AWS Simple Queue Service (SQS).
+Another AWS Lambda is triggered on such message, and it downloads all distributable files and creates a metafile with information about the release in AWS S3.
+These files are then served cached over the CDN from CloudFlare.
 
@@
     D.PDF Signer      &                 &                      & \checkmark         &                     & \doublecheck     & \checkmark        &                             &              \\
+    Podpisuj.sk       & \checkmark      & \doublecheck         & \checkmark         & \doublecheck        & \doublecheck     & \checkmark        & \doublecheck                &              \\
     \hline
-    Our - Octosign    & \doublecheck    & \doublecheck         & \doublecheck       & \doublecheck        & \checkmark       &                   & \doublecheck                & \doublecheck \\
+    Our - Octosign    & \doublecheck    & \doublecheck         & \doublecheck       & \doublecheck        & \checkmark       & \checkmark        & \doublecheck                & \doublecheck \\
   \end{tabular}
@@
 
-In the future, we can improve our application to be easier to setup for more users, and be more flexibile when it comes to the source of the private key and certificate, or timestamping.
+In the future, we can improve our application to be easier to set up for more users and be more flexible in the configuration of TSPs or used signature policy.
 That being said, we need to consider each configurable functionality we add as it increases complexity which could mean our software is harder to understand.
@@
 Goal of the website is to both support the application in the form of Download page and Help page and provide a to provide a general information on electronic signatures.
-The website is available at \url{https:\\octosign.com} and source code at \url{https://github.com/durasj/octosign-website}.
 The website is, at the time of writing, available in English and Slovak language.
@@
 \obrazok{./figures/website-home}{Home page of the octosign.com website.}{website-home}
+
+\subsection{Downloads}
 
@@
 
+To improve on this further, we are doing a \pojem{code-signing} for each release using a trusted way for each platform.
+On Mac OS X, this means enrolling in the Apple Developer Program that requires an annual fee and installation of signing certificates.
+Additionally, it is necessary to \textit{notarize} the application for use on Mac OS X.
+Notarization is an automated process required by Apple to prove that the application does not contain any malicious code and is properly signed and configured.
+If the application is not notarized, Gatekeeper will warn the user about the potentially unsafe software.
+On modern versions of Windows, applications that are not code-signed can trigger a warning or, depending on the settings, even completely prevent the user from using software that is not code-signed using a certificate that is cosigned with a trusted root certificate.
+Certification for individuals on the Mac OS seems to be fully automated, does not involve lengthy verification, and costs 99 USD a year.
+Certification on Windows is offered by various trusted third parties like DigiCert.com that we have used and involves lengthy, partially automated verification at a price starting from 60 USD a year.
+At the time of writing, there does not seem to be any service offering code-signing certificates to OSS for free.
+
+\subsection{Information and help}
+
 Our website provides a simple help page with different use cases that should allow users to start using the application quickly.
@@
 
+\subsection{Technologies}
+
+From the technological point of view, the website is a \pojem{Single Page Application} (SPA) written in React\footnote{React is a JavaScript library for building user interfaces, see \url{https://reactjs.org/}.}.
+It is built using Gatsby\footnote{Gatsby is a free and open-source framework based on React, see \url{https://www.gatsbyjs.org/}.}.
+Even though it is a SPA, it is \pojem{prerendered} to static HTML and CSS files that are served from CDN, which improves performance and allows for better \pojem{Search Engine Optimization} (SEO).
+
+The website is using CI/CD services from Netlify.com that is also responsible for distribution using a CDN.
+Each push on the master branch triggers a build and deploy on the live website.
+Each PR has a deploy preview that is basically a unique link that shows proposed changes built and deployed on the web.
+Such approach allows for easy iteration and allows others to contribute the content in markdown files without a need to build the website locally as they can see results as soon as they open the PR.
+
+Different language mutations are denoted by using a suffix, for example, the home page is written in English and Slovak using file names \textit{home.en.tsx} and \textit{home.sk.tsx}.
+Same applies for content written in markdown, for example, help page on basic usage is written in files \textit{basic.en.md} and \textit{basic.sk.md}.
+The URL of the website contains language code of the current language if it is other than English, for example, the download page is available in English on path \textit{/download} and in Slovak on path \textit{/sk/download}.
+By switching the language using the selector on the upper right, the URL changes to the required format.
+
+TODO: If you mind this blank space, I can add a scheme with source files in various languages > gatsby > CDN flow.
+
 \zaver
 
-We verified that all legal and technical preconditions are met for electronic signatures and should be, therefore, a viable alternative to handwritten signatures.
+We verified that the main legal and technical preconditions should be met for the use of electronic signatures.
+They should be, therefore, a viable alternative to handwritten signatures.
 
@@
 
-Therefore, we proposed a simple open-source software platform that aims to be straightforward for the end-user by choosing the most likely defaults and reducing the amount of information the user has to interact with, increasing the potential confidence of the user.
+Therefore, we proposed a simple open-source software platform that aims to be straightforward for the end-user by choosing the most likely defaults and reducing the amount of information the user has to interact with.
 Also, we made our application easy to translate using a third-party web UI where anyone from the community can contribute with translations.
-Our application was translated into three languages: English, Slovak, and Czech.
-It is also is fully modular, dividing the main module that contains the UI and modules performing the document manipulation.
-We implemented a module for signing using advanced electronic signatures that follow the European standards and should be fully compliant with the European law, and a module that allows signing using a scanned signature or drawn signature.
+Our application is available in three languages: English, Slovak, and Czech, with German and Spanish in translation.
+It is also fully modular, dividing the main module that contains the UI and modules performing the document manipulation.
+We implemented a module for signing using qualified electronic signatures that follow the European standards and should be compliant with the European law, and a module that allows signing using a scanned or drawn signature.
 The proposed software is a platform on which others can quickly build new ways to sign documents complying with different laws or to base further research.
@@
 On the other hand, there is some room for improvement in flexibility and ease of setup of our software.
-We also maintain a supporting website that provides download links for all major desktop platforms, a help page, and general information on electronic signatures.
+We also created a supporting website that provides download links for all major desktop platforms, a help page, and general information on electronic signatures.
+Our project should be open to people from different countries, speaking different languages, and open for contributions from people with various skills - not just from programmers.
+
+\chapter*{Resumé}
+\addcontentsline{toc}{chapter}{Resumé}
+
+Overili sme, že hlavné predpoklady pre použitie elektronických podpisov by mali byť naplnené.
+Mali by byť preto schodnou alternatívou pre vlastnoručné podpisy.
+
+Preskúmali sme existujúci softvér so zameraním na licenciu, kompatibilitu medzi platformami, lokalizáciu, podporu súborov, ľahkosť nastavenia, flexibilitu, schopnosť overiť podpisy a rozšíriteľnosť.
+Identifikovali sme pozitívne a negatívne stránky súčasných implementácií a našli sme medzeru v softvéri na podpisovanie pomocou kvalifikovaných elektronických podpisov.
+
+Preto sme navrhli jednoduchú otvorenú softvérovú platformu, ktorá sa snaží byť priamočiara pre koncového používateľa výberom pravdepodobných predvolených hodnôt a znížením množstva informácií s ktorými sa používateľ stretáva.
+Taktiež sme uľahčili preklad našej aplikácie pomocou webového rozhrania tretej strany, s ktorým môže ktokoľvek z komunity pomôcť prekladmi.
+Naša aplikácia je dostupná v troch jazykoch: angličtine, slovenčine a češtine s nemčinou a španielčinou v prekladaní.
+Je tiež plne modulárna s rozdelením hlavného modulu, ktorý obsahuje používateľské rozhranie, a modulov vykonávajúcich manipuláciu s dokumentmi.
+Implementovali sme modul na podpisovanie pomocou kvalifikovaného elektronického podpisu, ktorý spĺňa európske štandardy a mal by byť v súlade s európskymi zákonmi, a modul, ktorý umožňuje podpisovanie pomocou naskenovaného alebo nakresleného podpisu.
+Navrhovaný softvér je platforma, na ktorej môžu ostatní rýchlo vybudovať nové spôsoby podpisovania dokumentov v súlade s rôznymi zákonmi alebo založiť ďalší výskum.
+Venovali sme sa testovaniu a automatizácii vývoja Electron aplikácií, takže implementácia je robustná aj z pohľadu softvérového inžinierstva.
+Na druhej strane existuje priestor na zlepšenie flexibility a jednoduchosti prvotného nastavenia nášho softvéru.
+Vytvorili sme tiež podpornú webovú stránku s možnosťou stiahnutie pre hlavné počítačové operačné systémy, pomocou a všeobecnými informáciami o elektronických podpisoch.
+Náš projekt by mal byť otvorený ľuďom z rôznych krajín, rozprávajúcich rôznymi jazkymi, a zmenám od ľudí s rôznymi zručnosťami - nie len programátorom.
 
@@
 Available at: \url{https://humanities-digital-library.org/index.php/hdl/catalog/view/electronicsignatures/1/86-1}
-
-  \bibitem{licensing}
-  \osoba{Morin, A.}, et al.
-2012.
-A Quick Guide to Software Licensing for the Scientist-Programmer.
-In: \emph{PLoS computational biology} [online].
-Volume 8, issue 7 [cit.
-2020-01-27].
-Available at: \url{https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1002598}
 
@@
 
-  %\bibitem{2}
-  %\osoba{Beck, G.}, 2007. \emph{Zakázaná rétorika: 30 manipulatívních technik}. Preklad
-  %\osoba{Pomikálová, M.}. Praha: Grada Publishing. ISBN 978-80-247-1743-2.
-  %\bibitem{3}
-  %\osoba{Vojčík, P.}, 2010. \emph{Občianske právo hmotné II.} 3. prep. a dopl. vyd. Košice: UPJŠ v Košiciach. ISBN 978-80-7097-817-7.
-  %  2 autori
-  %\bibitem{4}
-  %\osoba{Šoltés, M.} a \osoba{Radoňák, J.}, 2013. \emph{Základné princípy laparoskopickej chirurgie.} Košice: UPJŠ v Košiciach. ISBN 978-80-8152-074-7.
-  % 3 autori
-  %\bibitem{5}
-  %\osoba{Guzanin, Š.}, \osoba{Sabovčík, R.} a \osoba{Kačmár, P.}, 2004.
-\emph{Selected Chapters of Plastic and Reconstructive Surgery: vysokoškolské učebné texty}.
-Košice: Univerzita Pavla Jozefa Šafárika v Košiciach, Lekárska fakulta.
-ISBN 80-7097-557-1.
-  % 4 a viac autorov
-  %\bibitem{6}
-  %\osoba{Nagyová, I.} et al. 2009. \emph{Measuring health and quality of life in the chronically ill}. Košice: Equilibria. ISBN 978-80-892-8446-7.
-  % Elektronická kniha
-  %\bibitem{7}
-  %\osoba{Speight, J.
-G.}, 2005.
-\emph{Lange's Handbook of Chemistry} [online].
-London: McGraw-Hill.
-[cit.
-2009.06.10.] ISBN 978-1-60119-261-5.
-Dostupné na: \url{http://www.knovel.com/web/portal/basic_search/display?_EXT_KNOVEL_DISPLAY_bookid=1347&_EXT_KNOVEL_DISPLAY_fromSearch=true&_EXT_KNOVEL_DISPLAY_searchType=basic}
-  % zborník
-  %\bibitem{8}
-  %\osoba{Bačkor, M.} a \osoba{Mihaličová, S.}, zost., 2013.
-\emph{Zborník príspevkov z konferencie 11.
-dni doktorandov experimentálnej biológie rastlín a 13.
-konferencie experimentálnej biológie rastlín} [online].
-Košice: Univerzita Pavla Jozefa Šafárika v Košiciach, Prírodovedecká fakulta [cit.
-2009-06-10].
-ISBN 9788081520327.
-Dostupné na: \url{
-  %  http://www.upjs.sk/public/media/5596/PF-Zbornik-prispevkov-konferencie-11-dni-doktorandov.pdf}
-  %
-  % 2. Časopis (ako celok)
-  %\bibitem{9}
-  %\emph{Thaiszia: Journal of Botany}. Košice: P.J.Safarik University, Botanic Garden, \mbox{1990--\ .} ISSN 1210-0420.
-  %\bibitem{10}
-  %\emph{Ikaros: elektronický časopis o informační bezpečnosti} [online], 2002. [Praha]: Ikaros. 1997--{} [cit. 2002-03-08]. Dostupné na: \url{http://www.ikaros.cz/}. ISSN 1212-5075.
-  % Jedno číslo časopisu
-  %\bibitem{11}
-  %\emph{CHIP: magazín informačních technologií}, 2013. Praha: Burda Praha, roč. 23, říjen. ISSN 1210-0684.
-  %  ------
-  %  3. Príspevok v knihe/zborníku
-  %  ------
-  %\bibitem{12}
-  %\osoba{Sabol, J.}, 2000. Jazyk ako ľudské posolstvo: (namiesto doslovu). In: \emph{O jazyku a štýle kriticky aj prakticky}. Prešov: Náuka, s. 149--159. ISBN 809676022X.
-  %\bibitem{13}
-  %\osoba{Tóthová, E.} a kol., 2013.
-A rare t(9,22,16)(q34,q11,q24) translocation in chronic myeloid leukemia for which imatinib mesylate was effective: a case report.
-In: \emph{XXVII.
-Olomoucké hematologické dny s mezinárodní účastí, 12.--14.5.2013, Olomouc: sborník abstrakt}.
-Olomouc: Univerzita Palackého v Olomouci, s.
-75--76.
-ISBN 9788024434803.
-  %  ------
-  % 4. Článok v časopise
-  %  ------
-  %\bibitem{14}
-  %\osoba{Beňačka, J.} et al., 2009.
-A better cosine approximate solution to pendulum equation.
-In: \emph{International Journal of Mathematical Education in Science and Technology}.
-Vol.
-40, no.
-2, p.
-206--215.
-ISSN 0020-739X.
-  %\bibitem{15}
-  %\osoba{Dubayová, T.} et al., 2010.
-The impact of the intensity of fear on patient's delay regarding health care seeking behavior: a systematic review vyhľadaní zdravotníckej starostlivosti.
-In: \emph{International Journal of Public Health}.
-Vol.
-55, no.
-5, p.
-459--468.
-ISSN 1661-8556.
-  %\bibitem{16}
-  %\osoba{Steinerová, J.}, 2000. Princípy formovania vzdelania v informačnej vede. In: \emph{Pedagogická revue}. Roč. 2, č. 3, s. 8--16. ISSN 1335-1982.
-  %\bibitem{17}
-  %\osoba{Hoggan, D.}, 2002.
-Challenges, Strategies, and Tools for Research Scientists.
-In: \emph{Electronic Journal of Academic and Special Librarianship} [online].
-Vol.
-3, no.
-3 [cit.
-2013-01-10].
-ISSN 1525-321X.
-Dostupné na: \url{http://southernlibrarianship.icaap.org/content/v03n03/Hoggan_d01.htm}
-  %\bibitem{18}
-  %\osoba{Srbecká, Gabriela}, 2010.
-Rozvoj kompetencí studentů ve vzdělávání.
-In: \emph{Inflow: information journal} [online].
-Roč.
-3, č.
-7 [cit.
-2013-08-06].
-ISSN 1802-9736.
-Dostupné na: \url{http://www.inflow.cz/rozvoj-kompetenci-studentu-ve-vzdelavani}
-  %  ------
-  % 5. Príspevok v zborníku na CD-ROM
-  %  ------
-  %\bibitem{19}
-  %\osoba{Zemánek, P.}, 2001.
-The machines for ``green works'' in vineyards and their economical evaluation.
-In: \emph{9th International Conference: proceedings.
-Vol.
-2.
-Fruit Growing and viticulture} [CD-ROM].
-Lednice: Mendel University of Agriculture and Forestry, p.
-262--268.
-ISBN 80-7157-524-0.
-  %  ------
-  % 6. Záverečné a kvalifikačné práce
-  %  ------
-  %\bibitem{20}
-  %\osoba{Mikulášiková, M.}, 1999. \emph{Didaktické pomôcky pre praktickú výučbu na hodinách výtvarnej výchovy pre 2. stupeň základných škôl}: diplomová práca. Nitra: UKF.
-  %\bibitem{21}
-  %\osoba{Urdzík, P.}, 2007. \emph{Predikcia intrauterinnej rastovej retardácie a preeklampsie pomocou biochemických a ultrazvukových markerov}: dizertačná práca. Košice: UPJŠ v Košiciach.
-  %  ------
-  % 7. Výskumné správy
-  %  ------
-  %\bibitem{22}
-  %\osoba{Baumgartner, J.} a kol., 1998. \emph{Ochrana a udržiavanie genofondu zvierat, šľachtenie zvierat}: výskumná správa. Nitra: VÚŽV.
-  %  ------
-  % 8. Normy
-  %  ------
-  %\bibitem{23}
-  %STN ISO 690: 2012. \emph{Informácie a dokumentácia. Návod na tvorbu bibliografických odkazov na informačné pramene a ich citovanie}.
-  %  ------
-  %  ------
-  % 10. Zákon
-  %  ------
-  %\bibitem{25}
-  %\emph{Zákon č.131/2002 Zb. o vysokých školách a o zmene a doplnení niektorých zákonov.}
-  %\bibitem{26}
-  %\emph{Zákon č. 313/2001 o verejnej službe.}
+  \bibitem{ecdss}
+  \osoba{Vandenbroucke, P.} and \osoba{Beliakov, A.} [online].
+\emph{Digital Signature Service}.
+Luxembourg: CEF Digital [cit.
+2020-05-16].
+Available at: \url{https://ec.europa.eu/cefdigital/DSS/webapp-demo/doc/dss-documentation.html}
+
+  \bibitem{verification}
+  \osoba{Hafemann, G.
+L.}, \osoba{Sabourin, R.}, and \osoba{Oliveira, S.
+L.}, 2017.
+Offline Handwritten Signature Verification - Literature Review.
+In: \emph{2017 Seventh International Conference on Image Processing Theory, Tools and Applications (IPTA)}, pp.
+1-8.
+Montreal: IEEE.
+ISBN 978-1-5386-1842-4.
+Available at: \url{https://arxiv.org/abs/1507.07909}
 
@@
 
+\prilohy
+An up-to-date version of each developed asset is available online on the URL specified in the list.
+The version available at the time of writing is part of the attached CD, and the user manual is also available as Appendix A.
+
+List of the most important assets available online and on the CD:
+
+\begin{itemize}
+  \item Software Octosign - repository with source code: \url{https://github.com/durasj/octosign}.
+  \item Software Octosign - released distributable files: \url{https://github.com/durasj/octosign/releases}.
+  \item Signing backend DSS - repository with source code: \url{https://github.com/durasj/octosign-dss}.
+  \item Signing backend Image - repository with source code: \url{https://github.com/durasj/octosign-image}.
+  \item Website octosign.com - repository with source code: \url{https://github.com/durasj/octosign-website}.
+  \item Documentation - short user manual: \url{https://octosign.com/help/basics}.
+  \item Documentation - signing backend specification: \url{https://github.com/durasj/octosign/wiki/Backend-specification}.
+\end{itemize}
+
+\priloha{User manual}
+
+\section{Downloading and installing the application}
+
+Open the subpage Download on website \url{https://octosign.com} where you can download the application by clicking on the button for your operating system.
+After the download is complete, open the file and follow the instructions.
+
+\section{Signing the document}
+
+First, it is needed to select a document. To select a document, click on the button \textit{SELECT FILES} or drag and drop your document on the application.
+The recommended document type is PDF.
+After selecting the document, a card with the document name will appear just like in the picture.
+
+\scaledobrazok{./figures/screenshot-manual-unsigned}{Screenshot with unsigned document.}{manual-unsigned}{0.5}
+
+Signing is possible using several different types of signatures visible in the upper part of the screen.
+The default type is \textit{Advanced electronic signature}, if available, else \textit{Simple image signature}.
+When signing with an image, there is only a drawn or picked signature image placed on the PDF.
+When signing with an advanced electronic signature, it is possible to use a qualified electronic signature.
+To sign a document, press the button \textit{SIGN} visible on the document card.
+Follow the instructions, and after the signing is successful, the document will be marked as \textit{Signed}.
+
+\scaledobrazok{./figures/screenshot-manual-signed}{Card with the signed document.}{manual-signed}{0.5}
+
+\section{Verifying of document signature}
+
+Verification of the document signature is automatically started after the document is selected, and it does not have to be triggered manually.
+
+\scaledobrazok{./figures/screenshot-manual-verifying}{Card with the document that is being verified.}{manual-verifying}{0.5}
+
+After the verification is completed, the status of the document will be visible on the card of the document, it can be: \textit{Signed}, \textit{Unsigned}, \textit{Invalid}, \textit{Unknown}, and \textit{Indeterminate}.
+
+If there is at least one signature attached to the document, it is possible to see more information about it by clicking on the button \textit{OPEN SIGNATURE DETAILS}.
+
+\section{Settings}
+
+Settings can be opened by clicking on the text \textit{Settings} in the lower right corner of the application.
+
+It is possible to set the language and options specific for the type of the signature.
+
+\scaledobrazok{./figures/screenshot-manual-settings}{Screenshot with open Settings.}{manual-settings}{0.5}
+
+Options available for the Advanced electronic signature backend are:
+
+\textbf{PKCS \#11 Library Path} - The path is automatically prefilled if there is a supported known software installed on the computer.
+If not, it is possible to fill it manually with the path that you can get from the supplier of the software bundled with your device.
+The path needs to lead to a library compliant with the PKCS \#11 standard in a 64-bit version.
+
+\textbf{Timestamping Server URL} - The URL is automatically prefilled with value "http://timestamp.digicert.com".
+In the case you would like to use your own timestamping server, feel free to change it, but please make sure to type in the full URL address, including the protocol ("http://" or "https://").
+
 \end{document}

### 2020-05-17 paper/main.tex
Fix typo in paper
Simplify acknowledgments
Add paper improvements based on feedback

@@
 \hyphenation{akých-koľ-vek}
+\hyphenation{zariadenia-mi}
+\hyphenation{ope-račné}
 
@@
 I want to thank my supervisor RNDr. Viliam Kačala for his continuous feedback and patience with my questions and late submissions.
-I would also like to acknowledge help received from RNDr. JUDr. Pavol Sokol, PhD. on legal questions, and access to resources from prof. RNDr. Gabriel Semanišin, PhD.
+I would also like to acknowledge help received from RNDr. JUDr. Pavol Sokol, PhD. and prof. RNDr. Gabriel Semanišin, PhD.
 Finally, I would like to thank my family and friends who contributed in various ways to my thesis.
@@
 We devote our attention to the exploration of legal preconditions relevant for the EU, technological preconditions, review of the existing software, and proposal of our own software platform.
-The result is an open software platform for operating systems Windows, Mac OS X, and Linux.
+The result is an open software platform for operating systems Windows, macOS, and Linux.
 Using the application, it is possible, following the eIDAS Regulation, to sign and verify the signature of any document with the legal effect of an attested signature.
@@
 V práci sa venujeme právnym predpokladom relevantným pre EÚ, technologickým predpokladom, prehľadu v súčasnosti dostupného softvéru a návrhu vlastnej softvérovej platformy.
-Výsledkom je otvorená softvérová platforma pre operačné systémy Windows, Mac OS X a Linux.
+Výsledkom je otvorená softvérová platforma pre operačné systémy Windows, macOS a Linux.
 Pomocou aplikácie je možné, v súlade s nariadením eIDAS, podpísať a overiť podpis ľubovoľného dokumentu, ktorý má rovnaký právny účinok ako osvedčený podpis.
@@
 
-While, specifically in the European Union, electronic signatures seem to be relatively extensively used in communication with the government, they do not seem to be widely adopted in the exchange between individuals or smaller organizations.
-The problem is that handwritten signatures are often unpractical for electronic communication due to the need of physical presence or printing and sending of documents.
+While, specifically in the European Union, electronic signatures seem to be relatively extensively used in communication with the government, they do not seem to be widely adopted in the exchange between individuals or smaller organizations, which still seem to prefer handwritten signatures.
+In many cases, however, they may be unpractical for electronic communication due to the need of physical presence or printing and sending of documents.
 
@@
 There are several preconditions that allow us to consider electronic signatures as an alternative.
-Firstly, the current state of the law and recent changes relating to the use of electronic signatures, secondly, well-established concepts in cryptography like asymmetric cryptography or hashing, and lastly, expertise from the field of software engineering - design, implementation, and maintenance of computer software.
+Firstly, the current state of the law and recent changes relating to the use of electronic signatures.
+Secondly, well-established concepts in cryptography like asymmetric cryptography or hashing.
+Lastly, expertise from the field of software engineering - design, implementation, and maintenance of computer software.
 We will concentrate on the first two in relation to the electronic signatures and where applicable open-source desktop computer software.
@@
 We can assume this is at least partially applicable outside of the SR as well.
-The \cite{law} shapes our overall view on the global legal status of signatures and chapters 1, 2, 3, 6, and 7 of \cite{osslicensing} our Chapter \ref{copyright}.
+Our overall view on the global legal status is shaped by \cite{law}.
+Subsection \ref{copyright} is largely based on Chapters 1, 2, 3, 6, and 7 of \cite{osslicensing}.
 
@@
 
-Signature is considered an "attested signature"\footnote{Translation from "osvedčený podpis" as used in the Slovak law.} if it is verified by an authorized third party.
-This process is referred to as legalization in Notary Law, § 56, and its purpose is to attest information that could form the basis for the exercise of rights or which could cause legal consequences \cite{notarylaw}.
+Signature is considered an \pojem{attested signature}\footnote{Translation from "osvedčený podpis" as used in the Slovak law.} if it is verified by an authorized third party.
+This process is referred to as \pojem{legalization} in Notary Law, § 56, and its purpose is to attest information that could form the basis for the exercise of rights or which could cause legal consequences \cite{notarylaw}.
 
@@
 
-It is not uncommon to see hybrid licensing - licensing under more than one license, with one being typically open-source and one proprietary.
+It is not uncommon to see hybrid licensing - which means licensing under more than one license, with one being typically open-source and one proprietary.
 In this scenario, users can choose which license they want to use based on their needs.
@@
 
-There are several issues that make symmetric encryption unsuitable.
+There are several attributes that make symmetric encryption unsuitable.
 Let us imagine we are creating a key for each pair of users in our network - imaginary country Isle of Alices and Bobs - for symmetric encryption and decryption of documents shared between each Alice and Bob.
@@
 
-Practical, widely used implementation of this idea is the \pojem{RSA} signature scheme that relies on the integer factorization problem published in 1978 \cite{rsa}.
+Practical and widely used implementation of this idea is the \pojem{RSA} signature scheme that relies on the integer factorization problem published in 1978 \cite{rsa}.
 
 In Slovakia, the private key is usually handed over to the user personally, saved on the ID card.
-This ID card (used together with a card reader) provides an \pojem{Application Programming Inteface} (API) for safe access, management, and cryptoprocessing.
+This ID card, together with a card reader, provides an \pojem{Application Programming Inteface} (API) for safe access, management, and cryptoprocessing.
 Such a device is in general called a \pojem{Hardware Security Module} (HSM), and the API relevant for this thesis is \pojem{PKCS \#11} \footnote{PKCS \#11 specification available at \url{http://docs.oasis-open.org/pkcs11/pkcs11-base/v2.40/os/pkcs11-base-v2.40-os.html}.}.
@@
   \item \textbf{Licensing} - is it freely available - both in terms of money and source code - best case is open-source; worst case is paid proprietary software.
-  \item \textbf{Cross-platform compatibility} - is it available for users of all major desktop platforms - best case is availability on Windows, Mac OS X, and Linux; worst case is availability only on one platform.
+  \item \textbf{Cross-platform compatibility} - is it available for users of all major desktop platforms - best case is availability on Windows, macOS, and Linux; worst case is availability only on one platform.
   \item \textbf{Localization} - is it translated and does it display data like dates in local format - best case is availability at least in English and Slovak language; worst case is availability only in one language with no support for localization of dates, etc..
@@
 
-JSignPdf is an open-source Java application that adds digital signatures to PDF documents. It can be used as a standalone application or as an Add-On in OpenOffice. We focus only on standalone use.
+JSignPdf is an open-source Java application that adds digital signatures to PDF documents. It can be used as a standalone application or as an add-on in OpenOffice. We focus only on standalone use.
 
@@
 Additionally, we have no way of knowing "how things work" in the background.
-File support is pretty extensive, but there is a strict restriction on file size at 4 MB - scanned documents or simply larger files will be a problem.
+File support is pretty extensive, but there is a strict restriction on file size at 4 MB - larger files such as documents scanned at home will possibly not be accepted.
 The website also provides valuable information, and the company seems to be open to questions from users - all completely for free.
@@
 Setup of the certificate for use with the SR eID can be a hit or miss.
-On one computer, the application would not include a certificate from the SR eID, and we had to go through a process of adding the PKCS \#11 library ourselves.
+On one of our computers, the application would not include a certificate from the SR eID, and we had to go through a process of adding the PKCS \#11 library manually.
 That meant we had to go through several steps and windows and disable the \textit{protected mode} option.
-The setup of this application should be done for the user ideally by the IT personnel.
+The setup of this application may not be straightforward and may require professional help.
 On another computer, we were simply able to choose from the list of the available certificates.
@@
 The application is proprietary and can not be extended or verified.
-It is available only on the Windows and Mac OS X.
+It is available only on the Windows and macOS.
 
@@
 It is straightforward and flexible.
-On the other thand, it lacks support for platforms other than Windows.
+On the other hand, it lacks support for platforms other than Windows.
 Additionally, the application is only partially localized to English.
@@
 Not everything is fully localized - especially the website, and any documentation we could find was available only in Slovak.
+Therefore it may be impractical for foreigners.
 It is not possible to extend it, but it is easy to use different tokens for signing.
@@
 
-The main module is an Electron\footnote{Electron is a framework allowing the development of desktop software using web technologies.} application written in the language TypeScript\footnote{TypeScript is statically typed superset of JavaScript that compiles to JavaScript.}.
+The main module is an Electron\footnote{Electron is a framework allowing the development of desktop software using web technologies.} application written in TypeScript\footnote{TypeScript is statically typed superset of JavaScript that compiles to JavaScript.}.
 This module features simple UI and abstraction around modules for signing that we call \pojem{backends}.
@@
 
-Signing modules - backends - are separate \pojem{Command Line Interface} (CLI) applications written in any programming language.
+Signing modules - backends - are separate command line interface (CLI) applications written in any programming language.
 They are responsible for the actual document manipulation - signing and verification.
@@
 We expect most of the languages have APIs for the consumption of the standard streams with excellent portability across different desktop platforms.
-Secondly, since we are working on separate applications, they can be written by different people under a different license.
+Secondly, since we are technically working on separate applications, they can be written by different people under a different license.
 That also means we can easily swap the main module or any of the backends for use in different environments or different bundles.
@@
 All communication is in UTF-8 encoded plaintext.
-It needs to implement operations \textit{meta}, \textit{sign}, and \textit{verify} that are run using the configured executable, for example, \textit{./backend-executable meta}, or, \textit{./backend-executable sign /home/path.pdf}.
+It needs to implement operations \texttt{meta}, \texttt{sign}, and \texttt{verify} that are run using the configured executable, for example, \texttt{./backend-executable meta}, or, \texttt{./backend-executable sign /home/path.pdf}.
 
@@
 The main module never sends anything over the STDIO unless it's asked to do so from the backend.
-That means the backend acts as a \textit{master} and the main module as a \textit{slave}.
+That means the backend acts as a \texttt{master} and the main module as a \texttt{slave}.
 The UI can end the process if any unexpected situation arises and displays an error.
@@
 
-\textit{TYPE} here is replaced by \textit{PROMPT}, \textit{GETOPTION}, or \textit{RESULT} based on the type of the message.
+\texttt{TYPE} here is replaced by \texttt{PROMPT}, \texttt{GETOPTION}, or \texttt{RESULT} based on the type of the message.
 Such messages should be written in one chunk and flushed. Communication on the STDERR is accepted and displayed as-is.
@@
 
-The backend that we call \pojem{Advanced Electronic Signature} (AES), is a Java CLI application that uses \pojem{Digital Signature Service} (DSS) framework. 
+The backend that we call \pojem{advanced electronic signature}, is a Java CLI application that uses \pojem{Digital Signature Service} (DSS) framework. 
 The DSS is a project from the Connecting Europe Facility (CEF) Digital, a European Union fund for pan-European infrastructure investment.
@@
 
-To provide users with the ability to sign using a simple image saved on the computer, which is probably a scanned signature, or signature drawn directly in our application, we created what we call \textit{Simple image signature} backend.
+To provide users with the ability to sign using a simple image saved on the computer, which is probably a scanned signature, or signature drawn directly in our application, we created what we call \textit{simple image signature} backend.
 It is useful when users do not have the resources required to sign the document using the QES, or in situations when they would rather not do so.
@@
 Electron applications have a main thread that is running a Node.js\footnote{Node.js is a JavaScript runtime for use outside of the browser with easier access to the OS APIs.} and a UI thread running the relevant part of the Chromium\footnote{Chromium is a popular open-source browser project, see \url{https://www.chromium.org/Home}.}.
-Their communication is done via \pojem{Inter-process communication} (IPC) and allows access to the Node.js APIs in the UI thread.
+Their communication is done via \pojem{inter-process communication} (IPC) and allows access to the Node.js APIs in the UI thread.
 Considering the JavaScript is interpreted at runtime, we are exposing ourselves to potential attacks (mostly XSS\footnote{XSS means Cross-site scripting, see \url{https://owasp.org/www-community/attacks/xss/}.}).
@@
 
-When approaching automated testing, we have decided to implement the testing pyramid as defined by the Mike Cohn\footnote{Mike Cohn came up with this concept in his book \textit{Succeeding with Agile}.}.
-One of the implementations of this idea was done by the Ham Vocke in the article \textit{The Practical Test Pyramid}\footnote{Available on the website of Martin Fowler at \url{https://martinfowler.com/articles/practical-test-pyramid.html}.}.
-It is implemented at a scale at Google, which explains their positive experience with it\footnote{Approach outlined in the article \textit{Just Say No to More End-to-End Tests} available at \url{https://testing.googleblog.com/2015/04/just-say-no-to-more-end-to-end-tests.html}.}.
+When approaching automated testing, we have decided to implement the testing pyramid as defined by Mike Cohn in \cite{succeedingagile}.
+One of the implementations of this idea was done by the Ham Vocke in \cite{practicalpyramid}.
+It is implemented at a scale at Google, which explains their positive experience with it in \cite{nomoree2e}.
 Following this approach, we write three types of tests: end-to-end, integration, and unit tests.
 The ratio of these types of tests should form a pyramid where the heavy bottom is formed by the unit tests, followed by a smaller set of component tests forming the middle of the pyramid, and a small set of end-to-end tests forming the tip of the pyramid.
-The motivation behind this approach is to write more tests that are faster and break less often (as they are more isolated) as opposed to tests that are slower and break more often.
-
-In our case, we have decided to write unit tests in the Jest\footnote{Jest is a testing framework and runner, see \url{https://jestjs.io/}.} testing both the UI and main thread code, component tests in Jest for the main thread and Cypress\footnote{Cypress is a testing framework and runner using the Chromium, see \url{https://www.cypress.io/}.} for the UI thread combined with the Percy\footnote{Percy is a SaaS for visual testing, see \url{https://percy.io/}.}, and end-to-end tests as Spectron tests which run and interact with the actual distributables (executable files).
+The motivation behind this approach is to write more tests that are faster and break less often due to better isolation, as opposed to tests that are slower and break more often.
+
+In our case, we have decided to write unit tests in Jest testing both the UI and main thread code, component tests in Jest for the main thread and Cypress for the UI thread combined with the Percy, and end-to-end tests as Spectron tests which run and interact with the actual distributables (executable files).
 We are able to fully isolate each type of tests because of our architecture that allows easy mocking.
@@
 
-Our code is kept in a git repository hosted at Github and is a source for the Continuous Integration (CI) and Continuous Delivery (CD) pipeline.
-For the CI/CD, we are using Azure Pipelines\footnote{Cloud service from Microsoft that is part of the Azure DevOps, see \url{https://docs.microsoft.com/en-us/azure/devops/pipelines/get-started/what-is-azure-pipelines?view=azure-devops}.} specifically, that are free for OSS.
+Our code is kept in a git repository\footnote{Repository is available at \url{https://github.com/durasj/octosign}.} hosted at Github and is a source for the Continuous Integration (CI) and Continuous Delivery (CD) pipeline.
+For the CI/CD, we are using Azure Pipelines specifically, that are free for OSS.
 On each push, apart from setting up the environment, we perform static analysis on the code checking the type safety using TypeScript, code style using the ESLint and Prettier, code quality using the CodeClimate.com, and check for potential vulnerabilities using the Snyk.io.
@@
 Build process build individual backends, the main thread, UI thread, and API code, and uploads bundled distributable files for all platforms back to the Github.
-The whole process runs for each platform - Windows, Linux, and Mac - separately; therefore, there is no need for cross-compiling and running of platform-dependent tools in virtualized environments.
+The whole process runs for each platform - Windows, Linux, and macOS - separately; therefore, there is no need for cross-compiling and running of platform-dependent tools in virtualized environments.
 Publishing of the new release is then done by saving the release in the Github.
@@
 Downloads are available for each major desktop platform in a format suitable for installation or integration in the operating system.
-For Windows, we are distributing an \textit{.exe} installer working on all supported 64-bit Windows versions.
-For Linux, a portable \pojem{AppImage} and a \textit{.deb} package for Debian-like distributions.
-For Mac OS X, a \textit{.dmg} disk image.
+For Windows, we are distributing an \texttt{EXE} installer working on all supported 64-bit Windows versions.
+For Linux, a portable \texttt{AppImage} and a \texttt{DEB} package for Debian-like distributions.
+For macOS, a \texttt{DMG} disk image.
 
@@
 To improve on this further, we are doing a \pojem{code-signing} for each release using a trusted way for each platform.
-On Mac OS X, this means enrolling in the Apple Developer Program that requires an annual fee and installation of signing certificates.
-Additionally, it is necessary to \textit{notarize} the application for use on Mac OS X.
+On macOS, this means enrolling in the Apple Developer Program that requires an annual fee and installation of signing certificates.
+Additionally, it is necessary to \textit{notarize} the application for use on macOS.
 Notarization is an automated process required by Apple to prove that the application does not contain any malicious code and is properly signed and configured.
@@
 On modern versions of Windows, applications that are not code-signed can trigger a warning or, depending on the settings, even completely prevent the user from using software that is not code-signed using a certificate that is cosigned with a trusted root certificate.
-Certification for individuals on the Mac OS seems to be fully automated, does not involve lengthy verification, and costs 99 USD a year.
+Certification for individuals on the macOS seems to be fully automated, does not involve lengthy verification, and costs 99 USD a year.
 Certification on Windows is offered by various trusted third parties like DigiCert.com that we have used and involves lengthy, partially automated verification at a price starting from 60 USD a year.
@@
 
-These pages are written in a Markdown\footnote{Markdown is a lightweight markup language, see \url{https://commonmark.org/}.}, making it more accessible for editing by the community that does not need to understand specific technologies used to display the content on the website.
+These pages are written in markdown - a lightweight markup language, making it more accessible for editing by the community that does not need to understand specific technologies used to display content on the website.
 
@@
 
-From the technological point of view, the website is a \pojem{Single Page Application} (SPA) written in React\footnote{React is a JavaScript library for building user interfaces, see \url{https://reactjs.org/}.}.
-It is built using Gatsby\footnote{Gatsby is a free and open-source framework based on React, see \url{https://www.gatsbyjs.org/}.}.
+From the technological point of view, the website is a \pojem{Single Page Application} (SPA) using Gatsby written in React.
 Even though it is a SPA, it is \pojem{prerendered} to static HTML and CSS files that are served from CDN, which improves performance and allows for better \pojem{Search Engine Optimization} (SEO).
@@
 
-Different language mutations are denoted by using a suffix, for example, the home page is written in English and Slovak using file names \textit{home.en.tsx} and \textit{home.sk.tsx}.
-Same applies for content written in markdown, for example, help page on basic usage is written in files \textit{basic.en.md} and \textit{basic.sk.md}.
-The URL of the website contains language code of the current language if it is other than English, for example, the download page is available in English on path \textit{/download} and in Slovak on path \textit{/sk/download}.
+Different language mutations are denoted by using a suffix, for example, the home page is written in English and Slovak using file names \texttt{home.en.tsx} and \texttt{home.sk.tsx}.
+Same applies for content written in markdown, for example, help page on basic usage is written in files \texttt{basic.en.md} and \texttt{basic.sk.md}.
+The URL of the website contains language code of the current language if it is other than English, for example, the download page is available in English on path \texttt{/download} and in Slovak on path \texttt{/sk/download}.
 By switching the language using the selector on the upper right, the URL changes to the required format.
 
-TODO: If you mind this blank space, I can add a scheme with source files in various languages > gatsby > CDN flow.
-
 \zaver
@@
 We verified that the main legal and technical preconditions should be met for the use of electronic signatures.
-They should be, therefore, a viable alternative to handwritten signatures.
+We therefore believe, they are a viable alternative to handwritten signatures.
 
@@
 Overili sme, že hlavné predpoklady pre použitie elektronických podpisov by mali byť naplnené.
-Mali by byť preto schodnou alternatívou pre vlastnoručné podpisy.
+Veríme, že sú preto schodnou alternatívou pre vlastnoručné podpisy.
 
@@
 
-Preto sme navrhli jednoduchú otvorenú softvérovú platformu, ktorá sa snaží byť priamočiara pre koncového používateľa výberom pravdepodobných predvolených hodnôt a znížením množstva informácií s ktorými sa používateľ stretáva.
+Preto sme navrhli jednoduchú otvorenú softvérovú platformu, ktorá sa snaží byť priamočiara pre koncového používateľa výberom pravdepodobných predvolených hodnôt a znížením množstva informácií, s ktorými sa používateľ stretáva.
 Taktiež sme uľahčili preklad našej aplikácie pomocou webového rozhrania tretej strany, s ktorým môže ktokoľvek z komunity pomôcť prekladmi.
 Naša aplikácia je dostupná v troch jazykoch: angličtine, slovenčine a češtine s nemčinou a španielčinou v prekladaní.
-Je tiež plne modulárna s rozdelením hlavného modulu, ktorý obsahuje používateľské rozhranie, a modulov vykonávajúcich manipuláciu s dokumentmi.
-Implementovali sme modul na podpisovanie pomocou kvalifikovaného elektronického podpisu, ktorý spĺňa európske štandardy a mal by byť v súlade s európskymi zákonmi, a modul, ktorý umožňuje podpisovanie pomocou naskenovaného alebo nakresleného podpisu.
+Je tiež plne modulárna, kde hlavný modul obsahuje používateľské rozhranie a niekoľko modulov vykonáva manipuláciu s dokumentmi.
+Implementovali sme modul na podpisovanie pomocou kvalifikovaného elektronického podpisu, ktorý spĺňa európske štandardy a mal by byť v súlade s európskymi zákonmi a modul, ktorý umožňuje podpisovanie pomocou naskenovaného alebo nakresleného podpisu.
 Navrhovaný softvér je platforma, na ktorej môžu ostatní rýchlo vybudovať nové spôsoby podpisovania dokumentov v súlade s rôznymi zákonmi alebo založiť ďalší výskum.
@@
 Na druhej strane existuje priestor na zlepšenie flexibility a jednoduchosti prvotného nastavenia nášho softvéru.
-Vytvorili sme tiež podpornú webovú stránku s možnosťou stiahnutie pre hlavné počítačové operačné systémy, pomocou a všeobecnými informáciami o elektronických podpisoch.
-Náš projekt by mal byť otvorený ľuďom z rôznych krajín, rozprávajúcich rôznymi jazkymi, a zmenám od ľudí s rôznymi zručnosťami - nie len programátorom.
+Vytvorili sme tiež podpornú webovú stránku s možnosťou stiahnutia pre hlavné počítačové operačné systémy, užívateľskou príručkou a všeobecnými informáciami o elektronických podpisoch.
+Náš projekt je otvorený ľuďom z rôznych krajín, rozprávajúcich rôznymi jazkymi a zmenám od ľudí s rôznymi zručnosťami - nie len programátorom.
 
@@
 
-  %  ------
-  % Article
-  %  ------
   \bibitem{law}
@@
 
+  %  ------
+  % Crypto
+  %  ------
   \bibitem{cryptotxtbook}
@@
 
+  %  ------
+  % Tech
+  %  ------
   \bibitem{sommerville}
   \osoba{SOMMERVILLE, Ian}, 2016. \emph{Software Engineering, 10th Edition}. Harlow: Pearson Education. ISBN 978-0-13-394303-0.
+
+  \bibitem{succeedingagile}
+  \osoba{Cohn, Mike}, 2009. \emph{Succeeding with Agile}. Boston: Addison-Wesley Professional. ISBN 978-0-32-166053-4.
+
+  \bibitem{practicalpyramid}
+  \osoba{Vocke, Ham}, 2018. \emph{The Practical Test Pyramid} [online]. Available at \url{https://martinfowler.com/articles/practical-test-pyramid.html}
+
+  \bibitem{nomoree2e}
+  \osoba{Wacker, Mike}, 2015. \emph{Just Say No to More End-to-End Tests} [online]. Available at \url{https://testing.googleblog.com/2015/04/just-say-no-to-more-end-to-end-tests.html}
 
@@
 
-Open the subpage Download on website \url{https://octosign.com} where you can download the application by clicking on the button for your operating system.
+Open the subpage \textit{Download} on website \url{https://octosign.com} where you can download the application by clicking on the button for your operating system.
 After the download is complete, open the file and follow the instructions.
@@
 First, it is needed to select a document. To select a document, click on the button \textit{SELECT FILES} or drag and drop your document on the application.
-The recommended document type is PDF.
+The recommended document type is PDF, although other file types are supported as well.
 After selecting the document, a card with the document name will appear just like in the picture.

### 2020-05-18 paper/main.tex
Fix suggestions from tutor

@@
 
-With regards to electronic signatures, law within the EU used to differ, with the law applicable in the SR being now-repealed \textit{Act No.
+Regarding the electronic signatures, law within the EU used to differ, with the law applicable in the SR being now-repealed \textit{Act No.
 215/2002 Coll.}.
@@
 Both when we are the ones using someone else's software and when someone else is using our software.
-Generally speaking, the more open the license is, the higher the probability it will not prevent adoption for legal reasons.
+Generally speaking, the more open the license is, the more likely it is it will not hinder adoption for legal reasons.
 We provide examples of categorized licenses in Figure \ref{fig:license-comparison}.

### 2020-05-18 paper/main.tex
Add proofreading fixes of abstract, intro, conclusion

@@
 With the recent changes in the legal status of electronic signatures in the world, there is an opportunity to expand the use of electronic signatures as an alternative to handwritten signatures outside the usual areas.
-We devote our attention to the exploration of legal preconditions relevant for the EU, technological preconditions, review of the existing software, and proposal of our own software platform.
+In this work, we devote our attention to the exploration of the legal preconditions relevant for the EU, technological preconditions, review of the existing software, and a proposal of our own software platform.
 The result is an open software platform for operating systems Windows, macOS, and Linux.
-Using the application, it is possible, following the eIDAS Regulation, to sign and verify the signature of any document with the legal effect of an attested signature.
+Using this application, it is possible, following the eIDAS Regulation, to sign and verify the signature of any document with the legal effect of an attested signature.
 Our solution supports ID cards of the Slovak Republic with the eID chip without any modifications or settings, which enables use outside the web applications of the state.
-The application is not tied only to the Slovak eID but is ready for further use with various devices in different countries and as a platform for other types of signatures.
-Another part is a website that, in addition to the possibility of downloading software and brief documentation, provides general information about electronic signatures.
-The emphasis is on simple extensibility of the program via language translations, documentation, functionality, and general information.
-The thesis can be further extended with novel ways of signing or by improving flexibility, which is not up to par with compared software.
+The application is not tied only to the Slovak eID but is ready for further use with various devices in other countries and as a platform for other types of signatures.
+There is also a website that, in addition to the possibility of downloading software and brief documentation, provides general information about electronic signatures.
+The emphasis is on the simple extensibility of the program via language translations, documentation, functionality, and general information.
+The thesis can be further extended with novel ways of signing or by improving flexibility, which is not completely up to par with compared software.
 }
@@
 Their practical usability, however, is influenced by their legal status and available infrastructure.
-With the recent changes in the relevant law around the world and the European Union specifically, they see increased adoption.
-
-While, specifically in the European Union, electronic signatures seem to be relatively extensively used in communication with the government, they do not seem to be widely adopted in the exchange between individuals or smaller organizations, which still seem to prefer handwritten signatures.
-In many cases, however, they may be unpractical for electronic communication due to the need of physical presence or printing and sending of documents.
-
-Our goal is to explore an accessible way to use electronic signatures.
-In the beginning, we go over the preconditions that make it a viable alternative to the handwritten signatures.
-We cover the legal preconditions relevant for the European Union and Slovakia in particular.
-Also, we explain technical preconditions focusing on the cryptographic concepts like asymmetric cryptography, timestamping, hashing, and standards that comply with the law within the European Union.
+With the recent changes to the relevant laws around the world, and in the European Union specifically, they are seeing increased adoption.
+
+While, specifically in the European Union, electronic signatures seem to be used relatively extensively in communication with the government, they do not seem to have been widely adopted in exchange between individuals or smaller organizations, which still seem to prefer handwritten signatures.
+In many cases, however, they may be impractical for electronic communication, due to the need of a person's physical presence or the printing and sending of documents.
+
+Our goal is to explore an accessible way of using electronic signatures.
+In the beginning, we go over the preconditions that make it a viable alternative to handwritten signatures.
+We cover the legal preconditions relevant to the European Union, and Slovakia in particular.
+Also, we explain technical preconditions, focusing on the cryptographic concepts like asymmetric cryptography, timestamping, hashing, and standards that comply with the laws of the European Union.
 Secondly, we do a review of the existing software, where we briefly cover several applications that are meant for basic use.
 We compare them using several criteria to find out the strengths and weaknesses of the current implementations and identify a potential gap.
-Finally, we devote our attention to our own implementation.
+Finally, we devote our attention to our own software's implementation.
 We explain the proposed desktop software and go into details on software engineering aspects.
 Specifically, we cover usability, localization, modularity, architecture, testing, and automation.
-In the end, we compare it with the existing software and identify possible future improvements.
-Another result we focus on is a supporting website and how it helps with the use of the software.
-All in a way that is open for extensibility by the community.
+In the end, we compare it to the existing software and identify possible future improvements.
+Another result we focus on is a supporting website and how it helps with the use of the software, all in a way that is open for extension by the community.
 
@@
 We verified that the main legal and technical preconditions should be met for the use of electronic signatures.
-We therefore believe, they are a viable alternative to handwritten signatures.
-
-We did a review of the existing software focusing on the licensing, cross-platform compatibility, localization, file support, ease of setup, flexibility, verification ability, and extensibility.
-We identified the positive and negative sides of the current implementations and found a gap in the software for signing using the qualified electronic signatures.
+We therefore believe they are a viable alternative to handwritten signatures.
+
+We did a review of the existing software, focusing on the licensing, cross-platform compatibility, localization, file support, ease of setup, flexibility, verification ability, and extensibility.
+We identified the positive and negative sides of the current implementations and found a gap in the software for signing using qualified electronic signatures.
 
 Therefore, we proposed a simple open-source software platform that aims to be straightforward for the end-user by choosing the most likely defaults and reducing the amount of information the user has to interact with.
-Also, we made our application easy to translate using a third-party web UI where anyone from the community can contribute with translations.
+Also, we made our application easy to translate using a third-party web UI where anyone from the community can contribute translations.
 Our application is available in three languages: English, Slovak, and Czech, with German and Spanish in translation.
 It is also fully modular, dividing the main module that contains the UI and modules performing the document manipulation.
-We implemented a module for signing using qualified electronic signatures that follow the European standards and should be compliant with the European law, and a module that allows signing using a scanned or drawn signature.
-The proposed software is a platform on which others can quickly build new ways to sign documents complying with different laws or to base further research.
-From the software engineering perspective, the implementation is robust, and we covered testing and automation in connection to Electron applications like ours.
+We implemented a module for signing using qualified electronic signatures that follow European standards and should be compliant with European law, and a module that allows signing using a scanned or drawn signature.
+The proposed software is a platform on which others can quickly build new ways of signing documents complying with different laws or basing further research.
+From a software engineering perspective, the implementation is robust, and we covered testing and automation in connection with Electron applications like ours.
 On the other hand, there is some room for improvement in flexibility and ease of setup of our software.

### 2020-05-18 paper/main.tex
Fix additional proofreading issues

@@
 The emphasis is on the simple extensibility of the program via language translations, documentation, functionality, and general information.
-The thesis can be further extended with novel ways of signing or by improving flexibility, which is not completely up to par with compared software.
+The thesis can be further extended with novel ways of signing or by improving flexibility, which is not completely up to par with the compared software.
 }
@@
 
-While, specifically in the European Union, electronic signatures seem to be used relatively extensively in communication with the government, they do not seem to have been widely adopted in exchange between individuals or smaller organizations, which still seem to prefer handwritten signatures.
+While, specifically in the European Union, electronic signatures seem to be used relatively extensively in communication with the government, they do not seem to have been widely adopted in exchanges between individuals or smaller organizations, which still seem to prefer handwritten signatures.
 In many cases, however, they may be impractical for electronic communication, due to the need of a person's physical presence or the printing and sending of documents.
@@
 We cover the legal preconditions relevant to the European Union, and Slovakia in particular.
-Also, we explain technical preconditions, focusing on the cryptographic concepts like asymmetric cryptography, timestamping, hashing, and standards that comply with the laws of the European Union.
+Also, we explain the technical preconditions, focusing on cryptographic concepts like asymmetric cryptography, timestamping, hashing, and standards that comply with the laws of the European Union.
 Secondly, we do a review of the existing software, where we briefly cover several applications that are meant for basic use.
@@
 
-We did a review of the existing software, focusing on the licensing, cross-platform compatibility, localization, file support, ease of setup, flexibility, verification ability, and extensibility.
+We did a review of the existing software, focusing on licensing, cross-platform compatibility, localization, file support, ease of setup, flexibility, verification ability, and extensibility.
 We identified the positive and negative sides of the current implementations and found a gap in the software for signing using qualified electronic signatures.

### 2020-05-18 paper/main.tex
Add final changes to the paper and info about thesis submission

@@
 
+\iffalse % This was not yet stable at the time of publishing
 We also support adding a visible signature for PDFs.
@@
 Labels are translated using the current language of the application as we assume this will be the language the receiver will most likely speak.
+\fi
 
@@
 
-To improve on this further, we are doing a \pojem{code-signing} for each release using a trusted way for each platform.
+To improve on this further, we tried to do a \pojem{code-signing} for each release using a trusted way for each platform.
 On macOS, this means enrolling in the Apple Developer Program that requires an annual fee and installation of signing certificates.
@@
 At the time of writing, there does not seem to be any service offering code-signing certificates to OSS for free.
+We were not able to finish the process on time for the publishing of this thesis - we feel like this process could be improved in the future for OSS.
 
